### 2015-10-18 PPM.tex
Yumeng's PPM description and test case (solid body and stirring) description

\paragraph{}
The dimensionally splitting scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian by \citep{Lin1996}. This dimensionally splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensional.
\section{Long time-step permitting PPM} 
\label{sec:PPM} 
\paragraph{} 
Both PPM and the long time steps technique use subgrid cell reconstruction. Although they are defined for varying grid spacing, we will only describe them for uniform grid interval, $\Delta x$. Long time-step permitting PPM is described on a domain with length $L$, which is divided into $n$ grid points. The position of $j$th grid is $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots, n-1$ and the corresponding tracer is $\phi_j$. Subgrid cell reconstruction uses the space average mass within the $j$th grid box at time $t_n$, which is defined as: 
\begin{equation} \label{eq:2.1} 
\phi^n_j = \frac{1}{\Delta x} \int^{x_{j+\frac{1}{2}}}_{x_{j-\frac{1}{2}}} \phi (x,t)dx
\end{equation}
where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid point. Thus, $x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$.
\paragraph{} 
The long time steps technique \citep{Leonard1995} calculates the flux through the edge of boundaries by transforming the time integral to a space integral in terms of the analytical solution of the advection equation and using the relationship $u\Delta t = c\Delta x$ and $u$dt = dx, where $c$ is the Courant number. This reveals that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$, and can be written as the following equation:
\begin{equation} \label{eq:2.2} 
c\phi_{j-\frac{1}{2}} = \frac{c}{\Delta t} \int^{\Delta t}_{0} \phi (x_j - \frac{\Delta x}{2}, t)dt = \frac{1}{\Delta x} \int^{x_j-\frac {\Delta x}{2}}_{x_j-\frac {\Delta x}{2}-c\Delta x} \phi (x)dx 
\end{equation} 
\paragraph{} 
\begin{figure}
% \vspace{-20pt}
\centering
\includegraphics[scale=0.6]{YumengsGraphics/nirvana.pdf}
% \vspace{-20pt}
\caption{The 1D long time permitting technique. The mass between $x_{j,L}$ and $x_{j-c_N-1,R}$ can be calculated using equation \ref{eq:2.3}, and the mass between $x_{j-c_N-1,R}$ and $x_{j-c}$ should be calculated using PPM \label{fig:2.1}}
\end{figure}
The cumulative mass from the start point to the $j$th left zone edges is defined as $M_{j-\frac{1}{2}}$ and can be estimated by the definition of $\phi_j$ in equation \ref{eq:2.1} and can be discretized as follows \citep{Colella1984}:
\begin{equation} \label{eq:2.3}
M_{j-\frac{1}{2}} = \sum_{k < j} \phi_{k}\Delta x
\end{equation} 
If the Courant number is an integer $c_N$, the flux at left boundary can be calculated directly using the difference between the cumulative mass at $x_{j-\frac{1}{2}}$, $M_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}-c_N}$, $M_{j-c_N-\frac{1}{2}}$. However, because the Courant number is not necessarily integer, the equation \ref{eq:2.2} is unable to be applied to all conditions. Hence, the Courant number is divided into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$. The Courant number thus becomes $c = c_N + c_r$ \citep{Leonard1995}. The long time step permitting technique is illustrated in figure \ref{fig:2.1}. Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$. In this way, the CFL restriction is removed, because the remnant Courant number is always less than one. The flux equation becomes:
\begin{equation} 
c\phi_{j-\frac{1}{2}} = M_{j-\frac{1}{2}} - M_{j-c_N-\frac{1}{2}} + \frac{1}{\Delta x} \int^{x_j-\frac {\Delta x}{2}-c_N\Delta x}_{x_j-\frac {\Delta x}{2}-c\Delta x} \phi (x)dx 
\end{equation} 
\paragraph{}
The remaining problem is to calculate the component of the flux related to the remnant Courant number, $c_r$, using PPM. PPM assumes that the profile at each interval is a parabola and thus can be represented by a second order polynomial \citep{Colella1984}. The equation can be achieved using $\phi(\xi = 0) = \phi_L$, $\phi(\xi = 1) = \phi_R$ and $ \int^{1}_{0} \phi (\xi) d\xi = \phi_j  $, which is obtained by \citet{Colella1984} as:
\begin{equation} \label{eq:2.4} 
\phi(x) = \phi_{L,j} + \xi (\Delta \phi_j + \phi_{6,j}(1-\xi)) 
\end{equation} 
where $ \xi = \frac{x-x_{j-\frac{1}{2}}}{\Delta x} $, $ \Delta \phi_j = \phi_{R,j} - \phi_{L,j} $, $ \phi_{6,j} = 6(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j})) $ and $\phi_{L,j}$ and $\phi_{R,j}$ represent the left and right boundary of $j$th interval respectively. Therefore, the flux can be calculated directly using the integral of $\phi(x)$
\paragraph{} 
From equation \ref{eq:2.4}, the boundary values of $\phi$ are required. A quartic polynomial mass interpolation \citep{Colella1984} is implemented to calculate $M_{j+\frac{1}{2}}$. Thus, the tracer value at the right boundary, $\phi_{j+\frac{1}{2}}$, can be calculated by differentiating $M_{j+\frac{1}{2}} $, the resulting formula is:
\begin{equation} \label{eq:2.5} 
\phi_{j+\frac{1}{2}} = \phi_j + \frac{1}{2}(\phi_{j+1}-\phi_{j}+\frac{\delta_m \phi_j - \delta_m \phi_{j+1}}{6}) 
\end{equation} 
where $\delta_m \phi_j$ is the slope in the $j$th interval and is calculated by:
\begin{equation}
\delta \phi_j = \frac{1}{2}(\phi_{j+1} - \phi_{j-1})
\end{equation}
% \begin{equation} \label{eq:2.6}
% \delta_m \phi_j =
%   \begin{cases}
%     \min(|\delta \phi_j|, 2|\phi_j - \phi_{j-1}|,2|\phi_{j+1} - \phi_{j}|)\text{sgn}(\delta \phi_j)    & (\phi_{j+1}-\phi_j)(\phi_j-\phi_{j-1})>0\\
%     0  & \text{ otherwise}
%   \end{cases}
% \end{equation}
% where $\delta \phi_j = \frac{1}{2}(\phi_{j+1} - \phi_{j-1})$.  
% \paragraph{} 
% Godunov's theorem shows that higher than first order accurate linear schemes cannot maintain monotonicity \citep{Durran2010} and PPM uses quartic interpolation, which is fourth order accurate. Hence, there might be spurious oscillations in the numerical solutions and flux limiters/filters can significantly reduce those oscillation and preserve monotonicity. Equation \ref{eq:2.6} is a limiter to guarantee monotonicity and can result in a steeper representation of tracer profile around discontinuity, otherwise $\delta \phi_j$ should be directly used in equation \ref{eq:2.5}. Some other constraints are also implemented by \cite{Colella1984} for $\phi_{R,j}$ and $\phi_{L,j}$ to prevent the spurious oscillations. For smooth profile, $\phi_{j+\frac{1}{2}} = \phi_{R,j} = \phi_{L,j+1}$. For discontinuous profiles, some limiters are applied. One of the limiters is: 
% \begin{eqnarray}  \label{eq:2.7}
% \phi_{L,j} = \phi_{L,j}(1-\eta_j) +\eta_j (\phi_{j-1}+\frac{1}{2} \delta_m \phi_{j-1}) \nonumber \\
% \phi_{R,j} = \phi_{R,j}(1-\eta_j) +\eta_j (\phi_{j+1}-\frac{1}{2} \delta_m \phi_{j+1})
% \end{eqnarray}
% where $\eta_j = \max(0,\min(\eta^{(1)}(\widetilde \eta_j-\eta^{(2)}),1)$.
% \paragraph{}
% $\widetilde \eta_j$ can be calculated by:
% \begin{eqnarray} \label{eq:2.8}
% \widetilde \eta_j &= -\frac{\phi_{j+2}-3\phi_{j+1}+3\phi_j-\phi_{j-1}}{6(\phi_{j+1}-\phi_{j-1})}\\
% &\quad \text{if } -\delta^2 \phi_{j+1} \delta^2 \phi_{j-1}>0, |\phi_{j+1}-\phi_{j-1}|-\varepsilon \min(|\phi_{j+1}|,|\phi_{j-1}|)>0\\
% &=0 \\&\quad \text{ otherwise,}
% \end{eqnarray}
% where $\delta^2 \phi_j = \frac{\phi_{j+1}-2\phi_{j}+\phi_{j-1}}{6\Delta x}$. The $\eta^{(1)}=20$,$\eta^{(2)}=0.05$ and $\epsilon=0.05$, which shows the level of discontinuity. These values ensure that the small jumps in discontinuity are not applied by such limiters and can maintain high order of accuracy. Because these constraints use a method with lower order of accuracy to reprensent the discontinuities, this could reduce the overall order of accuracy of high order polynomial interpolation \citep{Lauritzen2007}.
% \paragraph{}
% Another problem is that the polynomial interpolation could lead to unphysical estimation of left and right boundaries. Under this circumstance, other constraints should be applied \citep{Colella1984}. Without these constraints, the tracer at left boundaries and right boundaries would be much higher or lower than the real value.  The resulting $\phi_{L,j}$ and $\phi_{R,j}$ should be:
% \begin{eqnarray} \label{eq:2.9}
% &\phi_{R,j}\rightarrow 3\phi_j -2\phi_{L,j} &\text{if }& -\frac{(\phi_{R,j}-\phi_{L,j})^2}{6} >(\phi_{R,j} -\phi_{L,j})(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j})) \nonumber\\
% &\phi_{L,j}\rightarrow 3\phi_j -2\phi_{R,j} &\text{if }& (\phi_{R,j} -\phi_{L,j})(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j}))>\frac{(\phi_{R,j}-\phi_{L,j})^2}{6} \\
% &\phi_{L,j}\rightarrow \phi_j\text{, } \phi_{R,j}\rightarrow\phi_j &\text{if }& (\phi_{R,j} - \phi_j)(\phi_j-\phi_{L,j}) \leqslant 0 \nonumber
% \end{eqnarray}
\paragraph{}
The flux thus can be calculated in terms of the tracer at the boundaries of each interval. In terms of different upstream directions, the mass flux at the right boundary of $j$th interval can be written as:
\begin{eqnarray} \label{eq:2.10}
M_{c_r,R,j} = c_r \Delta x (\phi_{R,j}-\frac{c_r}{2}(\delta \phi_j-(1-\frac{2}{3} c_r)\phi_{6,j})) & \text{ if $c_r$}\geqslant 0 \nonumber \\
M_{c_r,R,j} = c_r \Delta x (\phi_{L,j+1}-\frac{c_r}{2}(\delta \phi_{j+1}-(1+\frac{2}{3} c_r)\phi_{6,j+1})) & \text{ if  $c_r$} < 0
\end{eqnarray}
\paragraph{}
For the complete Courant number, where the integer part of Courant number $c_N$ and remnant part $c_r$ is combined, the mass flux should be calculated by the following formulas:
\begin{eqnarray} \label{eq:2.11}
M_{R,j} = M_{c_r,R,j} + M_{j+\frac{1}{2}}-M_{j-N+\frac{1}{2}} & \text{c}>0 &\nonumber \\
M_{R,j} = -M_{c_r,R,j} +M_{j+N+\frac{1}{2}}- M_{j+\frac{1}{2}} & \text{c}<0 & \\
M_{R,j} = 0 & \text{otherwise} \nonumber 
\end{eqnarray}
Therefore, for westerly flows, the formula for next time step $\phi^{n+1}_j$ is: 
\begin{equation} \label{eq:2.12}
\phi^{n+1}_j = \phi^n_j + \frac{M^n_{in,j} - M^n_{out,j}}{\Delta x} = \phi^n_j + \frac{M^n_{R,j} - M^n_{R,j+1}}{\Delta x}
\end{equation}
where the subscript means the inflow and outflow of each cell.

\section{COSMIC Splitting}
\label{sec:COSMIC}
\paragraph{}
COSMIC splitting is an operator splitting scheme, which is used so that PPM can be implemented in two dimensions. Three of the properties expected of multidimensional advection schemes are: stability, conservation and constancy preservation \citep{Lin1996}. We will see how COSMIC splitting \citep{Leonard1996} maintains these three properties.
\begin{figure}
\vspace{-10pt}
\centering
\includegraphics[scale=0.6]{YumengsGraphics/conservative_form.pdf}
\vspace{-20pt}
\caption{The position of each flux and the position of $\phi$, the subscript w/e/n/s denote the direction of the flux relative to the position of $\phi$ \label{fig:2.2}}
\end{figure}
Operator splitting can use conservative-form or advective-form operators. Use of conservative-form operator with operator splitting leads to inherently mass conservation but cannot satisfy the constancy condition. The $x$ and $y$ direction conservative-form operator can be defined:
\begin{eqnarray}  \label{eq:2.13}
&X_{\scriptscriptstyle C}(\phi ) = c_{w}\phi_w(\phi )-c_e\phi_e(\phi ) 
&Y_{\scriptscriptstyle C}(\phi ) = c_{n}\phi_n(\phi )-c_s\phi_s(\phi )
\end{eqnarray}
where $c$ is the Courant number and the subscripts represent the direction of the boundary of each grid and w/e/n/s denote the compass direction relative to $\phi$. $\phi $ in the brackets means flux at the boundary $c\phi$ is calculated from cell average $\phi$ using the one-dimensional long time permitting PPM. As shown in figure \ref{fig:2.2}, this is an Arakawa C staggered grid. The subscript $C$ represents the conservative-form operator.
\paragraph{}
The conservative-form operator update of $\phi$ can thus be calculated as:
\begin{eqnarray} \label{eq:2.14}
&\phi^n_{\scriptscriptstyle CY} = \phi^n +Y_{\scriptscriptstyle C}(\phi^n)
&\phi^{n+1}_{\scriptscriptstyle CYX} = \phi^n_{\scriptscriptstyle CY} +X_{\scriptscriptstyle C}(\phi^n_{\scriptscriptstyle CY})
\end{eqnarray}
or
\begin{eqnarray*} 
&\phi^n_{\scriptscriptstyle CX} = \phi^n +X_{\scriptscriptstyle C}(\phi^n) 
&\phi^{n+1}_{\scriptscriptstyle CXY} = \phi^n_{\scriptscriptstyle CX} +Y_{\scriptscriptstyle C}(\phi^n_{\scriptscriptstyle CX})
\end{eqnarray*}
The conservative-form update calculates the flux at the boundary of each grid, which means it is necessary to use the velocity at boundaries.  By this means, the flux at each boundary is unique, maintaining conservation. 
\paragraph{}
However, this cannot satisfy constancy condition. If the initial tracer $\phi $ is homogeneous, the conservative-form $x$-direction update becomes:
\begin{equation} \label{eq:2.15}
\phi^{n+1}_{\scriptscriptstyle CX} = \phi^n(1 +c_{w}-c_e)
\end{equation}
The results destroy constancy condition, because in a 2D non-divergent velocity field, $c_{w}-c_e$ could be unequal. 
\paragraph{}
Advective-form operator splitting is another way to treat the advection scheme. Unlike the conservative-form, the advective-form splitting is not inherently conservative but can satisfy constancy condition. The advective-form operator can be defined as:
\begin{eqnarray} \label{eq:2.16}
&X_{\scriptscriptstyle A}(\phi ) = c^{cell}(\phi_w(\phi )-\phi_e(\phi )) 
&Y_{\scriptscriptstyle A}(\phi ) = c^{cell}(\phi_n(\phi )-\phi_s(\phi ))
\end{eqnarray}
where $c^{cell}$ is the cell centred Courant number, or the transverse velocity and the subscript $A$ denote advective-form operator. The cell centred Courant number should be ideally the average of two faces. However, \citep{Lin1996} pointed out that this could generate spurious flux, so the transverse velocity should be chosen as the upwind velocity. If the transverse velocity is in the opposite directions at two faces, the cell centred velocity should set to zero.
\paragraph{}
The advective-form update is:
\begin{eqnarray} \label{eq:2.17}
&\phi^{n}_{\scriptscriptstyle AY} = \phi^{n} +Y_{\scriptscriptstyle A}(\phi^{n}) 
&\phi^{n+1}_{\scriptscriptstyle AYX} = \phi^{n}_{\scriptscriptstyle AY} +X_{\scriptscriptstyle A}(\phi^{n}_{\scriptscriptstyle AY})
\end{eqnarray}
or
\begin{eqnarray*}
&\phi^{n}_{\scriptscriptstyle AX} = \phi^{n} +X_{\scriptscriptstyle A}(\phi^{n}) 
&\phi^{n+1}_{\scriptscriptstyle AXY} = \phi^{n}_{\scriptscriptstyle AX} +Y_{\scriptscriptstyle A}(\phi^{n}_{\scriptscriptstyle AX})
\end{eqnarray*}
\paragraph{}
If the initial tracer $\phi $ is unity, the advective-form update becomes:
\begin{equation} \label{eq:2.18}
\phi^{n+1}_{\scriptscriptstyle AX} = \phi^{n}+c^{cell}(\phi_w(\phi )-\phi_e(\phi )) = \phi^{n}
\end{equation}
The results maintain constancy condition. However, the use of transverse velocity cannot satisfy conservation condition. Both the conservative-form and the advective-form operator splitting  maintain stability, because they both introduce the cross term by using the intermediate $\phi^{n}_{\scriptscriptstyle AY}$ or $\phi^{n}_{\scriptscriptstyle CY}$, though they cannot both satisfy constancy and conservation \citep{Leonard1996}.
\paragraph{}
In order to satisfy both conservation and constancy condition, COSMIC splitting \citep{Leonard1996} combines the conservative-form and advective-form update. Thus, the COSMIC splitting can be written as:
\begin{equation} \label{eq:2.19}
\phi^{n+1} = \phi^n+X_{\scriptscriptstyle C}(\frac{1}{2}(\phi^n + \phi^n _{\scriptscriptstyle AY}) +Y_{\scriptscriptstyle C}(\frac{1}{2}(\phi^n + \phi^n _{\scriptscriptstyle AX})
\end{equation}
According to equation \ref{eq:2.19}, conservation is preserved by the outer conservative-form operator while constancy is maintained by the advective-form update. COSMIC splitting can be interpreted as the mass flux into control-volume cell. \cite{Lauritzen2007} showed that part of the mass comes from the departure cell, which is called local contribution  while there are nonlocal contributions if the Courant number is larger than one and the $\phi_{CX} \neq \phi_{AX}$.

### 2016-01-13 PPM.tex
update the description of results of test cases
Hilary adding more discussion and description to the paper

@@
-\paragraph{}
-The dimensionally splitting scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian by \citep{Lin1996}.
-This dimensionally splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensional.
-\section{Long time-step permitting PPM} 
+The dimensionally split scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian by \citep{Lin1996}.
+This dimension splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensions.
+\subsubsection{Long time-step permitting PPM} 
 \label{sec:PPM}
-\paragraph{} 
+
 Both PPM and the long time steps technique use subgrid cell reconstruction.
@@
 where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid point. Thus, $x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$.
-\paragraph{} 
+
 The long time steps technique \citep{Leonard1995} calculates the flux through the edge of boundaries by transforming the time integral to a space integral in terms of the analytical solution of the advection equation and using the relationship $u\Delta t = c\Delta x$ and $u$dt = dx, where $c$ is the Courant number.
@@
 \end{equation}
-\paragraph{} 
+
 \begin{figure}
@@
 \centering
-\includegraphics[scale=0.6]{YumengsGraphics/nirvana.pdf}
+\includegraphics[scale=0.4]{YumengsGraphics/nirvana.pdf}
 % \vspace{-20pt}
@@
 \end{equation}
-\paragraph{}
+
 The remaining problem is to calculate the component of the flux related to the remnant Courant number, $c_r$, using PPM.
@@
 Therefore, the flux can be calculated directly using the integral of $\phi(x)$
-\paragraph{} 
+
 From equation \ref{eq:2.4}, the boundary values of $\phi$ are required.
@@
 % where $\delta \phi_j = \frac{1}{2}(\phi_{j+1} - \phi_{j-1})$.
-% \paragraph{} 
+
 % Godunov's theorem shows that higher than first order accurate linear schemes cannot maintain monotonicity \citep{Durran2010} and PPM uses quartic interpolation, which is fourth order accurate.
@@
 % where $\eta_j = \max(0,\min(\eta^{(1)}(\widetilde \eta_j-\eta^{(2)}),1)$.
-% \paragraph{}
+
 % $\widetilde \eta_j$ can be calculated by:
@@
 Because these constraints use a method with lower order of accuracy to reprensent the discontinuities, this could reduce the overall order of accuracy of high order polynomial interpolation \citep{Lauritzen2007}.
-% \paragraph{}
+
 % Another problem is that the polynomial interpolation could lead to unphysical estimation of left and right boundaries.
@@
 % \end{eqnarray}
-\paragraph{}
+
 The flux thus can be calculated in terms of the tracer at the boundaries of each interval.
@@
 \end{eqnarray}
-\paragraph{}
+
 For the complete Courant number, where the integer part of Courant number $c_N$ and remnant part $c_r$ is combined, the mass flux should be calculated by the following formulas:
@@
 
-\section{COSMIC Splitting}
+\subsubsection{COSMIC Splitting}
 \label{sec:COSMIC}
-\paragraph{}
+
 COSMIC splitting is an operator splitting scheme, which is used so that PPM can be implemented in two dimensions.
@@
 \centering
-\includegraphics[scale=0.6]{YumengsGraphics/conservative_form.pdf}
-\vspace{-20pt}
+\includegraphics[scale=0.4]{YumengsGraphics/conservative_form.pdf}
+\vspace{-10pt}
 \caption{The position of each flux and the position of $\phi$, the subscript w/e/n/s denote the direction of the flux relative to the position of $\phi$ \label{fig:2.2}}
@@
 The subscript $C$ represents the conservative-form operator.
-\paragraph{}
+
 The conservative-form operator update of $\phi$ can thus be calculated as:
@@
 By this means, the flux at each boundary is unique, maintaining conservation.
-\paragraph{}
+
 However, this cannot satisfy constancy condition. If the initial tracer $\phi $ is homogeneous, the conservative-form $x$-direction update becomes:
@@
 The results destroy constancy condition, because in a 2D non-divergent velocity field, $c_{w}-c_e$ could be unequal.
-\paragraph{}
+
 Advective-form operator splitting is another way to treat the advection scheme.
@@
 If the transverse velocity is in the opposite directions at two faces, the cell centred velocity should set to zero.
-\paragraph{}
+
 The advective-form update is:
@@
 \end{eqnarray*}
-\paragraph{}
+
 If the initial tracer $\phi $ is unity, the advective-form update becomes:
@@
 Both the conservative-form and the advective-form operator splitting  maintain stability, because they both introduce the cross term by using the intermediate $\phi^{n}_{\scriptscriptstyle AY}$ or $\phi^{n}_{\scriptscriptstyle CY}$, though they cannot both satisfy constancy and conservation \citep{Leonard1996}.
-\paragraph{}
+
 In order to satisfy both conservation and constancy condition, COSMIC splitting \citep{Leonard1996} combines the conservative-form and advective-form update.

### 2016-01-17 PPM.tex
Hilary's discussion of the reuslts
update the PPM description

@@
-The dimensionally split scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian by \citep{Lin1996}.
+The dimensionally split scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian \citep{Lin1996}.
 This dimension splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensions.
@@
 \label{sec:PPM}
-
 Both PPM and the long time steps technique use subgrid cell reconstruction.
 Although they are defined for varying grid spacing, we will only describe them for uniform grid interval, $\Delta x$.
-Long time-step permitting PPM is described on a domain with length $L$, which is divided into $n$ grid points.
+The scheme is described on a domain with length $L$, which is divided into $n$ grid points.
 The position of $j$th grid is $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots, n-1$ and the corresponding tracer is $\phi_j$.
@@
 where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid point. Thus, $x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$.
-
-The long time steps technique \citep{Leonard1995} calculates the flux through the edge of boundaries by transforming the time integral to a space integral in terms of the analytical solution of the advection equation and using the relationship $u\Delta t = c\Delta x$ and $u$dt = dx, where $c$ is the Courant number.
-This reveals that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$, and can be written as the following equation:
-\begin{equation} \label{eq:2.2} 
-c\phi_{j-\frac{1}{2}} = \frac{c}{\Delta t} \int^{\Delta t}_{0} \phi (x_j - \frac{\Delta x}{2}, t)dt = \frac{1}{\Delta x} \int^{x_j-\frac {\Delta x}{2}}_{x_j-\frac {\Delta x}{2}-c\Delta x} \phi (x)dx 
-\end{equation}
-
 \begin{figure}
@@
 \end{figure}
-The cumulative mass from the start point to the $j$th left zone edges is defined as $M_{j-\frac{1}{2}}$ and can be estimated by the definition of $\phi_j$ in equation \ref{eq:2.1} and can be discretized as follows \citep{Colella1984}:
+
+The long time step permitting technique is illustrated in figure \ref{fig:2.1}.
+Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$.
+\cite{Leonard1995} shows that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.
+
+It is convinient to divide the Courant number, $c$, into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$ \citep{Leonard1995}.
+In this way, the CFL restriction is removed, because the remnant Courant number is always less than one.
+The flux equation becomes:
+\begin{eqnarray} \label{eq:2.2}
+c\cdot \Delta x \phi_{j-\frac{1}{2}} = M_{c_r,R,j-1} + M_{j-\frac{1}{2}}-M_{j-c_N-\frac{1}{2}} & \text{c}>0 &\nonumber \\
+c\cdot \Delta x \phi_{j-\frac{1}{2}} = -M_{c_r,R,j-1} +M_{j+c_N-\frac{1}{2}}- M_{j-\frac{1}{2}} & \text{c}<0 & \\
+c\cdot \Delta x \phi_{j-\frac{1}{2}} = 0 & \text{otherwise} \nonumber 
+\end{eqnarray}
+where $M_{j-\frac{1}{2}}$ is the cumulative mass from the start point to the $j$th left zone edges defined as \citep{Colella1984}:
 \begin{equation} \label{eq:2.3}
@@
 \end{equation}
-If the Courant number is an integer $c_N$, the flux at left boundary can be calculated directly using the difference between the cumulative mass at $x_{j-\frac{1}{2}}$, $M_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}-c_N}$, $M_{j-c_N-\frac{1}{2}}$.
-However, because the Courant number is not necessarily integer, the equation \ref{eq:2.2} is unable to be applied to all conditions.
-Hence, the Courant number is divided into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$.
-The Courant number thus becomes $c = c_N + c_r$ \citep{Leonard1995}.
-The long time step permitting technique is illustrated in figure \ref{fig:2.1}.
-Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$.
-In this way, the CFL restriction is removed, because the remnant Courant number is always less than one.
-The flux equation becomes:
-\begin{equation} 
-c\phi_{j-\frac{1}{2}} = M_{j-\frac{1}{2}} - M_{j-c_N-\frac{1}{2}} + \frac{1}{\Delta x} \int^{x_j-\frac {\Delta x}{2}-c_N\Delta x}_{x_j-\frac {\Delta x}{2}-c\Delta x} \phi (x)dx 
-\end{equation}
 
-The remaining problem is to calculate the component of the flux related to the remnant Courant number, $c_r$, using PPM.
-PPM assumes that the profile at each interval is a parabola and thus can be represented by a second order polynomial \citep{Colella1984}.
-The equation can be achieved using $\phi(\xi = 0) = \phi_L$, $\phi(\xi = 1) = \phi_R$ and $ \int^{1}_{0} \phi (\xi) d\xi = \phi_j  $, which is obtained by \citet{Colella1984} as:
-\begin{equation} \label{eq:2.4} 
-\phi(x) = \phi_{L,j} + \xi (\Delta \phi_j + \phi_{6,j}(1-\xi)) 
-\end{equation} 
-where $ \xi = \frac{x-x_{j-\frac{1}{2}}}{\Delta x} $, $ \Delta \phi_j = \phi_{R,j} - \phi_{L,j} $, $ \phi_{6,j} = 6(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j})) $ and $\phi_{L,j}$ and $\phi_{R,j}$ represent the left and right boundary of $j$th interval respectively.
-Therefore, the flux can be calculated directly using the integral of $\phi(x)$
+The $M_{c_r,R,j-1}$ in equation \ref{eq:2.2} is calculated by PPM, which assumes that the profile at each interval is a parabola \citep{Colella1984}.
+The parabolic equation can be achieved using $\phi(\xi = 0) = \phi_L$, $\phi(\xi = 1) = \phi_R$ and $ \int^{1}_{0} \phi (\xi) d\xi = \phi_j  $, where $\phi_L$ and $\phi_R$ is obtained by a quartic polynomial mass interpolation.
+The mass flux thus can be calculated in terms of the tracer at the boundaries of each interval.
 
-From equation \ref{eq:2.4}, the boundary values of $\phi$ are required.
-A quartic polynomial mass interpolation \citep{Colella1984} is implemented to calculate $M_{j+\frac{1}{2}}$.
-Thus, the tracer value at the right boundary, $\phi_{j+\frac{1}{2}}$, can be calculated by differentiating $M_{j+\frac{1}{2}} $, the resulting formula is:
-\begin{equation} \label{eq:2.5} 
-\phi_{j+\frac{1}{2}} = \phi_j + \frac{1}{2}(\phi_{j+1}-\phi_{j}+\frac{\delta_m \phi_j - \delta_m \phi_{j+1}}{6}) 
-\end{equation} 
-where $\delta_m \phi_j$ is the slope in the $j$th interval and is calculated by:
-\begin{equation}
-\delta \phi_j = \frac{1}{2}(\phi_{j+1} - \phi_{j-1})
-\end{equation}
-% \begin{equation} \label{eq:2.6}
-% \delta_m \phi_j =
-%   \begin{cases}
-%     \min(|\delta \phi_j|, 2|\phi_j - \phi_{j-1}|,2|\phi_{j+1} - \phi_{j}|)\text{sgn}(\delta \phi_j)    & (\phi_{j+1}-\phi_j)(\phi_j-\phi_{j-1})>0\\
-%     0  & \text{ otherwise}
-%   \end{cases}
-% \end{equation}
-% where $\delta \phi_j = \frac{1}{2}(\phi_{j+1} - \phi_{j-1})$.
-
-% Godunov's theorem shows that higher than first order accurate linear schemes cannot maintain monotonicity \citep{Durran2010} and PPM uses quartic interpolation, which is fourth order accurate.
-Hence, there might be spurious oscillations in the numerical solutions and flux limiters/filters can significantly reduce those oscillation and preserve monotonicity.
-Equation \ref{eq:2.6} is a limiter to guarantee monotonicity and can result in a steeper representation of tracer profile around discontinuity, otherwise $\delta \phi_j$ should be directly used in equation \ref{eq:2.5}.
-Some other constraints are also implemented by \cite{Colella1984} for $\phi_{R,j}$ and $\phi_{L,j}$ to prevent the spurious oscillations.
-For smooth profile, $\phi_{j+\frac{1}{2}} = \phi_{R,j} = \phi_{L,j+1}$.
-For discontinuous profiles, some limiters are applied.
-One of the limiters is:
-% \begin{eqnarray}  \label{eq:2.7}
-% \phi_{L,j} = \phi_{L,j}(1-\eta_j) +\eta_j (\phi_{j-1}+\frac{1}{2} \delta_m \phi_{j-1}) \nonumber \\
-% \phi_{R,j} = \phi_{R,j}(1-\eta_j) +\eta_j (\phi_{j+1}-\frac{1}{2} \delta_m \phi_{j+1})
-% \end{eqnarray}
-% where $\eta_j = \max(0,\min(\eta^{(1)}(\widetilde \eta_j-\eta^{(2)}),1)$.
-
-% $\widetilde \eta_j$ can be calculated by:
-% \begin{eqnarray} \label{eq:2.8}
-% \widetilde \eta_j &= -\frac{\phi_{j+2}-3\phi_{j+1}+3\phi_j-\phi_{j-1}}{6(\phi_{j+1}-\phi_{j-1})}\\
-% &\quad \text{if } -\delta^2 \phi_{j+1} \delta^2 \phi_{j-1}>0, |\phi_{j+1}-\phi_{j-1}|-\varepsilon \min(|\phi_{j+1}|,|\phi_{j-1}|)>0\\
-% &=0 \\&\quad \text{ otherwise,}
-% \end{eqnarray}
-% where $\delta^2 \phi_j = \frac{\phi_{j+1}-2\phi_{j}+\phi_{j-1}}{6\Delta x}$.
-The $\eta^{(1)}=20$,$\eta^{(2)}=0.05$ and $\epsilon=0.05$, which shows the level of discontinuity.
-These values ensure that the small jumps in discontinuity are not applied by such limiters and can maintain high order of accuracy.
-Because these constraints use a method with lower order of accuracy to reprensent the discontinuities, this could reduce the overall order of accuracy of high order polynomial interpolation \citep{Lauritzen2007}.
-
-% Another problem is that the polynomial interpolation could lead to unphysical estimation of left and right boundaries.
-Under this circumstance, other constraints should be applied \citep{Colella1984}.
-Without these constraints, the tracer at left boundaries and right boundaries would be much higher or lower than the real value.
-The resulting $\phi_{L,j}$ and $\phi_{R,j}$ should be:
-% \begin{eqnarray} \label{eq:2.9}
-% &\phi_{R,j}\rightarrow 3\phi_j -2\phi_{L,j} &\text{if }& -\frac{(\phi_{R,j}-\phi_{L,j})^2}{6} >(\phi_{R,j} -\phi_{L,j})(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j})) \nonumber\\
-% &\phi_{L,j}\rightarrow 3\phi_j -2\phi_{R,j} &\text{if }& (\phi_{R,j} -\phi_{L,j})(\phi_j-\frac{1}{2}(\phi_{L,j}+\phi_{R,j}))>\frac{(\phi_{R,j}-\phi_{L,j})^2}{6} \\
-% &\phi_{L,j}\rightarrow \phi_j\text{, } \phi_{R,j}\rightarrow\phi_j &\text{if }& (\phi_{R,j} - \phi_j)(\phi_j-\phi_{L,j}) \leqslant 0 \nonumber
-% \end{eqnarray}
-
-The flux thus can be calculated in terms of the tracer at the boundaries of each interval.
-In terms of different upstream directions, the mass flux at the right boundary of $j$th interval can be written as:
-\begin{eqnarray} \label{eq:2.10}
-M_{c_r,R,j} = c_r \Delta x (\phi_{R,j}-\frac{c_r}{2}(\delta \phi_j-(1-\frac{2}{3} c_r)\phi_{6,j})) & \text{ if $c_r$}\geqslant 0 \nonumber \\
-M_{c_r,R,j} = c_r \Delta x (\phi_{L,j+1}-\frac{c_r}{2}(\delta \phi_{j+1}-(1+\frac{2}{3} c_r)\phi_{6,j+1})) & \text{ if  $c_r$} < 0
-\end{eqnarray}
-
-For the complete Courant number, where the integer part of Courant number $c_N$ and remnant part $c_r$ is combined, the mass flux should be calculated by the following formulas:
-\begin{eqnarray} \label{eq:2.11}
-M_{R,j} = M_{c_r,R,j} + M_{j+\frac{1}{2}}-M_{j-N+\frac{1}{2}} & \text{c}>0 &\nonumber \\
-M_{R,j} = -M_{c_r,R,j} +M_{j+N+\frac{1}{2}}- M_{j+\frac{1}{2}} & \text{c}<0 & \\
-M_{R,j} = 0 & \text{otherwise} \nonumber 
-\end{eqnarray}
 Therefore, for westerly flows, the formula for next time step $\phi^{n+1}_j$ is: 
 \begin{equation} \label{eq:2.12}
-\phi^{n+1}_j = \phi^n_j + \frac{M^n_{in,j} - M^n_{out,j}}{\Delta x} = \phi^n_j + \frac{M^n_{R,j} - M^n_{R,j+1}}{\Delta x}
+\phi^{n+1}_j = \phi^n_j + \frac{M^n_{in,j} - M^n_{out,j}}{\Delta x} = \phi^n_j + c(\phi_{j-\frac{1}{2}} - \phi_{j+\frac{1}{2}})%\frac{M^n_{R,j} - M^n_{R,j+1}}{\Delta x}
 \end{equation}
@@
 \label{sec:COSMIC}
-
 COSMIC splitting is an operator splitting scheme, which is used so that PPM can be implemented in two dimensions.
 Three of the properties expected of multidimensional advection schemes are: stability, conservation and constancy preservation \citep{Lin1996}.
-We will see how COSMIC splitting \citep{Leonard1996} maintains these three properties.
 \begin{figure}
@@
 Operator splitting can use conservative-form or advective-form operators.
-Use of conservative-form operator with operator splitting leads to inherently mass conservation but cannot satisfy the constancy condition.
-The $x$ and $y$ direction conservative-form operator can be defined:
+Use of conservative-form operator with operator splitting leads to inherently mass conservation but cannot satisfy the constancy condition \citep{Leonard1996}.
+The conservative-form operator in $x$ and $y$ direction is:
 \begin{eqnarray}  \label{eq:2.13}
@@
 \end{eqnarray}
-where $c$ is the Courant number and the subscripts represent the direction of the boundary of each grid and w/e/n/s denote the compass direction relative to $\phi$.
-$\phi $ in the brackets means flux at the boundary $c\phi$ is calculated from cell average $\phi$ using the one-dimensional long time permitting PPM.
-As shown in figure \ref{fig:2.2}, this is an Arakawa C staggered grid.
+where $c$ is the Courant number and the subscripts represent the direction of the boundary of each grid and w/e/n/s denote the compass direction relative to $\phi$, as shown in figure \ref{fig:2.2}.
 The subscript $C$ represents the conservative-form operator.
-
-The conservative-form operator update of $\phi$ can thus be calculated as:
-\begin{eqnarray} \label{eq:2.14}
-&\phi^n_{\scriptscriptstyle CY} = \phi^n +Y_{\scriptscriptstyle C}(\phi^n)
-&\phi^{n+1}_{\scriptscriptstyle CYX} = \phi^n_{\scriptscriptstyle CY} +X_{\scriptscriptstyle C}(\phi^n_{\scriptscriptstyle CY})
-\end{eqnarray}
-or
-\begin{eqnarray*} 
-&\phi^n_{\scriptscriptstyle CX} = \phi^n +X_{\scriptscriptstyle C}(\phi^n) 
-&\phi^{n+1}_{\scriptscriptstyle CXY} = \phi^n_{\scriptscriptstyle CX} +Y_{\scriptscriptstyle C}(\phi^n_{\scriptscriptstyle CX})
-\end{eqnarray*}
-The conservative-form update calculates the flux at the boundary of each grid, which means it is necessary to use the velocity at boundaries.
-By this means, the flux at each boundary is unique, maintaining conservation.
-
-However, this cannot satisfy constancy condition. If the initial tracer $\phi $ is homogeneous, the conservative-form $x$-direction update becomes:
-\begin{equation} \label{eq:2.15}
-\phi^{n+1}_{\scriptscriptstyle CX} = \phi^n(1 +c_{w}-c_e)
-\end{equation}
-The results destroy constancy condition, because in a 2D non-divergent velocity field, $c_{w}-c_e$ could be unequal.
 
@@
 Unlike the conservative-form, the advective-form splitting is not inherently conservative but can satisfy constancy condition.
-The advective-form operator can be defined as:
+The advective-form operator \citep{Leonard1996} can be defined as:
 \begin{eqnarray} \label{eq:2.16}
@@
 
-The advective-form update is:
-\begin{eqnarray} \label{eq:2.17}
-&\phi^{n}_{\scriptscriptstyle AY} = \phi^{n} +Y_{\scriptscriptstyle A}(\phi^{n}) 
-&\phi^{n+1}_{\scriptscriptstyle AYX} = \phi^{n}_{\scriptscriptstyle AY} +X_{\scriptscriptstyle A}(\phi^{n}_{\scriptscriptstyle AY})
-\end{eqnarray}
-or
-\begin{eqnarray*}
-&\phi^{n}_{\scriptscriptstyle AX} = \phi^{n} +X_{\scriptscriptstyle A}(\phi^{n}) 
-&\phi^{n+1}_{\scriptscriptstyle AXY} = \phi^{n}_{\scriptscriptstyle AX} +Y_{\scriptscriptstyle A}(\phi^{n}_{\scriptscriptstyle AX})
-\end{eqnarray*}
-
-If the initial tracer $\phi $ is unity, the advective-form update becomes:
-\begin{equation} \label{eq:2.18}
-\phi^{n+1}_{\scriptscriptstyle AX} = \phi^{n}+c^{cell}(\phi_w(\phi )-\phi_e(\phi )) = \phi^{n}
-\end{equation}
-The results maintain constancy condition.
-However, the use of transverse velocity cannot satisfy conservation condition.
-Both the conservative-form and the advective-form operator splitting  maintain stability, because they both introduce the cross term by using the intermediate $\phi^{n}_{\scriptscriptstyle AY}$ or $\phi^{n}_{\scriptscriptstyle CY}$, though they cannot both satisfy constancy and conservation \citep{Leonard1996}.
-
 In order to satisfy both conservation and constancy condition, COSMIC splitting \citep{Leonard1996} combines the conservative-form and advective-form update.
@@
 According to equation \ref{eq:2.19}, conservation is preserved by the outer conservative-form operator while constancy is maintained by the advective-form update.
-COSMIC splitting can be interpreted as the mass flux into control-volume cell.
-\cite{Lauritzen2007} showed that part of the mass comes from the departure cell, which is called local contribution  while there are nonlocal contributions if the Courant number is larger than one and the $\phi_{CX} \neq \phi_{AX}$.
+COSMIC splitting can be interpreted as the mass flux into control-volume cell.% \cite{Lauritzen2007} showed that part of the mass comes from the departure cell, which is called local contribution  while there are nonlocal contributions if the Courant number is larger than one and the $\phi_{CX} \neq \phi_{AX}$.

### 2016-01-26 PPM.tex
Yumeng's change on PPM and COSMIC description and W coordinate

@@
-The dimensionally split scheme decribed in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian \citep{Lin1996}.
+The dimensionally split scheme described in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian \citep{Lin1996}.
 This dimension splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensions.
@@
 Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$.
+The deformational Courant number is $\Delta t \max(|u_x|,|u_y|,|v_x|,|v_x|)$ \citep{PS84} and the sufficient condition for the stability is $c_d \leq 1$.
 \cite{Leonard1995} shows that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.
@@
 Operator splitting can use conservative-form or advective-form operators.
-Use of conservative-form operator with operator splitting leads to inherently mass conservation but cannot satisfy the constancy condition \citep{Leonard1996}.
+Use of conservative-form operator with operator splitting leads to inherently mass conservation but cannot satisfy the constancy condition after several iteration \citep{Leonard1996}.
 The conservative-form operator in $x$ and $y$ direction is:

### 2016-01-26 getGraphics.sh
Hilary's version of the W co-ordindate

@@ -15,7 +15,7 @@ deformingAdvection/orthogonal/1600x800/5/T.pdf
 
 # List of files to copy and rename
 cpFiles=(
-deformingAdvection/nonOrthog/120x60/0/Tmesh.pdf
+deformingAdvection/nonOrthogW/120x60/0/Tmesh.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt1000CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt100CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt200CN_up/10000/Tall.pdf

### 2016-01-29 getGraphics.sh
Inserted more blank data files into the data directory for Yumeng's results

@@ -5,17 +5,17 @@ ROOTDIR=/home/hilary/OpenFOAM/*/run
 
 #List of files to convert to png
 pngFiles=(
-deformingAdvection/orthogonal/1600x800/0/T.pdf
-deformingAdvection/orthogonal/1600x800/1/T.pdf
-deformingAdvection/orthogonal/1600x800/2/T.pdf
-deformingAdvection/orthogonal/1600x800/3/T.pdf
-deformingAdvection/orthogonal/1600x800/4/T.pdf
-deformingAdvection/orthogonal/1600x800/5/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/0/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/1/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/2/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/3/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/4/T.pdf
+deformingAdvection/orthogonal_c1/1600x800/5/T.pdf
 )
 
 # List of files to copy and rename
 cpFiles=(
-deformingAdvection/nonOrthogW/120x60/0/Tmesh.pdf
+deformingAdvection/nonOrthogW_c1/120x60/0/Tmesh.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt1000CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt100CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt200CN_up/10000/Tall.pdf
@@ -52,8 +52,12 @@ solidBodyRotationOnPlane/nonOrthog/50x50/analytic/0/UT.pdf
 
 # List of files to convert from eps format to pdf
 epsFiles=(
-deformingAdvection/plots/l2errors.eps
-deformingAdvection/plots/linferrors.eps
+deformingAdvection/plots/l2errors_c1.eps
+deformingAdvection/plots/linferrors_c1.eps
+deformingAdvection/plots/l2errors_c10.eps
+deformingAdvection/plots/linferrors_c10.eps
+deformingAdvection/plots/l2errors_dt.eps
+deformingAdvection/plots/linferrors_dt.eps
 deformingAdvection/legends/Tmesh_T.eps
 deformingAdvection/legends/T.eps
[…]

### 2016-01-29 getGraphics.sh
Updated table 1

@@ -5,12 +5,12 @@ ROOTDIR=/home/hilary/OpenFOAM/*/run
 
 #List of files to convert to png
 pngFiles=(
-deformingAdvection/orthogonal_c1/1600x800/0/T.pdf
-deformingAdvection/orthogonal_c1/1600x800/1/T.pdf
-deformingAdvection/orthogonal_c1/1600x800/2/T.pdf
-deformingAdvection/orthogonal_c1/1600x800/3/T.pdf
-deformingAdvection/orthogonal_c1/1600x800/4/T.pdf
-deformingAdvection/orthogonal_c1/1600x800/5/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/0/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/1/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/2/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/3/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/4/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/5/T.pdf
 )
 
 # List of files to copy and rename

### 2016-01-29 PPM.tex
Modified definition of deformational Courant number
Added Hilary's alternative description of PPM. Is it correct?

@@
-The dimensionally split scheme described in this paper is known as Conservative Operator Splitting for Multidimensions with Inherently Constancy (COSMIC) splitting \citep{Leonard1996} or flux-form semi-Lagrangian \citep{Lin1996}.
-This dimension splitting scheme allows one-dimensional piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995} to be extended to two-dimensions.
+The one-dimensional scheme that we use is the piecewise parabolic method (PPM) \citep{Colella1984} combined with a long time steps technique \citep{Leonard1995}.
+This is extended to two-dimensions using COSMIC splitting \citep{Leonard1996} (Conservative Operator Splitting for Multidimensions with Inherently Constancy).
+PPM with a long time-step is a one-dimensional flux-form semi-Lagrangian scheme \citep{Lin1996}.
+
 \subsubsection{Long time-step permitting PPM} 
 \label{sec:PPM}
+
 Both PPM and the long time steps technique use subgrid cell reconstruction.
 Although they are defined for varying grid spacing, we will only describe them for uniform grid interval, $\Delta x$.
-The scheme is described on a domain with length $L$, which is divided into $n$ grid points.
-The position of $j$th grid is $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots, n-1$ and the corresponding tracer is $\phi_j$.
-Subgrid cell reconstruction uses the space average mass within the $j$th grid box at time $t_n$, which is defined as:
+We define the position of $j$th grid cell centre to be $x_j = j\Delta x$ and $j = 0, 1, 2, \cdots$ and the corresponding tracer value to be $\phi_j$.
+Subgrid cell reconstruction uses the space average mass within the $j$th grid box which is defined as:
 \begin{equation} \label{eq:2.1} 
-\phi^n_j = \frac{1}{\Delta x} \int^{x_{j+\frac{1}{2}}}_{x_{j-\frac{1}{2}}} \phi (x,t)dx
+\phi_j = \frac{1}{\Delta x} \int^{x_{j+\frac{1}{2}}}_{x_{j-\frac{1}{2}}} \phi (x,t)dx
 \end{equation}
-where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid point. Thus, $x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$.
+where $x_{j-\frac{1}{2}}$ is the midpoint between $x_{j-1}$ and $x_{j}$ and is also the boundary of each grid box ($x_{j-\frac{1}{2}} = x_j - \frac{\Delta x}{2}$).
 \begin{figure}
@@
 The long time step permitting technique is illustrated in figure \ref{fig:2.1}.
-Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point and the time step is restricted by the deformational Courant number $c_d$.
-The deformational Courant number is $\Delta t \max(|u_x|,|u_y|,|v_x|,|v_x|)$ \citep{PS84} and the sufficient condition for the stability is $c_d \leq 1$.
-\cite{Leonard1995} shows that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.
+Similar to semi-Lagrangian, the long time step permitting feature is achieved by tracing back to the departure point in one dimension.
+The time step is restricted by the deformational Courant number $c_d$:
+\begin{equation}
+c_d=
+\Delta t \max
+\left(\biggl|\frac{\partial u}{\partial x}\biggr|,\biggl|\frac{\partial u}{\partial y}\biggr|,\biggl|\frac{\partial u}{\partial z}\biggr|,\biggl|\frac{\partial v}{\partial x}\biggr|,\biggl|\frac{\partial v}{\partial y}\biggr|,\biggl|\frac{\partial v}{\partial z}\biggr|,\biggl|\frac{\partial w}{\partial x}\biggr|,\biggl|\frac{\partial w}{\partial y}\biggr|,\biggl|\frac{\partial w}{\partial z}\biggr|\right)
+\label{eq:cd}
+\end{equation}
+\citep{PS84} and the stability criterion is $c_d \leq 1$.
 
-It is convinient to divide the Courant number, $c$, into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$ \citep{Leonard1995}.
+\cite{Leonard1995} showed that the flux at the left boundary $x_{j-\frac{1}{2}}$ can be obtained using the average mass between $x_{j-\frac{1}{2}}$ and $x_{j-\frac{1}{2}}-c\Delta x$.
+
+It is convenient to divide the Courant number, $c$, into integer part, $c_N$, and decimal part (or remnant Courant number), $c_r$ \citep{Leonard1995}.
 In this way, the CFL restriction is removed, because the remnant Courant number is always less than one.
@@
 \begin{eqnarray} \label{eq:2.2}
-c\cdot \Delta x \phi_{j-\frac{1}{2}} = M_{c_r,R,j-1} + M_{j-\frac{1}{2}}-M_{j-c_N-\frac{1}{2}} & \text{c}>0 &\nonumber \\
-c\cdot \Delta x \phi_{j-\frac{1}{2}} = -M_{c_r,R,j-1} +M_{j+c_N-\frac{1}{2}}- M_{j-\frac{1}{2}} & \text{c}<0 & \\
-c\cdot \Delta x \phi_{j-\frac{1}{2}} = 0 & \text{otherwise} \nonumber 
+c \Delta x \phi_{j-\frac{1}{2}} = M_{c_r,R,j-1} + M_{j-\frac{1}{2}}-M_{j-c_N-\frac{1}{2}} & \text{c}>0 &\nonumber \\
+c \Delta x \phi_{j-\frac{1}{2}} = -M_{c_r,R,j-1} +M_{j+c_N-\frac{1}{2}}- M_{j-\frac{1}{2}} & \text{c}<0 & \\
+c \Delta x \phi_{j-\frac{1}{2}} = 0 & \text{otherwise} \nonumber 
 \end{eqnarray}

### 2016-01-29 PPM_Hilary.tex
Added Hilary's alternative description of PPM. Is it correct?

The one-dimensional scheme that we use is the piecewise parabolic method (PPM) \citep{Colella1984} without any monotonicity constraints combined with a long time steps technique \citep{Leonard1995}. This is extended to two-dimensions using COSMIC splitting \citep{Leonard1996} (Conservative Operator Splitting for Multidimensions with Inherently Constancy). PPM with a long time-step is a one-dimensional flux-form semi-Lagrangian scheme \citep{Lin1996}.

\subsubsection{Long time-step permitting PPM} 
\label{sec:PPM}

Both PPM and the long time steps technique use subgrid cell reconstruction. Although
\cite{Colella1984} defined PPM for varying grid spacing, we will define PPM and the long times-step technique for a uniform grid interval, $\Delta x$. The dependent variable, $\phi_j$, is the mean value of $\phi$ for each cell with centre at $x_j = j\Delta x$ for $j = 0, 1, 2, \cdots$. Consequently the mass in each cell is:
\[
M_j = \phi_j \Delta x.
\]
The PPM method entails finding a polynomial, $\phi(x)$, for each cell such that:
\[
\phi_j = \frac{1}{\Delta x} \int_{j-1/2}^{j+1/2} \phi(x) dx
\]
and then integrating this polynomial along the distance travlled in each time-step upwind of each cell boundary, $x_{j\pm 1/2}$. The polynomial is defined as:
\[
\phi(x) = \phi_{j-1/2}
        + \xi \biggl(\phi_{j+1/2} - \phi_{j-1/2}
                + (1-\xi)6\bigl(\phi_j - \frac{1}{2}(\phi_{j-1/2} + \phi_{j+1/2})\bigr)
              \biggr)
\]
where $\xi=(x-x_{j-1/2})/\Delta x$ and the value of $\phi$ at the cell boundaries are given by:
\[
\phi_{j+1/2} = \phi_j + \frac{1}{2}(\phi_{j+1} - \phi_j)
\biggl(
     1-\frac{1}{12}(\phi_{j+1} - \phi_{j-1})
\biggr)
\]

In order to cope with long time-steps, we follow \cite{Leonard1995} and divide the Courant number into an integer part, $c_N$, and a remainder, $c_r$. Then the flux through boundary $j-1/2$ between times $n\Delta t$ and $(n+1)\Delta t$ is given by:
\[
c \phi_{j-1/2}^{(n+1/2)} = M_{j-1/2} - M_{j-c_N-1/2} + \frac{1}{\Delta x}
\int_{x_j-\Delta x/2 - c_r \Delta x}^{x_j - \Delta x/2}
\phi(x) dx
\]
where $M_{j-1/2}$ is the cumulative mass from teh start point to position $x_{j-1/2}$:
\[
M_{j-1/2} - \sum_{k<j} \Delta x \phi_k.
\]

### 2016-01-29 getGraphics.sh

@@ -5,12 +5,12 @@ ROOTDIR=/home/hilary/OpenFOAM/*/run
 
 #List of files to convert to png
 pngFiles=(
-deformingAdvection/orthogonal_c1/1920x960/0/T.pdf
-deformingAdvection/orthogonal_c1/1920x960/1/T.pdf
-deformingAdvection/orthogonal_c1/1920x960/2/T.pdf
-deformingAdvection/orthogonal_c1/1920x960/3/T.pdf
-deformingAdvection/orthogonal_c1/1920x960/4/T.pdf
-deformingAdvection/orthogonal_c1/1920x960/5/T.pdf
+deformingAdvection/orthogonal_c1/960x480/0/T.pdf
+deformingAdvection/orthogonal_c1/960x480/1/T.pdf
+deformingAdvection/orthogonal_c1/960x480/2/T.pdf
+deformingAdvection/orthogonal_c1/960x480/3/T.pdf
+deformingAdvection/orthogonal_c1/960x480/4/T.pdf
+deformingAdvection/orthogonal_c1/960x480/5/T.pdf
 )
 
 # List of files to copy and rename

### 2016-01-31 getGraphics.sh
new graphics for deformational flow

@@ -5,12 +5,12 @@ ROOTDIR=/home/hilary/OpenFOAM/*/run
 
 #List of files to convert to png
 pngFiles=(
-deformingAdvection/orthogonal_c1/960x480/0/T.pdf
-deformingAdvection/orthogonal_c1/960x480/1/T.pdf
-deformingAdvection/orthogonal_c1/960x480/2/T.pdf
-deformingAdvection/orthogonal_c1/960x480/3/T.pdf
-deformingAdvection/orthogonal_c1/960x480/4/T.pdf
-deformingAdvection/orthogonal_c1/960x480/5/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/0/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/1/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/2/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/3/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/4/T.pdf
+deformingAdvection/orthogonal_c1/1920x960/5/T.pdf
 )
 
 # List of files to copy and rename

### 2017-01-19 advectionPaper.tex
Preparing manuscript for submission to qjrms

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,a4paper,english,british]{qjrms4}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{luainputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{color}
\usepackage{array}
\usepackage{verbatim}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}

\usepackage{todonotes}

\usepackage{amssymb}

\makeatother

\usepackage{babel}
\begin{document}

\title{Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme
for Atmospheric Transport}

\author{Yumeng Chen, Hilary Weller, Stephen Pring and James Shaw}

%\maketitle
\begin{comment}
\tableofcontents{}

\listoftodos
\end{comment}

\begin{abstract}
Dimensionally split advection schemes are attractive for atmospheric
modelling due to their efficiency and accuracy in each spatial dimension.
Accurate long time steps can be achieved without significant cost
using the flux-form semi-Lagrangian technique. The dimensionally split
scheme used in this paper is constructed from the one-dimensional
Piecewise Parabolic Method and extended to two dimensions using COSMIC
splitting. The dimensionally split scheme is compared with a genuinely
multi-dimensional, method of lines scheme with implicit time-stepping
which is stable for very large Courant numbers.

Two-dimensional advection test cases on Cartesian planes are proposed
that avoid the complexities of a spherical domain or multi-panel meshes.
These are solid body rotation, horizontal advection over orography
and deformational flow. The test cases use distorted meshes either
to represent sloping terrain or to mimic the distortions of a cubed
sphere.

Such mesh distortions are expected to accentuate the errors associated
with dimension splitting, however, the dimensionally split scheme
is very accurate on orthogonal meshes and accuracy decreases only
a little in the presence of large mesh distortions. The dimensionally
split scheme retains accuracy when long time-steps are used. The multi-dimensional
scheme is almost entirely insensitive to mesh distortions and asymptotes
to second-order accuracy at high resolution. As is expected for implicit
time-stepping, phase errors occur when using long time-steps but the
spatially well resolved features are advected at the correct speed
and the multi-dimensional scheme is always stable.

A naive estimate of computational cost (number of multiplies) reveals
that the implicit scheme is  the most expensive, particularly for
large Courant numbers. If the multi-dimensional scheme is used instead
with explicit time-stepping, the Courant number is restricted to less
than one but the cost becomes similar to the dimensionally split scheme.
\end{abstract}
\begin{keywords}
hello
\end{keywords}

\section{Introduction}

Weather and climate models are being developed on quasi-uniform meshes
in order to better exploit modern computers \citep[eg ][]{WTC12,LUJ+14,ST12,SG11,KNK15}
and so accurate and efficient transport (or advection) schemes on
non-orthogonal meshes are required. There is an abundance of desirable
properties of advection schemes, including:
\begin{enumerate}
\item Inherent local conservation of the advected quantity.
\item Stability in the presence of large Courant numbers.
\item Accuracy in the presence of large Courant numbers.
\item High order-accuracy.
\item Low computational cost, good parallel scaling and multi-tracer efficiency.
\item Low phase and dispersion errors (advection of all wavenumbers of the
advected quantity at close to the correct speed).
\item Low diffusion errors (maintaining amplitude of all wavenumbers of
the advected quantity).
\item Boundedness, monotonicity, positivity and maintaining correlations
between multiple advected tracers.
\end{enumerate}
Four (important) properties are listed together in the final item
because they will not be addressed here. In this paper, we address
the issue of conservative, accurate, efficient advection schemes for
logically rectangular, non-orthogonal meshes which are stable in the
presence of large Courant numbers. These schemes would be particularly
relevant for cubed-sphere meshes and for terrain following meshes.
Another novel aspect of this paper is that, for simplicity, we test
advection schemes entirely on planar meshes rather than on the sphere,
proposing test cases to challenge advection schemes on non-orthogonal
meshes without the need to implement meshes in spherical geometry.

Dimensionally split schemes (operating separately in each spatial
dimension) are attractive for atmospheric modelling due to their efficiency
and high accuracy in each spatial dimension \citep[eg. ][]{LR96,LLM96,BS99,PL07,KNK15}.
Inherent conservation is guaranteed by using the flux-form semi-Lagrangian
(FFSL or forward in time) technique \citep[eg. ][]{CW84} which integrates
the dependent variable over a swept distance upstream of every face
in order to calculate fluxes in and out of cells. Accurate long time
steps can be achieved without significant cost by calculating cumulative
mass fluxes along the domain \foreignlanguage{english}{\citep{LLM95}.
}This is done without remapping which can be an expensive procedure,
finding a conservative map between fields on overlapping meshes\foreignlanguage{english}{.
}One-dimensional schemes can be used with operator splitting to create
dimensionally split, second-order accurate schemes \citep[eg][]{LLM96}
on logically rectangular, multidimensional meshes. Dimensionally split
schemes have been found to give good accuracy on non-orthogonal meshes
such as the cubed-sphere \citep{PL07,KNK15} with special treatment
over cube edges. \citet{PL07} use the average of two one-sided schemes
at cube edges whereas \citet{KNK15} create ghost cells outside each
cube panel boundary. Without this special treatment, dimensionally
split schemes do not account properly for mesh distortion and errors
may occur. Accuracy between second and fourth order was found in practice
on a range of test cases on the cubed-sphere by \citet{KNK15}.

The same problems occur when using dimensionally split schemes over
orography since terrain following layers become non-orthogonal over
orography and special treatment cannot be used everywhere where there
is orography. A common solution has been to make the terrain following
layers as smooth as possible, reducing non-orthogonality \citep[eg][]{SLF+02}
or to use floating Lagrangian vertical co-ordinates \citep{Lin04}.
Dimension splitting may account for some of the errors over orography
reported by \citet{KUJ14} although they do not cite this as a reason
for errors.

Dimension splitting errors on distorted meshes can be eliminated by
using genuinely multi-dimensional advection schemes. These can be
either FFSL \citep[ie swept area, eg.][]{Las02,LR05b,Miu07b,TCD14},
method of lines \citep[MOL, discretising space and time separetely, eg.][]{WWF09,SG11,KNK15,SWMD1x}
or conservative semi-Lagrangian \citep[with conservative re-mapping, eg][]{IK04,LNU10,ZWS04}.
The FFSL and MOL multi-dimensional schemes have not previously been
extended to work with Courant numbers significantly larger than one.
FFSL multi-dimensional schemes could be extended to handle large time-steps
by integrating the upstream swept volume over a large upstream volume,
interacting with a large number of upstream cells. However the cost
would be proportional to the time-step since the larger the Courant
number, the more cells the upstream swept volume would need to overlap
with. This technique therefore offers no advantage over using a smaller
time-step. MOL multi-dimensional schemes can be extended to work with
Courant numbers larger than one by using implicit time-stepping. This
will increase the computational cost per tracer advected since the
solution of a matrix equation would be needed for every advected tracer.
Two other disadvantages of implicit time-stepping are the large phase
errors when long time-steps are used \citep[eg][]{DB12,LWW14} and
the difficulty of achieving monotonicity. This is in contrast to semi-Lagrangian
or FFSL schemes which maintain accuracy with long time-steps \citep{Pur76,PS84,LLM95}
although monotonocity with long time-steps is still challenging \citep{Bott10}.
Conservative semi-Lagrangian naturally extends to long time-steps
but the conservative remapping is complicated and expensive, particularly
on non-rectangular meshes and will not be investigated here. \citet{LUJ+14}
described how the FFSL technique with a long time-step can be made
equivalent to the conservative semi-Lagrangian.

It is therefore not clear what approach should be taken for achieving
long time-steps when advecting multiple tracers on distorted meshes.
In this paper, we show the effect of dimension splitting errors using
a FFSL dimensionally split scheme on a number of test-cases which
use distorted meshes and compare with a genuinely multi-dimensional
implicit MOL scheme using large and small Courant numbers.

The theoretical properties of dimensionally split advection schemes
are often tested on uniform, orthogonal meshes \citep[eg][]{LLM96}.
On the cubed-sphere, special treatment is needed for the cube edges
\citep[eg][]{LR96,KNK15}. Developing a transport scheme to the extent
that it can be used on a multi-panel cubed-sphere with special treatment
of cube edges is a considerable undertaking. Hence there is a need
for more challenging advection test cases which are simpler to implement,
without the need for spherical meshes. We therefore propose some modifications
of existing test cases to use distorted meshes, or distorted co-ordinate
systems, on a logically rectangular, two-dimensional plane.

The long time-step permitting, dimensionally split scheme and the
long time-step permitting multi-dimensional scheme are defined in
section \ref{sec:models}. In section \ref{sec:results} we present
results of three advection test cases on distorted meshes in two-dimensional
planes using Courant numbers above and below one. These are the solid
body rotation test case of \citet{LLM96} modified to use a mesh (or
co-ordinate system) with distortions similar to a cubed-sphere (section
\ref{sub:solidRotate}), the horizontal advection test case over orography
\citep{SLF+02}, examining sensitivity to time-step, resolution and
mountain height, all on the maximally distorted basic terrain following
mesh (section \ref{sub:overOrog}) and a modification of the deformational
flow test case of \citet{LSPT12} for a periodic rectangular plane
(section \ref{sub:deformFlow}). Some estimates are made of computational
cost in section \ref{sub:cost} and final conclusions are drawn in
section \ref{sec:concs}.

\section{Transport Schemes\label{sec:models}}

We present two conservative advection schemes suitable for long time-steps
(stable for Courant numbers significantly larger than one) for solving
the linear advection equation:
\begin{equation}
\frac{\partial\phi}{\partial t}+\nabla\cdot\mathbf{u}\phi=0
\end{equation}
where the dependent variable $\phi$ is advected by velocity field
$\mathbf{u}(\mathbf{x},t)$. The dimensionally split scheme is the\foreignlanguage{english}{
piecewise parabolic method \citep[PPM][]{CW84}} which uses the flux-form
semi-Lagrangian approach extended to long time-steps following \citet{LLM95}
with COSMIC splitting \citep{LLM96} to extend PPM to two dimensions.
The multi-dimensional scheme uses the method of lines approach (treating
space and time independently). The second-order accurate spatial discretistaion
of \citet{WS14} is combined with implicit, Crank-Nicholson time-stepping
to allow long time-steps. Neither scheme has monotonicity or positivity
preservation.

The code for dimensionally splitting scheme is available at \url{https://github.com/yumengch/COSMIC-splitting}.
The multi-dimensional scheme is implemented using \citet{OpenFOAM16}
and is available at \url{https://github.com/AtmosFOAM/}.

\subsection{One-dimensional PPM with Long Time-steps\label{sub:PPM}}

We describe a long time-step version of PPM\foreignlanguage{english}{
for solving the one-dimensional advection equation:
\begin{equation}
\frac{\partial\phi}{\partial t}+\frac{\partial u\phi}{\partial x}=0.
\end{equation}
\citet{CW84} defined PPM with monotonicity constraints and for variable
resolution but for simplicity (and for comparison with the multi-dimensional
scheme) we will define PPM without monotonicity constraints and for
a fixed resolution, $\Delta x$, and time-step, $\Delta t$. With
these restrictions, PPM should be fourth-order accurate in one dimension
for vanishing time-step. We define the dependent variable, $\phi_{i}^{(n)}$,
to be the mean value of $\phi$ in cell $i$ at time-level $n$ where
$x_{i}=i\Delta x$ and $t=n\Delta t$. Since PPM is a flux-form finite
volume method, $\phi_{i}^{(n)}$ is updated using:
\begin{equation}
\phi_{i}^{(n+1)}=\phi_{i}^{(n)}+X_{C}\left(\phi\right)=\phi_{i}^{(n)}-\frac{u_{i+1/2}\phi_{i+1/2}-u_{i-1/2}\phi_{i-1/2}}{\Delta x}\label{PPMgauss}
\end{equation}
where $X_{C}\left(\phi\right)$ is the conservative advection operator
for $\phi$ in the $x$ direction. The fluxes, $u_{i\pm1/2}\phi_{i\pm1/2}$,
are found by integrating a piecewise polynomial, $p$, along the distance
travelled in each time-step upwind of cell boundary $x_{i\pm1/2}$.
The polynomial is defined in each cell, $i$, such that:
\begin{equation}
\phi_{i}=\frac{1}{\Delta x}\int_{x_{i-1/2}}^{x_{i+1/2}}p_{i}(x)\ dx
\end{equation}
by
\begin{equation}
p_{i}(x)=p_{i-1/2}+\xi\left(p_{i+1/2}-p_{i-1/2}+\left(1-\xi\right)6\left(\phi_{i}-\frac{1}{2}\left(p_{i-1/2}+p_{i+1/2}\right)\right)\right)
\end{equation}
where $\xi=(x-x_{i-1/2})/\Delta x$ and 
\begin{equation}
p_{i+1/2}=\frac{7}{12}(\phi_{i}+\phi_{i+1})-\frac{1}{12}(\phi_{i+2}+\phi_{i-1})
\end{equation}
}

%
In order to cope with long time-steps, we follow \citet{LLM95} and
divide the Courant number into a signed integer part, $c_{N}$, and
a remainder, $c_{r}$. The departure point of location $x_{i-\frac{1}{2}}$
is thus computed as $x_{d}=x_{i-\frac{1}{2}}-u_{i-\frac{1}{2}}\Delta t$
and the departure cell, for cell edge $i-\frac{1}{2}$, is $i_{d}=i-c_{N}-1$
for $u_{i-\frac{1}{2}}>0$ and $i_{d}=i-c_{N}$ for $u_{i-\frac{1}{2}}<0$.
Then for $u_{i-\frac{1}{2}}>0$ the flux through $x_{i-1/2}$ between
times $n\Delta t$ and $(n+1)\Delta t$ is: 
\begin{equation}
u_{i-\frac{1}{2}}\phi_{i-\frac{1}{2}}=\frac{1}{\Delta t}\int_{x_{d}}^{x_{i-1/2}}p(x)\ dx=\frac{1}{\Delta t}\left(M_{i-1/2}-M_{i-c_{N}-1/2}+\int_{x_{i-1/2-c_{N}}-c_{r}\Delta x}^{x_{i-1/2-c_{N}}}p(x)\ dx\right)\label{eq:PPMflux}
\end{equation}
where $M_{i-1/2}$ is the cumulative mass from the start point to
position $x_{i-1/2}$: 
\begin{equation}
M_{i-1/2}=\sum_{k<i}\Delta x\phi_{k}.\label{eq:cumulativeMass}
\end{equation}
This departure point calculation assumes that the velocity is uniform
on the computational mesh which has a first-order error which could
be particularly damaging for long Courant numbers, when the wrong
departure cell could be found.

The velocity is derived from a stream function and the Jacobian of
the co-ordinate transform: \foreignlanguage{british}{
\begin{equation}
\left(\begin{array}{c}
u\\
v
\end{array}\right)=J\left(\begin{array}{c}
\;\Psi_{y}\\
-\Psi_{x}
\end{array}\right)
\end{equation}
}For stability, the time-step is restricted by the deformational Courant
number:
\begin{equation}
c_{d}=\Delta t\max\left(\biggl|\frac{\partial u}{\partial x}\biggr|,\biggl|\frac{\partial u}{\partial y}\biggr|,\biggl|\frac{\partial u}{\partial z}\biggr|,\biggl|\frac{\partial v}{\partial x}\biggr|,\biggl|\frac{\partial v}{\partial y}\biggr|,\biggl|\frac{\partial v}{\partial z}\biggr|,\biggl|\frac{\partial w}{\partial x}\biggr|,\biggl|\frac{\partial w}{\partial y}\biggr|,\biggl|\frac{\partial w}{\partial z}\biggr|\right)\label{eq:cd}
\end{equation}
\citep{PS84} such that $c_{d}\leq1$.

\subsection{COSMIC Splitting\label{sub:COSMIC}}

COSMIC operator splitting \citep{LLM96} allows single stage, one-dimensional
schemes such as PPM to be used stably in two or more dimensions whilst
retaining conservation, constancy preservation and second-order accuracy
(on orthogonal meshes). As we are now considering two spatial dimensions,
we define $\phi_{ij}$, $u_{ij}$ and $v_{ij}$, the values of $\phi$
and the velocity components, $u$ and $v$ in cell $(i,j)$ where
$x=i\Delta x$ and $y=j\Delta y$. COSMIC splitting uses both advective
and conservative advection operators in the $x$ and $y$ directions:
\begin{align}
X_{C}(\phi)=-\frac{1}{\Delta x}\left(u_{e}\phi_{e}-u_{w}\phi_{w}\right) &  & Y_{C}(\phi)=-\frac{1}{\Delta y}\left(v_{n}\phi_{n}-v_{s}\phi_{s}\right)\label{eq:COSMIC_A}\\
X_{A}(\phi)=X_{c}(\phi)+\frac{\phi_{ij}}{\Delta x}\left(u_{e}-u_{w}\right) &  & Y_{A}(\phi)=Y_{c}(\phi)+\frac{\phi_{ij}}{\Delta y}\left(v_{n}-v_{s}\right)\label{eq:COSMIC_C}
\end{align}
where $\phi_{n}=\phi_{i,j+1/2}$, $\phi_{s}=\phi_{i,j-1/2}$, $\phi_{e}=\phi_{i+1/2,j}$
, $\phi_{w}=\phi_{i-1/2,j}$, $v_{n}=v_{i,j+1/2}$, $v_{s}=v_{i,j-1/2}$,
$u_{e}=u_{i+1/2,j}$ and $u_{w}=u_{i-1/2,j}$ are the values of $\phi$,
$u$ and $v$ at the cell boundaries. If COSMIC is being used to extend
PPM to two spatial dimensions then $\phi_{n,s,e,w}$ are calculated
from equation \ref{eq:PPMflux}. Assuming C-grid staggering, $v_{n}$,
$v_{s}$, $u_{e}$ and $u_{w}$ are dependent variables. Instead of
using cell centered velocity \citep{LR96} or upwind velocity \citep{LLM96},
the advective operators are calculated in a similar manner to \citet{Lin04}.

%
Mesh distortions can be included in the advection equation with a
co-ordinate transform with Jacobian $J$:\foreignlanguage{english}{
\begin{equation}
\frac{\partial|J|^{-1}\phi}{\partial t}+\frac{\partial|J|^{-1}u\phi}{\partial x}+\frac{\partial|J|^{-1}v\phi}{\partial y}=0.
\end{equation}
}The operators $X_{c}$, $Y_{c}$ $X_{A}$ and $Y_{A}$ are then:\foreignlanguage{english}{
\begin{align}
X_{C}(\phi)=-\frac{1}{\Delta x}\left(|J|_{e}^{-1}u_{e}\phi_{e}-|J|_{w}^{-1}u_{w}\phi_{w}\right) &  & Y_{C}(\phi)=-\frac{1}{\Delta y}\left(|J|_{n}^{-1}v_{n}\phi_{n}-|J|_{s}^{-1}v_{s}\phi_{s}\right)\label{eq:COSMIC_A-1}\\
X_{A}(\phi)=X_{c}(\phi)+\frac{\phi_{ij}}{\Delta x}\left(|J|_{e}^{-1}u_{e}-|J|_{w}^{-1}u_{w}\right) &  & Y_{A}(\phi)=Y_{c}(\phi)+\frac{\phi_{ij}}{\Delta y}\left(|J|_{n}^{-1}v_{n}-|J|_{s}^{-1}v_{s}\right)\label{eq:COSMIC_C-1}
\end{align}
which are combined to update $\phi_{ij}$ in each cell by: 
\begin{equation}
\phi_{ij}^{(n+1)}=\phi_{ij}^{n}+|J|_{ij}X_{C}\left(\phi_{ij}^{(n)}+\frac{|J|_{ij}}{2}Y_{A}\left(\phi_{ij}^{(n)}\right)\right)+|J|_{ij}Y_{C}\left(\phi_{ij}^{(n)}+\frac{|J|_{ij}}{2}X_{A}\left(\phi_{ij}^{(n)}\right)\right).\label{eq:COSMICupdate}
\end{equation}
where $|J|_{ij}$ is the determinant of Jacobian at the center of
cell $(i,j)$, and $\Delta x$ and $\Delta y$ give the cell size
in uniform computational domain.}

\subsection{Multi-dimensional Method of Lines (MOL) Scheme\label{sub:MMmolscheme}}

The MOL scheme uses the finite volume method for arbitrary meshes
and is implemented in \citep{OpenFOAM16}. This uses a cubic upwind
spatial discretisation \citep{WS14,SWMD1x} combined with Crank-Nicholson
in time. Although the interpolation uses a cubic polynomial, cell
centre values are approximated as cell average values and face centre
values are approximated as face average values so the method is limited
to 2nd order accuracy in space. The advection scheme uses Gauss's
divergence theorem to approximate the divergence term of the advection
equation:
\begin{equation}
\nabla\cdot\mathbf{u}\phi\approx\frac{1}{V}\sum_{f\in c}\phi_{f}\mathbf{u}_{f}\cdot\mathbf{S}_{f}\label{eq:multidGauss}
\end{equation}
where $V$ is the cell volume, the summation is over all faces, $f$,
of cell $c$, $\phi_{f}$ is the value of the dependent variable,
$\phi$ interpolated onto face $f$, $\mathbf{u}_{f}$ is the velocity
at face $f$ and $\mathbf{S}_{f}$ is the vector normal to face $f$
with magnitude equal to the area of face $f$ (ie the face area vector).
The advection scheme uses a fit to a 2d (or 3d) polynomial using an
upwind-biased stencil of cells (fig \ref{fig:stencil}) in order to
interpolate from known cell values onto a face. In 2d, the cubic polynomial
is:\foreignlanguage{english}{
\begin{equation}
\phi=a+bx+cy+dx^{2}+exy+fy^{2}+gx^{3}+hx^{2}y+ixy^{2}
\end{equation}
omitting the $y^{3}$ term, where $x$ is the direction normal to
a cell face and $y$ is perpendicular to $x$. (The $y^{3}$ term
is omitted because it cannot be set with a stencil that is narrow
in the direction of the flow.) Coefficients $a$ to $i$ are set from
a least squares fit to the cell data in the stencil. The least-squares
problem involves a $9\times m$ matrix singular value decomposition
(where $m$ is the size of the stencil) for every face and for both
orientations of each face . However this is purely a geometric calculation
and is therefore a pre-processing activity since the mesh is fixed.
This generates a set of weights for calculating $\phi_{f}$ from the
cell values in the stencil, leaving $m$ multiplies for each face
for each call of the advection operator. The stencils are found for
three-dimensional, arbitrarily structured meshes by finding the face(s)
closest to upwind of the face we are interpolating onto, taking the
two cells either side of the upwind face(s) and then taking the vertex
neighbours of those central cells (fig \ref{fig:stencil}). For each
face there are two possible stencils depending on the upwind direction.
Both of the stencils are stored and the interpolation weights for
both stencils are calculated. }

\begin{figure}
\noindent \begin{centering}
\scalebox{0.5}[0.5]{\input{figs/cubicUpCPCstencil.pstex_t}}\hfill{}\scalebox{0.5}[0.5]{\input{figs/hexStencil.pstex_t}}
\par\end{centering}

\caption{Stencils of upwind-biased cells for interpolating onto face $f$ using
a 2d cubic polynomial for two different mesh structures.\label{fig:stencil}}
\end{figure}

In order to ensure that the fit is accurate in the cells either side
of face $f$ and to ensure that the values in these adjacent cells
have the strongest control over $\phi_{f}$, rows associated with
these values in the least squares fit matrix are weighted a factor
of 1000 relative to the other rows \citep[following][]{Las02}. This
does not affect the order of accuracy. Mathematically, an arbitrarily
large value of the weight can be used to ensure that the fit goes
exactly through the upwind and downwind cell. However if a value too
large is used, the singular value problem becomes ill conditioned.
We do not use any of the other stabilisation procedures as described
by \citet{SWMD1x}. The value $\phi_{f}$ is then calculated as a
higher-order correction to first-order upwind: 
\begin{equation}
\phi_{f}=\phi_{\text{up}}+\sum_{c}w_{c}\phi_{c}\label{eq:multiDflux}
\end{equation}
where $w_{c}$ are the weights for each cell of the stencil calculated
from the least squares fit, with $w_{\text{up}}$ reduced by 1 to
make the fit a correction on upwind.

\subsubsection{Implicit Solution, Matrix Solvers and Tolerances \label{sub:implicitMethod}}

The trapezoidal implicit or Crank-Nicholson time-stepping leads to
a matrix equation which needs to be solved to find all the $\phi$s
at the next time-step. In order to ensure that the matrix is diagonally
dominant for arbitrary time-steps, the cubic interpolation applied
is a deferred correction on first-order upwind so that only the coefficient
corresponding to the upwind cells are included in the matrix. This
means that more than one implicit solves are needed per time-step
so that the higher-order terms are solved to be second-order accurate
in time. If the Courant number is less than or close to one, we use
two implicit solves per time-step. Consequently, assuming that the
velocity field and mesh are constant in time, the time-stepping scheme
is defined as:
\begin{eqnarray}
\frac{\phi^{\prime}-\phi^{n}}{\Delta t} & = & -\frac{1}{2V}\sum_{f\in c}\left(\phi_{\text{up}}^{n}+\phi_{\text{up}}^{\prime}+2\sum_{c}w_{c}\phi_{c}^{n}\right)\mathbf{u}_{f}\cdot\mathbf{S}_{f}\label{eq:CN_1}\\
\frac{\phi^{n+1}-\phi^{n}}{\Delta t} & = & -\frac{1}{2V}\sum_{f\in c}\left(\phi_{\text{up}}^{n}+\phi_{\text{up}}^{n+1}+\sum_{c}w_{c}\phi_{c}^{n}+\sum_{c}w_{c}\phi_{c}^{\prime}\right)\mathbf{u}_{f}\cdot\mathbf{S}_{f}.\label{eq:CN_2}
\end{eqnarray}
For larger Courant numbers we use four implicit solves per time-step
although sensitivity to this choice has not been investigated.

If this implicit scheme is applied on a logically rectangular, two-dimensional
mesh with horizontal and vertical Courant numbers $c_{x}$ and $c_{z}$,
then the diagonal coefficients of the matrix would be $1+c_{x}/2+c_{z}/2$
and, assuming two upwind directions, there would be exactly two off-diagonal
elements, $-c_{x}/2$ and $-c_{z}/2$. Consequently the matrix is
very sparse, asymmetric and diagonally dominant for all time-steps.
It is solved using the OpenFOAM bi-conjugate gradient solver using
DILU pre-conditioning to a tolerance of $10^{-8}$ every iteration.
Sensitivity to the solver or solver tolerance have not been investigated.
Information about the number of solver iterations for different test
cases in given in section \ref{sub:cost}.

\section{Results of Test Cases in Planar Geometry\label{sec:results}}

In order to make test cases as simple as possible without the need
for incorporating spherical geometry, multi-panel meshes or non-rectangular
cells, our computational domain consists of a periodic two-dimensional
plane with deformations in the co-ordinate system (or mesh) mimicking
the kind of distortions which are produced by a cubed-sphere mesh.
We also use a two-dimensional vertical slice test case over orography
using terrain following co-ordinates (or terrain following meshes).

A range of test cases are undertaken on uniform and distorted meshes
using the dimensionally split scheme and the multi-dimensional scheme
in order to evaluate the influences of mesh distortions, the validity
of using a dimensionally split scheme on a distorted mesh and the
schemes' accuracy and stability for long time-steps.

\subsection{Solid Body Rotation\label{sub:solidRotate}}

The solid body rotation test case of \citet{LLM96} is used to compare
the accuracy of the dimensionally split and multi-dimensional schemes
on orthogonal and non-orthogonal meshes. We define this test case
on a domain that is $10^{4}\times10^{4}\text{ m}^{2}$. The velocity
is defined by numerically differentiating the streamfunction which
can be defined at mesh vertices, $\mathbf{x}$, by: 
\begin{equation}
\psi(\mathbf{x},t)=A|\mathbf{x}-\mathbf{x}_{c}|^{2}
\end{equation}
where $\mathbf{x}_{c}$ is the centre of the domain, and $A=5\pi/3000\text{ s}^{-1}$
so that the angular velocity is $2A$. The initial tracer takes a
Gaussian distribution in order to ensure that all advection schemes
achieve their theoretical order of accuracy:
\begin{equation}
\phi(\mathbf{x})=\exp\left(-\frac{1}{2}\frac{|\mathbf{x}-\mathbf{x}_{\phi}|^{2}}{r_{\phi}^{2}}\right)\label{eq:7}
\end{equation}
where $\mathbf{x}_{\phi}=\mathbf{x}_{c}+r_{c\phi}\mathbf{j}$ is the
initial centre of the tracer distribution, $r_{c\phi}=2500\text{ m}$,
$r_{\phi}=500\text{ m}$ and $\mathbf{j}$ is the unit vector in the
$y$ direction. The analytic solution has the same tracer distribution
but with the centre of the tracer at: 
\begin{equation}
\mathbf{x}_{\phi}=\mathbf{x}_{c}+r_{c\phi}\left(\begin{array}{c}
\cos(\pi/2+2At)\\
\sin(\pi/2+2At)
\end{array}\right)
\end{equation}
and the Gaussian rotates anti-clockwise exactly one revolution in
$600\text{ s}$.

The solid body rotation test case is performed on a uniform, orthogonal
mesh and on a non-orthogonal mesh on a plane with non-orthogonality
similar to that of a cubed-sphere mesh. For the dimensionally split
scheme, non-orthogonality is achieved using the co-ordinate transform:
\begin{equation}
\begin{array}{ccc}
X=x &  & Y=\begin{cases}
y_{m}\left(1+\frac{y-f}{2y_{m}-f}\right) & \text{ for }y\ge f\\
y_{m}\left(1+\frac{y-f}{f}\right) & \text{ for }y<f
\end{cases}\end{array}
\end{equation}
where $y_{m}=5000\text{ m}$ and $f$ is the equation for a $y$ position
of uniform $Y$ half way up the domain. In order to create angles
of $120^{o}$ in the mesh, similar to a cubed-sphere, $f$ is given
by:
\begin{equation}
f=\begin{cases}
y_{m}\left(1+\frac{1}{2\sqrt{3}}\right)-\frac{x}{\sqrt{3}} & \text{ for }x\le x_{m}\\
y_{m}\left(1-\frac{1}{2\sqrt{3}}\right)+\frac{x-x_{m}}{\sqrt{3}} & \text{ for }x>x_{m}
\end{cases}.\label{eq:f_for_Vmesh}
\end{equation}
where $x_{m}=5000$ m. For a $50\times50$ mesh this gives the $x$
and $y$ co-ordinate locations as shown in figure \ref{fig:cubedGrid}.
The multi-dimensional scheme model uses Cartesian co-ordinates and
a distorted mesh rather than a non-orthogonal co-ordinate system on
a Cartesian mesh. However this does not affect the numerical results
assuming that the co-ordinate transforms are implemented in a consistent
way to the distorted mesh in Cartesian co-ordinates.

For the dimensionally split scheme, bi-periodic boundary conditions
are applied. For the multi-dimensional scheme, it was more straightforward
to impose fixed value boundary conditions of $\phi=0$  where the
velocity is into the domain and zero normal gradient where the velocity
is out of the domain. However $\phi$ remains almost zero near the
boundaries so these boundary conditions do not affect the accuracy.

\begin{figure}
\begin{centering}
\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
\par\end{centering}

\caption{Two non-orthogonal mesh on a plane with $50\times50$ mesh points
and angles up to $120^{\text{o}}$ designed to have similar angles
to a cubed-sphere mesh. \label{fig:cubedGrid}}
\end{figure}

Results of this test case on the orthogonal and non-orthogonal meshes
of $100\times100$ cells with $\Delta t=1\text{ s}$ are shown in
figure \ref{fig:solidBody} for both advection schemes (which gives
a maximum Courant number close to one). The contours show the tracer
value every 100 seconds and the colours show the errors summed every
100 seconds. The dimensionally split scheme outperforms the multi-dimensional
scheme on both meshes due to the higher-order accuracy of the split
scheme. The dimensionally split scheme introduces a small error at
300 seconds where the tracer goes through the change in direction
of the mesh which would be ameliorated if we were using monotonicity
constraints. The second-order, multi-dimensional scheme shows phase
lag but errors are almost entirely insensitive the mesh distortions.

\begin{figure}
\begin{tabular}{cc}
Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_Tdiff}}\tabularnewline
\end{tabular}

\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=1\text{ s}$
giving a maximum Courant number of around 1. The contours show the
tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
The dashed contour shows the initial conditions. The min and max text
in each plot gives the minimum and maximum errors. \label{fig:solidBody}}
\end{figure}

The multi-dimensional and dimensionally split schemes take very different
approaches to handling large Courant numbers. The multi-dimensional
scheme uses implicit time-stepping whereas the dimensionally split
scheme uses a flux-form semi-Lagrangian approach, integrating over
a line of cells in order to calculate the flux across a face. Implicit
schemes are known to suffer from phase errors \citep[eg][]{DB12,LWW14}
for long time-steps whereas the accuracy of semi-Lagrangian is less
sensitive to time-step \citep{PS84}. Therefore we present results
of both schemes on orthogonal and non-orthogonal meshes for time-steps
10 times those used in figure \ref{fig:solidBody} ($\Delta t=10\text{ s}$)
giving maximum Courant numbers of around 10 using $100\times100$
cells in figure \ref{fig:solidBody-dt10}. The error of the multi-dimensional
scheme is again much larger than the dimensionally split scheme on
both meshes. The dimensionally split scheme is accurate at large Courant
numbers despite the first-order calculation of departure points. However,
the dimensionally split scheme introduces oscillations on the non-orthogonal
mesh, particularly where the mesh changes direction whereas for the
multi-dimensional scheme, errors are not strongly affected by the
non-orthogonality. Figure \ref{fig:solidBody-dt10} clearly shows
phase errors of the implicit time-stepping but, despite the large
Courant number, the well resolved part of the profile is propagating
at close to the correct speed. Dispersion analysis \citep{LWW14}
shows that high \emph{frequency} oscillations (which are poorly resolved
in time \emph{and} space) will be slowed dramatically but fast moving
features which are \emph{well resolved in space} will propagate at
a much more realistic speed, supporting the results shown in figure
\ref{fig:solidBody-dt10}.

\begin{figure}
\begin{tabular}{cc}
Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_TdiffBig_Tdiff}}\tabularnewline
\end{tabular}

\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$
giving a maximum Courant number of around 10. The contours show the
tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
The dashed contour shows the initial conditions. The min and max text
in each plot gives the minimum and maximum errors. \label{fig:solidBody-dt10}}
\end{figure}

In order to compare convergence with resolution of the different numerical
methods, we use the $\ell_{2}$ and $\ell_{\infty}$ error norms defined
in the usual way:
\begin{eqnarray}
\ell_{2} & = & \sqrt{\int_{V}\left(\phi-\phi_{T}\right)^{2}dV\bigg/\int_{V}\phi_{T}^{2}\ dV}\\
\ell_{\infty} & = & \max|\phi-\phi_{T}|\big/\max|\phi_{T}|
\end{eqnarray}
where $\phi_{T}$ is the analytic solution and the integrations and
maxima are over the whole domain, with volume $V$. Figure \ref{fig:solidBodyConverge}
shows convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$
error measures for meshes of $50\times50$, $100\times100$, $200\times200$
and $400\times400$ cells with time-steps scaled in order to maintain
a maximum Courant number of 1 ($\Delta t=2,\ 1,\ 0.5,\ 0.25\text{ s}$)
or scaled to achieve a maximum Courant number of 10 ($\Delta t=20,\ 10,\ 5,\ 2.5\text{ s}$).
The error norms are calculated at $t=500\ \text{seconds}$, when the
tracer has made $5/6$ of one revolution in order to avoid error cancellation.
On both orthogonal and non-orthogonal meshes, the multi-dimensional
scheme has second order convergence once errors are low enough to
avoid error saturation (stable errors are bounded at around one).
With a large Courant number, both schemes are less accurate. For the
multi-dimensional scheme, this is due to phase errors of the implicit
time-stepping whereas for the multi-dimensional scheme the first-order
errors in calculating the departure point and trajectory will be emerging
and there could also be significant errors from the second-order COSMIC
splitting. On the orthogonal mesh, the dimensionally split scheme
has  third order convergence for the Courant number close to one and
second order for the larger Courant number \citep[consistent with the results of ][]{CW84,LLM96}.
On the non-orthogonal mesh, the dimensionally split scheme has second
order converges for both Courant numbers.

\begin{figure}
\begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx1$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx1$\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC1} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC1}\tabularnewline
$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx10$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx10$\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC10} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC10}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Convergence with spatial resolution of the $\ell_{2}$ and $\ell_{\infty}$
errors for the solid body rotation test case at $t=500\ \text{seconds}$
($5/6$ of one revolution) on orthogonal and non-orthogonal meshes
using multi-dimensional and dimensionally split schemes. \label{fig:solidBodyConverge} }
\end{figure}

The different time-stepping schemes of the two models also affect
accuracy. The $\ell_{2}$ and $\ell_{\infty}$ error measures as a
function of time-step for meshes of $100\times100$ cells are shown
in figure \ref{fig:solidBodyConverge-dt}. The dimensionally split
scheme, which uses flux-form semi-Lagrangian time-stepping, has errors
reducing as time-step increases, up to a Courant number of 2 ($\Delta t=2\ \text{s}$)
whereas the multi-dimensional scheme, which uses the method of lines
to treat space and time separately, always has error reducing as time-steps
reduces. The flux-form semi-Lagrangian technique discretises space
and time together and the error is not very sensitive to time-step.
However, the shorter the time-step, the more time-steps need to be
taken so errors can actually accumulate more by taking more time-steps.
(This is consistent with the order of accuracy of semi-Lagrangian
being $\Delta x^{p}/\Delta t$ for interpolation using polynomials
of degree $p$, as described by \citet{Dur10}.)

\begin{figure}
\begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$
errors for the solid body rotation test case at $t=500\ \text{seconds}$
($5/6$ of one revolution) using orthogonal and non-orthogonal meshes
of $100\times100$ cells using multi-dimensional and dimensionally
split schemes. Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$
to $ $$c=10$ (for $\Delta t=10\ \text{s})$. \label{fig:solidBodyConverge-dt} }
\end{figure}

In summary, the dimensionally split scheme has excellent behaviour
at large and small Courant numbers on the orthogonal meshes with up
to third order convergence for small Courant numbers and the errors
increase and order of convergence decreases on non-orthogonal meshes.
In contrast, the multi-dimensional scheme is insensitive to the orthogonality,
converges with second-order and suffers from phase errors at large
Courant numbers.

\subsection{Horizontal Advection over Orography\label{sub:overOrog}}

Non-orthogonal meshes (or co-ordinate systems) are usually necessary
for representing orography which could be a challenge for dimensionally
split schemes. Horizontal-vertical split schemes are commonly used
in this context \citep[eg.][]{DEE+12,WGZ+13}. We present results
of the \citet{SLF+02} horizontal advection over orography test case
for a range of resolutions and Courant numbers for the dimensionally
split and multi-dimensional schemes.

All simulations use basic terrain following co-ordinates \citep[BTF,][]{GCS75}
in order to present a challenging test case that maximises the non-orthogonality.
The transformation is given by:
\begin{equation}
\begin{array}{ccc}
X=x &  & Z=H\frac{z-h(x)}{H-h(x)}\end{array}
\end{equation}
where $H$ is the domain height and $h$ is the terrain height. The
test case uses a domain of width 300\ km, height, $H=25\text{ km}$
and a mountain range defined by
\begin{equation}
h=\begin{cases}
h_{0}\cos^{2}\frac{\pi x}{\lambda}\cos^{2}\frac{\pi x}{2a} & \text{ for }|x|\le a\\
0 & \text{ otherwise}
\end{cases}
\end{equation}
with the maximum mountain height, $h_{0}=3\text{ km}$, half-width
$a=25\text{ km}$ and wavelength $\lambda=8\text{ km}$. These values
give a maximum terrain gradient of close to $45^{o}$. The wind is
given by a streamfunction which is defined at vertices so that the
wind field is discretely divergence free. The streamfunction at vertices
is calculated analytically from the wind profile:
\begin{equation}
u(z)=u_{0}\begin{cases}
1 & \text{ for }z_{2}\le z\\
\sin^{2}\left(\frac{\pi}{2}\frac{z-z_{1}}{z_{2}-z_{1}}\right) & \text{ for }z_{1}<z\le z_{2}\\
0 & \text{ for }z<z_{1}
\end{cases}
\end{equation}
with $u_{0}=10\text{ ms}^{-1}$, $z_{1}=4\text{ km}$ and $z_{2}=5\text{ km}$.
The initial tracer position is given by:
\begin{equation}
\phi=\begin{cases}
\cos^{2}\frac{\pi r}{2} & \text{ for }r\le1\\
0 & \text{ otherwise}
\end{cases}\text{ with }r=\sqrt{\left(\frac{x-x_{0}}{A_{x}}\right)^{2}+\left(\frac{z-z_{0}}{A_{z}}\right)^{2}}
\end{equation}
with initial tracer centre, $(x_{0},z_{0})=(-50\text{ km},\ 9\text{ km})$
and halfwidths $A_{x}=25\text{ km}$, $A_{z}=9\text{ km}$. At time
$t=5000\text{ s}$ the tracer is above the mountain and the simulation
finishes at $t=10,000\text{ s}$ by which time the analytic solution
is centred at $(50\text{km },8\text{km})$.
\begin{figure}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.14\textwidth}>{\centering}p{0.42\textwidth}c}
 & Dimensionally split & Multi-dimensional\tabularnewline
$\Delta t=25\ $s

max $c$ = 0.74

horiz $c$ =0.25

$c_{d}=0.22$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall}\tabularnewline
$\Delta t=50\ $s

max $c$ = 1.48

horiz $c$ =0.5

$c_{d}=0.44$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall}\tabularnewline
$\Delta t=100\ $s

max $c$ = 2.96

horiz $c$ =1

$c_{d}=0.87$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall}\tabularnewline
$\Delta t=200\ $s

max $c$ = 5.93

horiz $c$ =2

$c_{d}=1.75$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall}\tabularnewline
$\Delta t=500\ $s

max $c$ = 14.8

horiz $c$ =5

$c_{d}=4.4$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt500CN_up_10000_Tall}\tabularnewline
$\Delta t=1000\ $s

max $c$ = 29.6

horiz $c$ =10

$c_{d}=8.7$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt1000CN_up_10000_Tall}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.85\textwidth]{HilarysGraphics/vSlice_horizontalAdvection_orography_legends_Tall_TdiffAll}}\tabularnewline
\end{tabular*}

\caption{Tracer contours after 0, 5000 and 10000s for the horizontal advection
over orography and errors in colour. Different plots use different
time-steps. Spatial resolution is $\Delta x=1\ $km, $\Delta z=500\ $m.
Maximum Courant numbers, horizontal Courant number, $c$, and maximum
deformational Courant number, $c_{d}$ are also given. \label{fig:horizontalAdvectionWithDt}}
\end{figure}

The tracer advection over orography is shown in figure \ref{fig:horizontalAdvectionWithDt}
for the split and multi-dimensional schemes at a resolution of $\Delta x=1\text{ km}$,
$\Delta z=500\text{ m}$ and for a range of Courant numbers. The horizontal
Courant number is defined as $u_{0}\Delta t/\Delta x$ and ranges
from 0.25 to 10. The maximum Courant number is the maximum of the
multi-dimensional Courant number which is defined for cell with faces
$f$ as:
\begin{equation}
c=\frac{1}{2V}\sum_{f}|\mathbf{u}_{f}\cdot\mathbf{S}_{f}|\Delta t
\end{equation}
(see section \ref{sub:MMmolscheme} for definitions of variables)
and ranges from 0.74 to 29.6. The time-step restriction for the split
scheme is based on the deformational Courant number, $c_{d}\le1$,
(eqn \ref{eq:cd}). The maximum deformational Courant number is also
given in figure \ref{fig:horizontalAdvectionWithDt}. The contours
in figure \ref{fig:horizontalAdvectionWithDt} show the tracer values
at 0, $5000\text{ s}$ and $10,000\text{ s}$ after initialisation
and the colours show the errors from the analytic solution. For horizontal
Courant numbers less than one (maximum Courant number up to 3), both
schemes give accurate results with the dimensionally split scheme
tending to give oscillations and the multi-dimensional scheme producing
more diffusion. For larger Courant numbers, when the deformational
Courant number is greater than one, the split scheme is unstable while
the multi-dimensional scheme produces large phase errors due to the
errors associated with implicit time-stepping. The term responsible
for the large deformational Courant number is $\partial u/\partial z$
where the velocity shears from $u_{0}=10$m/s at $z=z_{2}$ to zero
at $z=z_{1}$.

We examine the convergence with resolution in figure \ref{fig:horizontalAdvectConverge}
which shows the $\ell_{2}$ and $\ell_{\infty}$ error norms as a
function of $\Delta x$. These simulations all use a maximum Courant
number less than one, a horizontal Courant number of 0.25, a maximum
deformational Courant number of about 0.2 and fixed ratios of $\Delta x$,
$\Delta z$ and $\Delta t$. Both schemes give similar accuracy with
the dimensionally split scheme having faster convergence with resolution.

\begin{figure}
\noindent \begin{centering}
\begin{tabular}{cc}
$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
\includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_l2errors} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_linferrors}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metrics
for horizontal advection over a 3000m and a 6000m mountains after
10,000s for the dimensionally split (PPM) scheme and multi-dimensional
MOL scheme, all using a horizontal Courant number of 0.25. Lines also
show theoretical first- and second-order convergence are also shown.
\label{fig:horizontalAdvectConverge}}
\end{figure}

In summary, the dimensionally split scheme has good accuracy over
orography for modest Courant numbers but larger errors for larger
Courant numbers and is unstable when the deformational Courant number
is greater than one whereas the multi-dimensional scheme with implicit
time-stepping is stable for all Courant numbers. The multi-dimensional
scheme second-order convergent and the dimensionally split scheme
converges faster.

\subsection{Deformational Flow \label{sub:deformFlow}}

In deformational flow, there is no analytical solution and therefore
we follow the approach of \citet{NL10,LSPT12} and define an evolving
velocity field that reverses direction half way through the simulation,
taking the tracer back to the initial conditions. Error norms can
then be calculated by comparing the final and initial tracer fields.
The deformational velocity field is added to a fixed solid body rotation
that does not reverse so as to avoid error cancellation between the
forwards and backwards periods (solid body here meaning horizontal
wind on a periodic domain).

We define a Cartesian version of the deformational, non-divergent
test case of \citet{LSPT12} with a domain between $-\pi$ and $\pi$
in the $x$ direction ($L_{x}=2\pi$) and between $-\frac{\pi}{2}$
and $\frac{\pi}{2}$ in the $y$ direction ($L_{y}=\pi$) with periodic
boundary conditions in the $x$ direction and zero gradient, zero
flow boundary conditions in the $y$ direction. The stream function
adapted to Cartesian co-ordinate is:
\begin{eqnarray}
\psi\left(x,y,t\right) & = & \frac{\hat{\psi}}{T}\left(\frac{L_{x}}{2\pi}\right)^{2}\sin^{2}\left(2\pi\left(\frac{x}{L_{x}}-\frac{t}{T}\right)\right)\cos^{2}\left(\pi\frac{y}{L_{y}}\right)\cos\left(\pi\frac{t}{T}\right)-L_{x}\frac{y}{T}
\end{eqnarray}
where $\hat{\psi}=10$ and $T=5$ is the time for one complete revolution
of the periodic domain for the solid body rotation part of the flow.
In order to test the order of convergence of the advection schemes,
we use the infinitely smooth Gaussian distribution for the initial
tracer concentration:
\begin{equation}
\phi=0.95\exp-\frac{|\mathbf{x}-\mathbf{x}_{0}|^{2}}{A}+0.95\exp-\frac{|\mathbf{x}-\mathbf{x}_{1}|^{2}}{A}
\end{equation}
where $\mathbf{x}^{T}=(x,y)$, $\mathbf{x}_{0}^{T}=(\frac{5}{12}L_{x},0)$,
$\mathbf{x}_{1}^{T}=(\frac{7}{12}L_{x},0)$ and $A=\frac{1}{5}$.

A mesh with distortion similar to a cubed-sphere is defined by the
co-ordinate transform:
\begin{equation}
\begin{array}{ccc}
X=x &  & Y=\begin{cases}
L_{y}\frac{y-f}{L_{y}-2f} & \text{ for }y\ge f\\
L_{y}\frac{y-f}{L_{y}+2f} & \text{ for }y<f
\end{cases}\end{array}
\end{equation}
where $f$ is:
\[
f=\begin{cases}
\frac{1}{\sqrt{3}}\left(\frac{\pi}{4}-|x|\right) & \text{ for }|x|\le\frac{\pi}{2}\\
\frac{1}{\sqrt{3}}\left(|x|-\frac{3\pi}{4}\right) & \text{ for }|x|>\frac{\pi}{2}
\end{cases}.
\]
The initial conditions and a $120\times60$ mesh with distortion defined
by this co-ordinate transform is shown in figure \ref{fig:deformGrid}.

\begin{figure}
\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_120x60_0_TUmesh}

\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_Tmesh_T}

\caption{The non-orthognal mesh/co-ordinate system for the deformational flow
with $120\times60$ cells and the initial tracer conditions.\label{fig:deformGrid}}
\end{figure}

The tracer concentrations after 1, 2, 3, 4 and 5 time units are shown
in figure \ref{fig:deformResults} using the dimensionally split and
multi-dimensional schemes on a non-orthogonal mesh of $480\times240$
cells, a time-step of 0.0025 units (ie 2000 time-steps to reach 5
time units) which gives a maximum Courant number of 1.03. The tracer
is stretched out, wound up, advected around and then wound back into
its original position with some numerical errors. Both advection schemes
preserve fine filaments but suffer from some dispersion errors which
generate small oscillations around zero behind sharp gradients in
the direction of the flow since neither scheme is monotonic or positive
preserving. The dimensionally split scheme returns the tracer to a
more accurate final solution.

\begin{figure}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
$t$ & Dimensionally split & Multi-dimensional\tabularnewline
1 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_1_T}\tabularnewline
2 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_2_T}\tabularnewline
3 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T3} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_3_T}\tabularnewline
4 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T4} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_4_T}\tabularnewline
5 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T5} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_5_T}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_T}}\tabularnewline
\end{tabular*}

\caption{Results of the deformational flow test case after 1, 2, 3, 4 and 5
time units on the non-orthogonal meshes of 480$\times$240 cells using
the multi-dimensional and split schemes. \label{fig:deformResults}}
\end{figure}

\begin{table}
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline 
Resolution & $60\times30$ & $120\times60$ & $240\times120$ & $480\times240$ & $960\times480$ & $1920\times960$\tabularnewline
\hline 
$\Delta t$ $\downarrow$ & \multicolumn{6}{c|}{Maximum Courant numbers {\footnotesize{}(deformational)} $\downarrow$}\tabularnewline
\hline 
\hline 
0.2 & 10.3 {\footnotesize{}(0.639)} &  &  &  &  & \tabularnewline
\hline 
0.1 &  & 10.3{\footnotesize{} (0.312)} &  &  &  & \tabularnewline
\hline 
0.05 &  & 5.17{\footnotesize{} (0.154)} & 10.3 {\footnotesize{}(0.154)} &  &  & \tabularnewline
\hline 
0.025 &  &  &  & 10.3{\footnotesize{} (0.077)} &  & \tabularnewline
\hline 
0.02 &  & 2.06{\footnotesize{} (0.061)} &  &  &  & \tabularnewline
\hline 
0.0125 & 1.03 &  &  &  & 10.3 & \tabularnewline
\hline 
0.01 &  & 1.03 {\footnotesize{}(0.030)} &  &  &  & \tabularnewline
\hline 
0.00625 &  &  &  &  &  & 10.3\tabularnewline
\hline 
0.005 &  & 0.517 {\footnotesize{}(0.015)} & 1.03{\footnotesize{} (0.015)} &  &  & \tabularnewline
\hline 
0.0025 &  & 0.254 {\footnotesize{}(0.008)} &  & 1.03{\footnotesize{} (0.008)} &  & \tabularnewline
\hline 
0.00125 &  &  &  &  & 1.03{\footnotesize{} (0.004)} & \tabularnewline
\hline 
0.000625 &  &  &  &  &  & 1.03\tabularnewline
\hline 
\end{tabular}

\caption{Resolutions, time-steps and Courant numbers for the non-orthogonal
meshes for the deformational flow test case. Maximum Courant numbers
are shown inside the table for simulations which are presented in
graphs. The maximum Courant numbers on the orthogonal meshes are 70\%
of those on the non-orthogonal meshes. The maximum deformational Courant
numbers are in brackets for some cases \label{tab:res_Courant}}
\end{table}

Sensitivity to orthogonality, resolution and time-step are explored
using a range of simulations with maximum Courant numbers shown in
table \ref{tab:res_Courant}. The Courant numbers on the uniform,
orthogonal meshes (with no mesh distortions) are 70\% of those on
the non-orthogonal mesh since the non-orthogonal mesh has clustering
of mesh points. The deformational Courant number is less than one
for all simulations. The convergence with resolution of the $\ell_{2}$
and $\ell_{\infty}$ error norms at the final time are shown in figure
\ref{fig:deformationalConverge} for both advection schemes using
a maximum Courant number close to one and for a maximum Courant number
close to ten. We will first consider the behaviour of the schemes
at modest Courant number (maximum close to one). All of the schemes
give first-order convergence with resolution at coarse resolutions
due to error saturation (if the scheme is stable, errors are bounded
above by about one). At higher resolution, the split scheme converges
with nearly third-order for both meshes whereas the multi-dimensional
scheme approaches second-order. For large Courant numbers (maximum
close to 10), the split scheme converges with first-order for both
mesh types due to the crude estimation of the departure point (section
\ref{sub:PPM}) whereas the multi-dimensional scheme is much less
accurate but approaches second-order  at high resolution. The dimensionally
split scheme is sensitive to mesh distortions only for large Courant
numbers.

\begin{figure}
\begin{tabular}{cc}
$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
\end{tabular}

\caption{Convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$
errors for the deformational flow on orthogonal and non-orthogonal
meshes using multi-dimensional and dimensionally split schemes. \label{fig:deformationalConverge} }
\end{figure}

We inspect sensitivity to time-step of both schemes on orthogonal
and non-orthogonal meshes of $120\times60$ cells in figure \ref{fig:deformational_dt}
using the time-steps shown in table \ref{tab:res_Courant} giving
maximum Courant numbers ranging from 0.25 to 10 (deformational Courant
numbers always less than 0.32). This demonstrates potentially useful
properties of the flux-form semi-Lagrangian time-stepping used by
the split scheme. Ignoring errors in calculating the departure point,
semi-Lagrangian schemes have errors proportional to $\Delta t^{-1}$
\citep{Dur10} which explains the reduction in error as time-step
increases for the split scheme. Once the maximum Courant number reaches
1 or 2 ($\Delta t=0.01$ or 0.02), errors of the split scheme do grow
with the time-step, due to the deformational nature of the flow and
errors in calculating the departure points. In contrast, using the
multi-dimensional scheme which uses method of lines time-stepping,
errors always increase as the time-step is increased. Comparisons
between schemes in figures \ref{fig:deformResults} and \ref{fig:deformationalConverge}
used maximum Courant numbers of 1 and 10 which showed that the split
scheme on the orthogonal mesh gave better accuracy than the multi-dimensional
scheme. However, figure \ref{fig:deformational_dt} shows that this
advantage disappears at lower Courant numbers since the multi-dimensional
scheme (method of lines) gets more accurate with lower Courant numbers
whereas the split scheme (semi-Lagrangian) gets less accurate since
more time-steps have to be taken. Both schemes are stable for all
time-steps considered.

\begin{figure}
\begin{tabular}{cc}
$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_dt} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_dt}\tabularnewline
\end{tabular}

\caption{Variation of $\ell_{2}$ and $\ell_{\infty}$ errors with time-step
for the deformational flow on orthogonal and non-orthogonal meshes
of $120\times60$ cells using multi-dimensional and dimensionally
split schemes. Different time-steps give maximum Courant numbers of
0.25, 0.5, 1, 2, 5 and 10. \label{fig:deformational_dt} }
\end{figure}

In summary, the dimensionally split scheme has good accuracy for deformational
flow independent of mesh orthogonality. The multi-dimensional scheme
is competitive at small Courant numbers but the semi-Lagrangian nature
of the split scheme means that errors are very low for Courant numbers
close to one.

\subsection{Computational Cost\label{sub:cost}}

We cannot compare CPU time, wall clock time or parallel efficiency
of the two advection schemes because the multi-dimensional scheme
is written in C++ using OpenFOAM and the split scheme is written in
Python, both codes have been run on different hardware and the split
advection scheme code has not been parallelised. Instead we consider
the number of multiply operations performed per cell per time-step
for calculating the fluxes of each tracer. We appreciate that this
is not a good predictor of efficiency as it does not consider memory
read and write requirements or cache coherency. However all data that
is multiplied has to be fetched from memory and so, assuming that
all data can be arranged optimally in memory to enable fewest cache
misses, the number of multiplies should be related to the wall clock
time. Neither advection scheme does significantly more work per memory
fetch than the other.

For both schemes we consider the number of multiplies needed to calculate
the flux of $\phi$ at each face. That is equation \ref{eq:PPMflux}
for the dimensionally scheme and equation (\ref{eq:multiDflux}) for
the multi-dimensional scheme. We do not consider the computational
cost of updating the cell averages from the fluxes using Gauss's theorem
(eqns (\ref{PPMgauss}) and (\ref{eq:multidGauss})) as these are
the same for each scheme. For the multi-dimensional scheme, we do
include an estimate of the amount of work done by the linear equation
solver but we do not consider the scalability of this solver.

\subsubsection{Dimensionally Split Scheme}

The computational cost of PPM with COSMIC splitting is not strongly
time-step dependent due to the swept area approach of flux-form semi-Lagrangian.
One-dimensional PPM uses 4 cells for interpolation to find face values
used by the reconstruction. Assuming as much as possible is pre-computed,
this interpolation uses 3 multiplies on a non-uniform mesh. The reconstruction
then uses 6 multiplies to calculate the flux on each face. Only one
additional memory access and one additional  multiply are needed per
cell for Courant numbers greater than one (eqn \ref{eq:cumulativeMass})
making 10 multiplies per cell for applying PPM in one direction for
a Cournat number greater than one. The COSMIC splitting in two dimensions
involves four applications of the one-dimensional PPM. This makes
40 multiplies per cell in total for applying PPM with COSMIC splitting
in two dimensions. In three dimensions, COSMIC splitting requires
12 applications of PPM \citep{LLM96} leading to 120 multiplies.

\subsubsection{Multi-dimensional Scheme}

The number of multiplies involved in the multi-dimensional scheme
includes the number of multiplies to update the higher order advection
and the number of multiplications for each iteration of the linear
equation solver. We will also consider the cost of an explicit version
of the multi-dimensional scheme using an RK2 time-stepping scheme
\citep[eg as used by][]{SW16} which is stable and accurate up to
a Courant number of one for this spatial discretistion and gives very
similar results to the implicit scheme (not shown).

The explicit version of the dimensionally split scheme uses RK2 or
Heun time-stepping, in which $\phi_{\text{up}}^{\prime}$ on the RHS
of eqn (\ref{eq:CN_1}) is replaced by $\phi_{\text{up}}^{n}$ and
$\phi_{\text{up}}^{n+1}$ on the RHS of eqn (\ref{eq:CN_2}) is replaced
by $\phi_{\text{up}}^{\prime}$. On a logically rectangular two-dimensional
mesh there are 12 cells in each stencil for each face (fig \ref{fig:stencil}).
Each cell has four faces and the interpolation onto each face is used
to calculate the flux between two cells. This leads to 24 multiplies
per cell for each RK2 stage and hence 48 multiplies per cell per time-step.
When using implicit time-stepping, there will be 24 multiplies per
cell for every evaluation of the right hand side of the matrix for
the higher-order correction on first-order upwind.

We have not explored the sensitivity of the accuracy and stability
to the stencil size and shape in three dimensions. The three dimensional
equivalent of the stencil of quadrilaterals in figure \ref{fig:stencil}
is likely to contain 36 (rather than 12) cells although it may be
possible to omit some corner cells and use a stencil of 20 cells.
For a stencil of 36 cells, the number of multiplies per cell per time-step
would be $36\times3\times2=216$.

The implicit version of the multi-dimensional scheme (section \ref{sub:implicitMethod})
requires the solution of an asymmetric, diagonally dominant matrix
with three non-zero elements per row for a logically rectangular two-dimensional
mesh. The pre-conditioner is implemented in file \url{DILUPreconditioner.C}
in OpenFOAM 3.0.1 and the solver in file \url{PBiCG.C}. From these
files, we estimate that, for a mesh of quadrilaterals, the solver
will use 24 multiplies (or divides) per cell, per solver iteration,
including pre-conditioning. The average number of iterations of the
preconditioned bi-CG solver per time-step for each of the simulations
is shown in table \ref{tab:iterations} (including the number of solver
iterations in each outer iteration). These simulations use two outer
iterations per time-step when the maximum Courant number is $\le1.1$
(as in equations (\ref{eq:CN_1}) and (\ref{eq:CN_2})) but, for stability,
use four outer iterations per time-step for the larger Courant numbers.
This partly explains the greater number of iterations for larger Courant
numbers. A solver tolerance of $10^{-8}$ is used for each of the
outer solves. Table \ref{tab:iterations} shows that the total number
of iterations per time-step reduces slightly as resolution increases.
The total number of iterations for a complete simulation is reduced
by using larger Courant numbers because the number of iterations per
time-step increases less than linearly with increasing Courant numbers.
In fact, simulations with larger Courant numbers are considerably
cheaper because there are fewer evaluations of the right hand side
of the matrix equation.

\noindent 
\begin{table}
\noindent \begin{centering}
\begin{tabular}{|c||c|c|c|c|c|}
\hline 
Resolution $\rightarrow$ & $120\times60$ & $240\times120$ & $480\times240$ & $960\times480$ & $1920\times960$\tabularnewline
\hline 
Max $c$ $\downarrow$ & \multicolumn{5}{c|}{Iterations per time-step orthogonal/non-orthogonal $\downarrow$}\tabularnewline
\hline 
\hline 
10 & 33.4/39.5 & 31.3/38.0 & 26.0/32.8 & 19.3/25.0 & 13.5/18.6\tabularnewline
\hline 
5 & 22.4/25.7 &  &  &  & \tabularnewline
\hline 
2 & 13.1/14.6 &  &  &  & \tabularnewline
\hline 
1 & 6.0/6.5 & 5.4/5.8 & 4.9/5.3 & 4.6/4.9 & 3.9/4.2\tabularnewline
\hline 
0.5 & 4.9/5.0 &  &  &  & \tabularnewline
\hline 
0.25 & 3.9/3.9 &  &  &  & \tabularnewline
\hline 
\end{tabular}
\par\end{centering}

\caption{Number of solver iterations per time-step of the multi-dimensional
scheme for all the simulations on orthogonal/non-orthogonal meshes.
\label{tab:iterations}}
\end{table}

Combining the number of solver iterations, the number of multiply
operations per solve and the number of multiply operations in calculating
the explicit higher order part of the advection scheme, the total
number of multiply operations per cell per time-step for the multi-dimensional
scheme is shown in table \ref{tab:multiplies} for the non-orthogonal
mesh of $120\times60$ cells. Table \ref{tab:multiplies} also shows
the number of multiplies for using the explicit, RK2 version of the
multi-dimensional scheme and the dimensionally split scheme.

\begin{table}
\noindent \begin{centering}
\begin{tabular}{|c|c|c|c|}
\hline 
Max $c$ & Implicit & Explicit & Dimensionally\tabularnewline
\hline 
 & \multicolumn{2}{c|}{Multi-dimensional} & Split\tabularnewline
\hline 
\hline 
10 & $39.5\times24+48\times2=1,044$ & - & 40\tabularnewline
\hline 
5 & $25.7\times24+48\times2=712.8$ & - & 40\tabularnewline
\hline 
2 & $14.6\times24+48\times2=446.4$ & - & 40\tabularnewline
\hline 
1 & $6.5\times24+48=204$ & 48 & 40\tabularnewline
\hline 
0.5 & $5.0\times24+48=168$ & 48 & 40\tabularnewline
\hline 
0.25 & $3.9\times24+48=$141.6 & 48 & 40\tabularnewline
\hline 
\end{tabular}
\par\end{centering}

\caption{Total number of multiply operations per cell per time-step for different
Courant numbers for the multi-dimensional scheme, an explicit version
of the multi-dimensional scheme (using RK2 or Heun time-stepping)
and the dimensionally split scheme on two-dimensional, logically rectangular
meshes. The multi-dimensional scheme using RK2 time-stepping is stable
for Courant numbers $\le1$ and gives very similar solutions to the
implicit version. \label{tab:multiplies}}
\end{table}

Table \ref{tab:multiplies} shows that the implicit scheme always
uses more multiply operations but particularly uses more multiplies
for large Courant numbers. The explicit version of the multi-dimensional
scheme always uses 48 multiply operations but is not stable for all
time-steps whereas the dimensionally split scheme (using flux-form
semi-Lagrangian time-stepping) is stable for all Courant numbers (at
this spatial resolution) and always uses the fewest number of multiply
operations per cell per time-step.

There is considerable flexibility in the solver configuration: the
number of outer iterations per time-step determines how frequently
the high-order correction on the right hand side of the matrix equation
is updated, and the solver tolerance per outer iteration could be
modified by using a weaker tolerance on all but the final matrix solve
per time-step. These options have not been explored. It may also be
beneficial to create more non-zero matrix entries rather than having
the higher-order correction entirely a deferred correction on first-order
upwind, but such a change would need to ensure that the matrix remains
diagonally dominant.

\section{Summary and Conclusions\label{sec:concs}}

We examine the errors associated with using a dimensionally split
advection scheme and a multi-dimensional advection scheme on distorted
meshes. The dimensionally split scheme is very accurate on orthogonal
meshes and only loses a little accuracy on highly distorted meshes,
despite a first-order departure point calculation. The multi-dimensional
scheme with implicit time-stepping is less accurate on orthogonal
meshes than the dimensionally split scheme but the accuracy is not
sensitive to mesh distortions and the stability is less sensitive
to Courant number.

The dimensionally split scheme is the piecewise polynomial method
\citep[PPM,][]{CW84} with COSMIC splitting \citep{LLM96} that extends
it to two spatial dimensions. PPM is converges with third-order in
one dimension and COSMIC splitting enables second-order convergence
in two orthogonal directions. PPM is a flux-form semi-Lagrangian scheme
and so can handle large Courant numbers accurately without significant
additional computational cost, with a time-step restriction based
on the deformational Courant number. The second-order accurate multi-dimensional
scheme is split in space and time (method of lines) and uses a cubic
polynomial fit over a stencil of cells for spatial discretisation
and trapezoidal implicit in time to retain stability for large Courant
numbers. We use versions of both schemes without any monotonicity
constraints in order to compare the handling of multi-dimensionality
of the two schemes and order of convergence rather than comparing
the limiters of the two schemes.

Three two-dimensional advection test cases on Cartesian planes are
proposed without the complexities of a spherical domain or multi-panel
meshes but with distorted meshes to mimic the distortions of a cubed-sphere
or terrain following co-ordinates. We therefore propose that these
test cases could be used in the initial testing of advection schemes,
before the generation of meshes on the sphere. The first test case
is an extension of the \citet{LLM96} solid body rotation using a
distorted mesh. The second test case is the established horizontal
advection over orography \citep{SLF+02} using a basic terrain following
mesh in order to maximise the distortions and a version using higher
orography. The third test case is the deformational flow test case
of \citet{LSPT12} adapted to a planar, Cartesian domain and a distorted
mesh. We use the version of this test case with smooth initial conditions
(the sum of two Gaussians) in order to examine order of convergence.

The dimensionally split scheme is extremely accurate on orthogonal
meshes and retains accuracy when long time-steps are used. However
on distorted meshes, particularly at changes in direction such as
those that appear at cube sphere edges or over orography, the split
scheme loses some accuracy. In contrast, the multi-dimensional scheme
is almost entirely insensitive to mesh distortion and asymptotes to
second-order convergence at high resolution. As is expected for implicit
time-stepping, phase errors occur when using long time-steps but the
spatially well resolved features are advected at the correct speed
and the multi-dimensional scheme is always stable.

The matrix solver associated with the implicit time-stepping of the
multi-dimensional scheme means that it is always considerably more
expensive than the split scheme with cost increasing with Courant
number. We haven't investigated sensitivity to solver and pre-conditioner
choices and it may be possible to do  better. But the flux-form semi-Lagrangian
method enables long time-steps without any matrix solutions and this
will always be difficult to beat. It is possible to use the multi-dimensional
scheme with explicit time-stepping such as Runge-Kutta. In that case,
the Courant number is restricted to be less than one but the cost
is similar to the dimensionally split scheme.

The conclusions of this paper are consistent with those of \citet{KNK15}
who found a dimensionally split scheme to be as accurate as a multi-dimensional
scheme on a cubed-sphere mesh with special interpolations so that
the dimensionally split scheme could cope with the cube edges. In
addition, we find that special treatment is not needed at cubed-sphere
edges to maintain accuracy when using dimension splitting.  When dimensionally
split schemes are used on cubed-sphere meshes, they do usually have
special interpolations at cube edges \citep[eg][]{LR96}. However
orography is everywhere and so special treatment over steep orography
would not be practical.

The dimensionally split scheme is hard to beat, providing close to
third-order convergence even in the presence of mesh distortions and
can be very cheaply extended for large Courant numbers.

\bibliographystyle{abbrvnat}
\bibliography{numerics}

\end{document}

### 2017-01-19 makeFigs.sh

@@ -1,0 +1,14 @@
+#!/bin/bash -e
+cd figures
+
+for file in stencil sbrc1 sbrc10 sbr_dx sbr_dt overMountains overMountains_dx deform_init deform deform_dx deform_dt; do
+    fileR=`filename -r $file`
+    lyx --export pdflatex -f $fileR
+    pdflatex $fileR
+    pdfcrop $fileR.pdf ../$fileR.pdf
+    evince ../$fileR.pdf &
+done
+
+cd ..
+
+cp -u HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf .

### 2017-01-24 advectionPaper.tex
Version submitted to QJRMS and to arxiv

@@
 %% Do not edit unless you really know what you are doing.
-\documentclass[12pt,a4paper,english,british]{qjrms4}
+\documentclass[12pt,a4paper,english,british,times,doublespace]{qjrms4}
 \usepackage{mathptmx}
 \usepackage[T1]{fontenc}
-\usepackage[utf8]{luainputenc}
+\usepackage[latin9]{inputenc}
 \setcounter{secnumdepth}{3}
 \setcounter{tocdepth}{3}
-\usepackage{color}
-\usepackage{array}
-\usepackage{verbatim}
 \usepackage{url}
@@
 \providecommand{\tabularnewline}{\\}
-%% A simple dot to overcome graphicx limitations
-\newcommand{\lyxdot}{.}
 
@@
 
+\usepackage{lineno}
+%\linenumbers
+
 \makeatother
@@
 \begin{document}
-
-\title{Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme
-for Atmospheric Transport}
-
-\author{Yumeng Chen, Hilary Weller, Stephen Pring and James Shaw}
-
-%\maketitle
-\begin{comment}
-\tableofcontents{}
-
-\listoftodos
-\end{comment}
-
-\begin{abstract}
-Dimensionally split advection schemes are attractive for atmospheric
-modelling due to their efficiency and accuracy in each spatial dimension.
-Accurate long time steps can be achieved without significant cost
-using the flux-form semi-Lagrangian technique. The dimensionally split
-scheme used in this paper is constructed from the one-dimensional
-Piecewise Parabolic Method and extended to two dimensions using COSMIC
-splitting. The dimensionally split scheme is compared with a genuinely
-multi-dimensional, method of lines scheme with implicit time-stepping
-which is stable for very large Courant numbers.
-
-Two-dimensional advection test cases on Cartesian planes are proposed
-that avoid the complexities of a spherical domain or multi-panel meshes.
-These are solid body rotation, horizontal advection over orography
-and deformational flow. The test cases use distorted meshes either
-to represent sloping terrain or to mimic the distortions of a cubed
-sphere.
-
-Such mesh distortions are expected to accentuate the errors associated
-with dimension splitting, however, the dimensionally split scheme
-is very accurate on orthogonal meshes and accuracy decreases only
-a little in the presence of large mesh distortions. The dimensionally
-split scheme retains accuracy when long time-steps are used. The multi-dimensional
-scheme is almost entirely insensitive to mesh distortions and asymptotes
-to second-order accuracy at high resolution. As is expected for implicit
-time-stepping, phase errors occur when using long time-steps but the
-spatially well resolved features are advected at the correct speed
-and the multi-dimensional scheme is always stable.
-
-A naive estimate of computational cost (number of multiplies) reveals
-that the implicit scheme is  the most expensive, particularly for
-large Courant numbers. If the multi-dimensional scheme is used instead
-with explicit time-stepping, the Courant number is restricted to less
-than one but the cost becomes similar to the dimensionally split scheme.
+\title{Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme for Atmospheric Transport}
+\author{Yumeng Chen,\affil{a} Hilary Weller,\affil{b}\corrauth\ Stephen Pring\affil{c} and James Shaw\affil{b}}
+
+\runningheads{Y. Chen, H. Weller, S. Pring and J. Shaw}{Dimension splitting for Advection}
+
+\address{
+\affilnum{a}Universit\"at Hamburg\\
+\affilnum{b}Meteorology, University of Reading \\
+\affilnum{c}Met Office, UK}
+\corraddr{E-mail: <h.weller@reading.ac.uk>}
+
+\begin{abstract}\normalfont
+Dimensionally split advection schemes are attractive for atmospheric modelling due to their efficiency and accuracy in each spatial dimension.
+Accurate long time steps can be achieved without significant cost using the flux-form semi-Lagrangian technique.
+The dimensionally split scheme used in this paper is constructed from the one-dimensional Piecewise Parabolic Method and extended to two dimensions using COSMIC splitting.
+The dimensionally split scheme is compared with a genuinely multi-dimensional, method of lines scheme with implicit time-stepping which is stable for very large Courant numbers.
+
+Two-dimensional advection test cases on Cartesian planes are proposed that avoid the complexities of a spherical domain or multi-panel meshes.
+These are solid body rotation, horizontal advection over orography and deformational flow.
+The test cases use distorted meshes either to represent sloping terrain or to mimic the distortions of a cubed sphere.
+
+Such mesh distortions are expected to accentuate the errors associated with dimension splitting, however, the dimensionally split scheme is very accurate on orthogonal meshes and accuracy decreases only a little in the presence of large mesh distortions.
+The dimensionally split scheme also loses some accuracy when long time-steps are used.
+The multi-dimensional scheme is almost entirely insensitive to mesh distortions and asymptotes to second-order accuracy at high resolution.
+As is expected for implicit time-stepping, phase errors occur when using long time-steps but the spatially well resolved features are advected at the correct speed and the multi-dimensional scheme is always stable.
+
+A naive estimate of computational cost (number of multiplies) reveals that the implicit scheme is  the most expensive, particularly for large Courant numbers.
+If the multi-dimensional scheme is used instead with explicit time-stepping, the Courant number is restricted to less than one but the cost becomes similar to the dimensionally split scheme.
 \end{abstract}
-\begin{keywords}
-hello
-\end{keywords}
+
+\keywords{
+Multi-dimensional; advection; stability; accuracy; long time-steps
+}
+\maketitle
+
+%\tableofcontents
+%\listoftodos
+\onecolumn
 
@@
 \noindent \begin{centering}
-\scalebox{0.5}[0.5]{\input{figs/cubicUpCPCstencil.pstex_t}}\hfill{}\scalebox{0.5}[0.5]{\input{figs/hexStencil.pstex_t}}
+\includegraphics{stencils}
 \par\end{centering}
@@
 \begin{centering}
-\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
+\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
 \par\end{centering}
@@
 \begin{figure}
-\begin{tabular}{cc}
-Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
-\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
-Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
-\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_Tdiff}}\tabularnewline
-\end{tabular}
-
-\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
-meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=1\text{ s}$
-giving a maximum Courant number of around 1. The contours show the
-tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
-The dashed contour shows the initial conditions. The min and max text
-in each plot gives the minimum and maximum errors. \label{fig:solidBody}}
+\includegraphics{sbrc1.png}
+
+\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=1\text{ s}$ giving a maximum Courant number of around 1.
+The contours show the tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
+The dashed contour shows the initial conditions.
+The min and max text in each plot gives the minimum and maximum errors.
+\label{fig:solidBody}}
 \end{figure}
@@
 \begin{figure}
-\begin{tabular}{cc}
-Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
-\includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
-Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
-\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{HilarysGraphics/solidBodyRotationOnPlane_legends_TdiffBig_Tdiff}}\tabularnewline
-\end{tabular}
-
-\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
-meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$
-giving a maximum Courant number of around 10. The contours show the
-tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
-The dashed contour shows the initial conditions. The min and max text
-in each plot gives the minimum and maximum errors. \label{fig:solidBody-dt10}}
+\includegraphics{sbrc10.png}
+
+\caption{Results of the solid body rotation test case on orthogonal and non-orthogonalmeshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$ giving a maximum Courant number of around 10.
+The contours show the tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
+The dashed contour shows the initial conditions.
+The min and max text in each plot gives the minimum and maximum errors.
+\label{fig:solidBody-dt10}}
 \end{figure}
@@
 \begin{centering}
-\begin{tabular}{cc}
-$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx1$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx1$\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC1} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC1}\tabularnewline
-$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx10$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx10$\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errorsC10} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrorsC10}\tabularnewline
-\end{tabular}
+\includegraphics{sbr_dx}
 \par\end{centering}
 
-\caption{Convergence with spatial resolution of the $\ell_{2}$ and $\ell_{\infty}$
-errors for the solid body rotation test case at $t=500\ \text{seconds}$
-($5/6$ of one revolution) on orthogonal and non-orthogonal meshes
+\caption{Convergence with spatial resolution of the $\ell_{2}$ and $\ell_{\infty}$ errors for the solid body rotation test case at $t=500\ \text{seconds}$ ($5/6$ of one revolution) on orthogonal and non-orthogonal meshes
 using multi-dimensional and dimensionally split schemes. \label{fig:solidBodyConverge} }
@@
 \begin{centering}
-\begin{tabular}{cc}
-$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
-\end{tabular}
+\includegraphics{sbr_dt}
 \par\end{centering}
 
-\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$
-errors for the solid body rotation test case at $t=500\ \text{seconds}$
-($5/6$ of one revolution) using orthogonal and non-orthogonal meshes
-of $100\times100$ cells using multi-dimensional and dimensionally
-split schemes. Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$
-to $ $$c=10$ (for $\Delta t=10\ \text{s})$. \label{fig:solidBodyConverge-dt} }
+\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$errors for the solid body rotation test case at $t=500\ \text{seconds}$ ($5/6$ of one revolution) using orthogonal and non-orthogonal meshes of $100\times100$ cells using multi-dimensional and dimensionally split schemes.
+Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$ to $ $$c=10$ (for $\Delta t=10\ \text{s})$.
+\label{fig:solidBodyConverge-dt} }
 \end{figure}
@@
 is centred at $(50\text{km },8\text{km})$.
+
 \begin{figure}
-\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.14\textwidth}>{\centering}p{0.42\textwidth}c}
- & Dimensionally split & Multi-dimensional\tabularnewline
-$\Delta t=25\ $s
-
-max $c$ = 0.74
-
-horiz $c$ =0.25
-
-$c_{d}=0.22$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall}\tabularnewline
-$\Delta t=50\ $s
-
-max $c$ = 1.48
-
-horiz $c$ =0.5
-
-$c_{d}=0.44$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall}\tabularnewline
-$\Delta t=100\ $s
-
-max $c$ = 2.96
-
-horiz $c$ =1
-
-$c_{d}=0.87$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall}\tabularnewline
-$\Delta t=200\ $s
-
-max $c$ = 5.93
-
-horiz $c$ =2
-
-$c_{d}=1.75$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall}\tabularnewline
-$\Delta t=500\ $s
-
-max $c$ = 14.8
-
-horiz $c$ =5
-
-$c_{d}=4.4$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt500CN_up_10000_Tall}\tabularnewline
-$\Delta t=1000\ $s
-
-max $c$ = 29.6
-
-horiz $c$ =10
-
-$c_{d}=8.7$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt1000CN_up_10000_Tall}\tabularnewline
-\multicolumn{3}{c}{\includegraphics[width=0.85\textwidth]{HilarysGraphics/vSlice_horizontalAdvection_orography_legends_Tall_TdiffAll}}\tabularnewline
-\end{tabular*}
-
-\caption{Tracer contours after 0, 5000 and 10000s for the horizontal advection
-over orography and errors in colour. Different plots use different
-time-steps. Spatial resolution is $\Delta x=1\ $km, $\Delta z=500\ $m.
-Maximum Courant numbers, horizontal Courant number, $c$, and maximum
-deformational Courant number, $c_{d}$ are also given. \label{fig:horizontalAdvectionWithDt}}
+\includegraphics{overMountains.png}
+
+\caption{Tracer contours after 0, 5000 and 10000s for the horizontal advection over orography and errors in colour.
+Different plots use different time-steps.
+Spatial resolution is $\Delta x=1\ $km, $\Delta z=500\ $m.
+Maximum Courant numbers, horizontal Courant number, $c$, and maximum deformational Courant number, $c_{d}$ are also given.
+\label{fig:horizontalAdvectionWithDt}}
 \end{figure}
@@
 \noindent \begin{centering}
-\begin{tabular}{cc}
-$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
-\includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_l2errors} & \includegraphics[width=0.49\textwidth]{HilarysGraphics/orography_linferrors}\tabularnewline
-\end{tabular}
+\includegraphics{overMountains_dx}
 \par\end{centering}
 
-\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metrics
-for horizontal advection over a 3000m and a 6000m mountains after
-10,000s for the dimensionally split (PPM) scheme and multi-dimensional
-MOL scheme, all using a horizontal Courant number of 0.25. Lines also
-show theoretical first- and second-order convergence are also shown.
+\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metricsfor horizontal advection over a 3000m and a 6000m mountains after 10,000s for the dimensionally split (PPM) scheme and multi-dimensional MOL scheme, all using a horizontal Courant number of 0.25.
+Lines also show theoretical first- and second-order convergence are also shown.
 \label{fig:horizontalAdvectConverge}}
@@
 \begin{figure}
-\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_120x60_0_TUmesh}
-
-\includegraphics[width=1\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_Tmesh_T}
-
-\caption{The non-orthognal mesh/co-ordinate system for the deformational flow
-with $120\times60$ cells and the initial tracer conditions.\label{fig:deformGrid}}
+\includegraphics{deform_init}
+
+\caption{The non-orthognal mesh/co-ordinate system for the deformational flow with $120\times60$ cells and the initial tracer conditions.\label{fig:deformGrid}}
 \end{figure}
@@
 \begin{figure}
-\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
-$t$ & Dimensionally split & Multi-dimensional\tabularnewline
-1 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_1_T}\tabularnewline
-2 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_2_T}\tabularnewline
-3 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T3} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_3_T}\tabularnewline
-4 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T4} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_4_T}\tabularnewline
-5 & \includegraphics[width=0.49\textwidth,valign=m]{YumengsGraphics/deformationalData/plots/T5} & \includegraphics[width=0.49\textwidth,valign=m]{HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_5_T}\tabularnewline
-\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{HilarysGraphics/deformingAdvectionCyclic_legends_T}}\tabularnewline
-\end{tabular*}
-
-\caption{Results of the deformational flow test case after 1, 2, 3, 4 and 5
-time units on the non-orthogonal meshes of 480$\times$240 cells using
-the multi-dimensional and split schemes. \label{fig:deformResults}}
+\includegraphics{deform.png}
+
+\caption{Results of the deformational flow test case after 1, 2, 3, 4 and 5 time units on the non-orthogonal meshes of 480$\times$240 cells using the multi-dimensional and split schemes.
+\label{fig:deformResults}}
 \end{figure}
@@
 \begin{figure}
-\begin{tabular}{cc}
-$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
-\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
-$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
-\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
-\end{tabular}
-
-\caption{Convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$
-errors for the deformational flow on orthogonal and non-orthogonal
-meshes using multi-dimensional and dimensionally split schemes. \label{fig:deformationalConverge} }
+\includegraphics{deform_dx}
+
+\caption{Convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$ errors for the deformational flow on orthogonal and non-orthogonal meshes using multi-dimensional and dimensionally split schemes.
+\label{fig:deformationalConverge} }
 \end{figure}
@@
 \begin{figure}
-\begin{tabular}{cc}
-$\ell_{2}$ error & $\ell_{\infty}$ error\tabularnewline
-\includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_l2errors_dt} & \includegraphics[width=0.48\linewidth]{HilarysGraphics/deformingAdvectionCyclic_plots_linferrors_dt}\tabularnewline
-\end{tabular}
-
-\caption{Variation of $\ell_{2}$ and $\ell_{\infty}$ errors with time-step
-for the deformational flow on orthogonal and non-orthogonal meshes
-of $120\times60$ cells using multi-dimensional and dimensionally
-split schemes. Different time-steps give maximum Courant numbers of
-0.25, 0.5, 1, 2, 5 and 10. \label{fig:deformational_dt} }
+\includegraphics{deform_dt}
+
+\caption{Variation of $\ell_{2}$ and $\ell_{\infty}$ errors with time-step for the deformational flow on orthogonal and non-orthogonal meshes of $120\times60$ cells using multi-dimensional and dimensionally split schemes.
+Different time-steps give maximum Courant numbers of 0.25, 0.5, 1, 2, 5 and 10.
+\label{fig:deformational_dt} }
 \end{figure}

### 2017-01-24 makeFigs.sh

@@ -12,3 +12,10 @@ done
 cd ..
 
 cp -u HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf .
+
+eps2png2 deform.pdf
+eps2png2 sbrc10.pdf
+eps2png2 sbrc1.pdf
+
+zip figures.zip stencil.pdf solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf sbrc1.png sbrc10.png sbr_dx.pdf sbr_dt.pdf overMountains.png overMountains_dx.pdf deform_init.pdf deform.png deform_dx.pdf deform_dt.pdf
+

### 2017-04-11 revision/reply_to_reviewer.tex
a first start to reply to the reviewers

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,a4paper,english,british,times,doublespace]{qjrms4}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{esint}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\pdfpageheight\paperheight
\pdfpagewidth\paperwidth

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}

\usepackage{todonotes}

\usepackage{amssymb}

\usepackage{lineno}
%\linenumbers

\makeatother

\usepackage{babel}
\begin{document}
Q: quasi-uniform resolution doesn't help better exploit modern architectures; they're largely orthogonal concepts. Quasi-uniform resolution just improves the global time step restriction.

A: Quasi-uniform resolution indeed help better exploit modern architectures. The lat-lon grid leads to convergence of resolution toward poles and restricts the time stepping. These problems are usually solved by semi-implicit semi-Lagrangian schemes. The modern computers use increasing number of cores and the efficiency is limited by the scalability. Implicit solvers and semi-Lagrangian advection result in massive data communications \citep{ST12}, which limits the scalability around poles.

\bibliographystyle{abbrvnat}
\bibliography{numerics}

\end{document}

### 2017-04-24 revision/reply_to_reviewer.tex
organizing comments from reviewers

@@
 \begin{document}
+Q: Might want to say "distorted, non-orthogonal meshes" to make it's clear that you are indeed directly testing non-orthogonality.
+
+Yumeng: Changed as suggested.
+
 Q: quasi-uniform resolution doesn't help better exploit modern architectures; they're largely orthogonal concepts. Quasi-uniform resolution just improves the global time step restriction.
 
-A: Quasi-uniform resolution indeed help better exploit modern architectures.
+Yumeng: Quasi-uniform resolution indeed help better exploit modern architectures.
 The lat-lon grid leads to convergence of resolution toward poles and restricts the time stepping.
@@
 
+Q: I feel the strong need for a caveat here.
+I appreciate the desire to create a simpler test case not involving spherical geometry, because the cubed-sphere really is easy to mess up.
+However, my personal experience is that dimensional splitting  on the cubed-sphere incurs significantly more error compared to dimensional splitting on a plane.
+This could potentially be because spherical geometry coupled with non-orthogonality incurs more splitting error than non-orthogonality alone.
+Without directly confirming the applicability of this framework to the cubed-sphere by correlating error measures between them, I think the authors need to qualify more clearly that this applicability remains a *hypothesis.*
+
+Yumeng: The error on the sphere indeed is larger than on the plane for dimensionally splitting scheme, even on the lat-lon grid.
+On cubed sphere, the mesh around corner of each panel has abrupt changes and is not differentiable.
+\cite{TN17} implemented shallow water equations on generalized curvilinear coordinate on the plane as a test before the implementation on the sphere.
+The comparison between plane and non-orthogonal mesh with abrupt change is to show the difference between orthogonal grid and non-orthogonal grid, which has similar feature as cubed sphere.
+The test certainly cannot be equivalent to the comparison between orthogonal lat-lon grid and cubed-sphere grid.
+
+Q: Why would remapping otherwise be needed? This sounds like Yin-Yang kind of stuff. If so, please explicitly mention that.
+
+Yumeng:  The remapping here refers to integrating the reconstruction function to get the mass fluxes along the trajectory of each cell face.
+The cumulative mass technique \citep{CW84,LLM95} avoids the mass calculation along the trajectory and we just need to get the integral at departure cell with simple add/subtraction from cumulative mass.
+We did not remap between different grids like Yin-Yang.
+
+Q: temporally second-order accurate
+
+Yumeng: changed as suggested
+
+Q: There's another paper that actually does do dimensional splitting on the cubed-sphere:
+
+<http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>
+
+And they incur very large L-inf errors when the advected shape passes cubed-sphere corners.
+Their errors are smaller with their Strang carry-over scheme than one would experience with an alternating Strang splitting, but the errors are still quite high.
+Also, these errors at the cubed-sphere corners reduce significantly at smaller time steps, as expected, but it puts a bit of a damper on the desire to use larger time steps.
+
+Yumeng: 
+On this paper, on page 460: 
+"It is difficult to design a numerical scheme that preserves a constant field in the dimensional splitting framework.
+It is our ongoing work to design a nonsplitting SLDG scheme that preserves the constant field when the velocity field is nondivergent." 
+Is this because of the filter or the splitting? How is the performance of the scheme under 1-D divergnet flow, because the 2-D non-divergent field could lead to 1-D divergent flow in each dimension.
+
+On page 464, "It is clear that the second order of convergence is observed, which comes from the splitting error."
+I believe in solid body rotation test case on Cartesian grid, the splitting error is zero in COSMIC splitting if the splitting scheme is 2nd order in time.
+So I think the splitting scheme is not comparable to the COSMIC splitting.
+
+On page 469, "This is an indication that such error comes from the dimensional splitting, and there exists a certain symmetry property for different dimensional-splitting orderings.
+Such symmetry property, together with the symmetry of the cosine bell profile, may contribute to the dropping of the $l_\infty$ error after the cosine bell passed the corner."
+If this is the reason for the large error, COSMIC splitting may give a better results as COSMIC splitting get a symmetric results by averaging two splitting scheme.
+
+Q: It's important to note that all Schar et al did was compress all of the errors to the lowest few levels, where the metric Jacobian determinants vary even more.
+
+A: ??
+
+Q: Remapping is where the error in incurred here. The advantage is that one can do that less often with floating layers.
+
+A: ??
+
+Q: Ullrich & Norman, 2012 QJRMS performed a multi-dimensional FFSL method with hyperdiffusion out to a CFL of roughly 2.5
+
+A: ??
+
+Q: I'm not so sure this is true.
+The reconstruction cost would be the same regardless.
+It's the integration cost that would be proportional with CFL.
+Of course, as the CFL increases, integration increasingly becomes more and more of the model cost.
+So this is asymptotically true, but not necessarily true for more moderate CFL numbers.
+
+A: ??
+
+Q: A third disadvantage is that they require MPI_all_reduce for the linear solves, which is increasingly prohibitive in its expense on large computers.
+
+A: ??
+
+Q: I believe this has been solved with SSP implicit RK and SSP DIRK schemes, though, right?
+
+A:
+
+Q: It's worth mentioning that significantly less special treatment is needed for element-based schemes (e.g., Galerkin) because no halo is needed.
+They merely transform vector quantities at element edges that coincide with panel boundaries.
+
+A:
+
+Q: What do you mean by maximally? Do you mean "highly", or is something being optimized here?
+
+A:
+
+Q: Thanks for mentioning this. Do you think using quadrature in time would be worthwhile? It'd have to be separate over each individual cell because it requires continuity, so it'd be pretty expensive
+
+A:
+
+Q: Is this done for sake of expense, since advective SL schemes only need one upwind evaluation, and not an integral; or are the other reasons for this?
+
+A: Using the same scheme both for inner and outer operator can avoid stability problem. PPM is a finite volume scheme. It is easier to make use of PPM in both inner and outer operator.
+
+Q: I find it confusing that you show a hexagonal mesh here and yet are only using rectangular meshes in the study.
+
+A: ??
+
+Q: Can you use a_1, a_2, etc here, because "i" has a meaning that could otherwise be confusing.
+
+A: ??
+
+Q: It seems to me that the implicit scheme just isn't a very good comparison point. I think it would be more helpful to see a smaller time step, fully explicit, multi-dimensional comparison point.
+
+If the purpose of this paper were solely to introduce new test cases, I think the existing methods would be OK.
+But since the purpose is to also introduce a new scheme, I think having a better comparison point is important.
+Since for many models, the state of the art is using smaller-time-step explicit schemes, it's important to see how the accuracy compares to that.
+
+A:
+
+Q: Have you thought about using a more accurate implicit time stepping than CN? It would probably improve the phase errors.
+
+A:
+
+Q: This is really a comparative statement, and I think the baseline is fairly poor.
+
+A:
+
+Q: Does it really "maximize" the non-orthogonality or just make for a "highly" non-orthogonal test case?
+
+A:
+
+Q: I think it's worth mentioning why we dimensionally split off the vertical: the high aspect ratio present in most weather / climate models.
+
+A:
+
+Q: I'd remove words like "very" and "extremely." They aren't well-defined, and it's better to use comparative language with a clear baseline.
+
+A:
+
+Q: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: <http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>.
+Additionally, Katta et al was cited as a dimensionally split scheme, but it is not.
+The reconstruction is performed in dimensional sweeps, but the scheme itself is multi-dimensional.
+
+A:
+
+Q: Most importantly, to say the newly proposed dimensionally split scheme is accurate compared to multi-dimensional schemes, I believe a better comparison point is needed.
+Users select the scheme that gives the best results with the fewest resources.
+I believe the authors need to include a time-explicit semi-discrete (method of lines) multi-dimensional scheme with cubic interpolants using smaller time steps to serve as a comparison point for this scheme.
+I know the focus is on large time steps, but it's necessary to consider if one can, in fact, get more accuracy for a given amount of work by simply using the status quo with smaller time steps.
+
+A:
+
+Q: It is stated that results in this 2-D framework will generalize well to the cubed-sphere, but I believe that in order to truly make that statement, one would have to implement a given scheme on both meshes and then compare.
+My experience on the cubed-sphere mesh is that it.
+
+A:
+
+Q:  However, I am concerned that the comparison is somewhat limited by the fact that the multi-dimensional scheme is of lower-order than the dimensionally-split scheme.
+How can the lower order of accuracy of the multi-dimensional scheme justified?
+
+A:
+
+Q: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
+The effect of this is most visible in Figure 4, where the more accurate dimensionally-split scheme has more grid-scale noise than the lower-order multi-dimensional scheme, likely because of the larger implicit diffusion in the lower-order scheme.
+A proper monotone FFSL scheme would greatly reduce these ripples with only a modest introduction of diffusive errors.
+
+A:
+
+Q: Given that a fundamentally new scheme is not presented here, a better title might be something like, “Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps”
+
+A:
+
+Q: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
+LR96 is designed for cartesian and lat-lon grids.
+Also, only LR96 uses FFSL methods, while PL07 has dropped the semi-Lagrangian extension (which was most useful near the poles of lat-lon grids).
+
+A:
+
+Q: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
+
+A:
+
+Q: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
+
+A: 
 \bibliographystyle{abbrvnat}

### 2017-04-24 firstDraft/advectionPaper.tex
Re-arranged the files so that the first draft is in subdirectory firstDraft

--- 2017-01-24 advectionPaper.tex
@@
 These are solid body rotation, horizontal advection over orography and deformational flow.
-The test cases use distorted meshes either to represent sloping terrain or to mimic the distortions of a cubed sphere.
+The test cases use \textcolor{red}{distorted, non-orthognal} meshes either to represent sloping terrain or to mimic the distortions of a cubed sphere.
 
@@
 }One-dimensional schemes can be used with operator splitting to create
-dimensionally split, second-order accurate schemes \citep[eg][]{LLM96}
+dimensionally split, \textcolor{red}{temporally} second-order accurate schemes \citep[eg][]{LLM96}
 on logically rectangular, multidimensional meshes. Dimensionally split

### 2017-04-24 firstDraft/qjrms/qjrms4doc.tex
Re-arranged the files so that the first draft is in subdirectory firstDraft

% qjrms4doc.tex V1.10, 4 October 2013

\documentclass[times]{qjrms4}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{moreverb}

\def\volumeyear{2013}
%\def\volumenumber{00}

\begin{document}

\runningheads{A.~N.~Other}{A demonstration of the \emph{Q.~J.~R.
Meteorol. Soc.} class file}

\title{A demonstration of the \LaTeXe\ class file for
the \itshape{Quarterly Journal of the Royal Meteorological
Society}\footnotemark[2]}

\author{A.~N.~Other\corrauth}

\address{John Wiley \& Sons, Ltd, The Atrium, Southern Gate, Chichester,
West Sussex, PO19~8SQ, UK}

\corraddr{Journals Production Department, John Wiley \& Sons, Ltd,
The Atrium, Southern Gate, Chichester, West Sussex, PO19~8SQ, UK.}

\begin{abstract}
This paper describes the use of the \LaTeXe\ \textsf{qjrms4.cls}
class file for setting papers for the \emph{Quarterly Journal of
the Royal Meteorological Society}.
\end{abstract}

\keywords{class file; \LaTeXe; \emph{Q.~J.~R. Meteorol. Soc.}}

\maketitle

\footnotetext[2]{Please ensure that you use the most up to date
class file,
available from the QJRMS Home Page at\\
\href{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1477-870X}{\tiny\texttt{http://onlinelibrary.wiley.com/journal/10.1002/(ISSN)1477-870X}}%
}

\section{Introduction} Many authors submitting to research journals
now use \LaTeXe\ to prepare their papers. This paper describes the
\textsf{qjrms4.cls} class file which can be used to convert
articles produced with other \LaTeXe\ class files into the correct
form for publication in the \emph{Quarterly Journal}.

The \textsf{qjrms4.cls} class file preserves much of the standard
\LaTeXe\ interface so that any document which was produced using
the standard \LaTeXe\ \textsf{article} style can easily be
converted to work with the \textsf{qjrms4} style. However, the
width of text and typesize will vary from that of
\textsf{article.cls}; therefore, \emph{line breaks will change}
and it is likely that displayed mathematics and tabular material
will need re-setting.

In the following sections we describe how to lay out your code to
use \textsf{qjrms4.cls} to reproduce the typographical look of the
\emph{Quarterly Journal}. However, this paper is not a guide to
using \LaTeXe\ and we would refer you to any of the many books
available (see, for example, \cite{R1}, \cite{R2} and \cite{R3}).

\section{The three golden rules} Before we proceed, we would like to
stress \emph{three golden rules} that need to be followed to
enable the most efficient use of your code at the typesetting
stage:
\begin{enumerate}
\item[(i)] keep your own macros to an absolute minimum;
\item[(ii)] as \TeX\ is designed to make sensible spacing
decisions by itself, do \emph{not} use explicit horizontal or
vertical spacing commands, except in a few accepted (mostly
mathematical) situations, such as \verb"\," before a
differential~d, or \verb"\quad" to separate an equation from its
qualifier;
\item[(iii)] follow the \emph{Quarterly Journal}
reference style.
\end{enumerate}

\section{Getting started} The \textsf{qjrms4} class file should run
on any standard \LaTeXe\ installation. If any of the fonts, class
files or packages it requires are missing from your installation,
they can be found on the \emph{\TeX\ Collection} DVDs or from
CTAN.

The \emph{Quarterly Journal} is published using a proprietary
font. A reasonable match can be achieved by using the Times fonts
and this is achieved by using the \verb"times" option as
\verb"\documentclass[times]{qjrms4}". If for any reason you have a
problem using Times you can easily resort to Computer Modern fonts
by removing the \verb"times" option.

\section{The article header information}
The heading for any file using \textsf{qjrms4.cls} is shown in
Figure~\ref{F1}.

\begin{figure*}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\documentclass[times]{qjrms4}
%\documentclass[times,doublespace]{qjrms4}%For paper submission

\begin{document}

\runningheads{<Initials and Surnames>}{<Short title>}

\title{<Initial cap, lower case>}

\author{<An Author,\affil{a}
Someone Else\affil{b}\corrauth\ and Perhaps Another\affil{a}>}

\address{<\affilnum{a}First author's address
(in this example it is the same as the third author)\\
\affilnum{b}Second author's address>}

\corraddr{<Corresponding author's address (the second author in
this example)>. E-mail: <corresponding author's email address>}

\begin{abstract}
<Text>
\end{abstract}

\keywords{<List up to eight key words>}

\maketitle

\section{Introduction}
.
.
.
\end{boxedverbatim}
\end{center}
\caption{Example header text.\label{F1}}
\end{figure*}

\subsection{Remarks}
\begin{enumerate}
\item[(i)] In \verb"\runningheads", keep the short title to no
more than 50 characters; use `\emph{et~al.}' if there are three or
more authors.

\item[(ii)] Note the use of \verb"\affil" and \verb"\affilnum" to
link names and addresses. The author for correspondence is marked
by \verb"\corrauth" and \verb"\corraddr" is used to give that
author's address, which will be printed prefaced by
`Correspondence to:'.

\item[(iii)] For submitting a double-spaced manuscript, add
\verb"doublespace" as an option to the documentclass line.

\item[(iv)] Keywords are separated by semicolons.
\end{enumerate}

\section{The body of the article}

\subsection{Mathematics} \textsf{qjrms4.cls} makes the full
functionality of \AmS\/\TeX\ available. We encourage the use of
the \verb"align", \verb"gather" and \verb"multline" environments
for displayed mathematics.

\subsection{Figures and tables} \textsf{qjrms4.cls} uses the
\textsf{graphicx} package for handling figures.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}
Recall that the\\
\verb"\begin{figure*}...\end{figure*}"\\
commands are needed for a
figure spanning both columns.

For further details on how to size figures, etc, with the
\textsf{graphicx} package see, for example, \cite{R1} or
\cite{R3}. If figures are available in an acceptable format (for
example, .eps, .ps) they will be used but a printed version should
always be provided. \medbreak

The standard coding for a table is shown in Figure~\ref{F2}.

\begin{figure}
\begin{verbatim}
\begin{table}
\caption{<Table caption>}
\centering
\begin{tabular}{<table alignment>}
\toprule
<column headings>\\
\midrule
<table entries
(separated by & as usual)>\\
<table entries>\\
.
.
.\\
\bottomrule
\end{tabular}
\end{table}
\end{verbatim}
\caption{Example table layout.\label{F2}}
\end{figure}

\subsection{Cross-referencing}
The use of the \LaTeX\ cross-reference system
for figures, tables, equations, etc, is encouraged
(using \verb"\ref{<name>}" and \verb"\label{<name>}").

\subsection{Acknowledgements}
An Acknowledgements section is started with \verb"\ack" or
\verb"\acks" for \textit{Acknowledgement} or
\textit{Acknowledgements}, respectively. It must be placed just
before the References (or before the appendix when applicable).

\subsection{Bibliography}
Please note that the file \textsf{wileyqj.bst} is available from
the same download page for those authors using \BibTeX.

Otherwise, the normal commands for producing the reference list
are:
\begin{verbatim}
\begin{thebibliography}{99}
\item <Reference details>
.
.
.
\item <Reference details>
\end{thebibliography}
\end{verbatim}

\subsection{Double spacing}
If you need to double space your document for submission please
use the \verb+doublespace+ option as shown in the sample layout in
Figure~\ref{F1}.

\section{Support for \textsf{qjrms4.cls}}
We offer on-line support to participating authors.
Please contact us via e-mail at
\href{mailto:qjauth-cls@wiley.co.uk}{\texttt{qjauth-cls@wiley.co.uk}}.

We would welcome any feedback, positive or otherwise, on your
experiences of using \textsf{qjrms4.cls}.

\section{Copyright statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ John Wiley \& Sons, Ltd, The Atrium,
Southern Gate, Chichester, West Sussex, PO19~8SQ, UK.  All rights reserved.

\subsection{Rules of use}
This class file is made available for use by authors who wish to
prepare an article for publication in the \emph{Quarterly Journal
of the Royal Meteorological Society} published by John Wiley \&
Sons, Ltd.  The user may not exploit any part of the class file
commercially.

This class file is provided on an \emph{as~is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantablility or fitness for a particular purpose. There will
be no duty on the author[s] of the software or  John Wiley \&
Sons, Ltd to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\ack This class file was developed by Sunrise Setting Ltd,
Paignton, Devon, UK. Website:\\
\href{http://www.sunrise-setting.co.uk}{\texttt{www.sunrise-setting.co.uk}}

\begin{thebibliography}{99}
\bibitem[Kopka and Daly(2003)]{R1}
Kopka~H, Daly~PW. 2003. \emph{A Guide to \LaTeX} (4th~edn).
Addison-Wesley.

\bibitem[Lamport(1994)]{R2}
Lamport~L. 1994. \emph{\LaTeX: a Document Preparation System}
(2nd~edn). Addison-Wesley.

\bibitem[Mittelbach and Goossens(2004)]{R3}
Mittelbach~F, Goossens~M. 2004. \emph{The \LaTeX\ Companion}
(2nd~edn). Addison-Wesley.

\end{thebibliography}
\end{document}

### 2017-04-24 revision/advectionPaper.tex
a first start to reply to the reviewers
Re-arranged the files so that the first draft is in subdirectory firstDraft

--- 2017-01-24 advectionPaper.tex
@@
 These are solid body rotation, horizontal advection over orography and deformational flow.
-The test cases use distorted meshes either to represent sloping terrain or to mimic the distortions of a cubed sphere.
+The test cases use \textcolor{red}{distorted non-orthogonal meshes} either to represent sloping terrain or to mimic the distortions of a cubed sphere.
 
@@
 
+\section*{Acknowledgements}
+
+Thanks to Nigel Wood of the UK Met Office for advice and proof reading. Weller acknowledges support from NERC grant NE/K006797/1.
+
 \bibliographystyle{abbrvnat}

### 2017-04-24 revision/figures/deform12.tex
Re-arranged the files so that the first draft is in subdirectory firstDraft

%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{array}
\usepackage[pdftex]{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amssymb}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
$t$ & Dimensionally split & Multi-dimensional\tabularnewline
1 & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_1_T}\tabularnewline
2 & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_nonOrthogW_c1_480x240_2_T}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{/home/hilary/latex/myPapers/YumengHilaryAdvectionPaper/HilarysGraphics/deformingAdvectionCyclic_legends_T}}\tabularnewline
\end{tabular*}
\end{document}

### 2017-04-24 comments/initialSolidBody/initialSolidBody.py

@@ -1,0 +1,45 @@
+from __future__ import division
+import numpy as np
+import matplotlib.pyplot as plt
+
+nx = 51
+ny = 1
+Lx = 1e4
+Ly = 0
+X = np.linspace(0, Lx, nx)
+Y = np.linspace(0, Ly, ny)
+dx = X[1] - X[0]
+dy = Y[1] - Y[0]
+
+# tracer distribution
+A= np.pi*5/3/1000.
+r = 2500.
+theta0 = np.pi/2.
+x0 = 0.5*nx*dx+r*np.cos(theta0)
+y0 = 0.5*ny*dy+ r*np.sin(theta0)
+phi = np.exp(- 0.5*(((X-x0)/500)**2 + ((Y-y0)/500)**2))
+
+# streamfunction
+r = ((X-0.5*dx)-(0.5*nx)*dx)**2+((Y-0.5*dy)-0.5*Ly)**2
+r1 = 3500.
+r2 = 4500.
+psi = np.where(r<=r1**2, A*r, \
+    np.where((r>r1**2) & (r<=r2**2), A*r1*np.sqrt(r), A*r1*r2))
+
+
+Also, I tried    
+# tracer distribution
+    A= np.pi*5/3/1000.
+    r = 2500.
+    theta0 = np.pi/2.
+    x0 = 0.5*nx*dx+r*np.cos(theta0)
+    y0 = 0.5*ny*dy+ r*np.sin(theta0)
+    phi = np.exp(- 0.5*(((X-x0)/500)**2 + ((Y-y0)/500)**2))
+
+# streamfunction
+    r = np.sqrt(((X-0.5*dx)-(0.5*nx)*dx)**2+((Y-0.5*dy)-0.5*Ly)**2)
+    r1 = 3500.
+    r2 = 4500.
+    psi = np.where(r<=r1, A*r**2, \
+        np.where((r>r1) & (r<=r2), A*r1*r, A*r1*r2))
+

### 2017-04-27 revision/reply_to_reviewer.tex
modifying advectionPaper.lyx and export to tex file and reformatt of reply_to_reviewer.tex

--- 2017-04-11 revision/reply_to_reviewer.tex
@@
-%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
-%% Do not edit unless you really know what you are doing.
-\documentclass[12pt,a4paper,english,british,times,doublespace]{qjrms4}
+\documentclass[12pt,a4paper]{article}
 \usepackage{mathptmx}
@@
 \usepackage[authoryear]{natbib}
-
-\makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-\pdfpageheight\paperheight
-\pdfpagewidth\paperwidth
-
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
 
@@
 
-\usepackage{babel}
+% \usepackage{babel}
 \begin{document}
-Q: quasi-uniform resolution doesn't help better exploit modern architectures; they're largely orthogonal concepts. Quasi-uniform resolution just improves the global time step restriction.
+\begin{description}
+\item [Page 1, Line 41]: Might want to say ``distorted, non-orthogonal meshes'' to make it's clear that you are indeed directly testing non-orthogonality.
 
-A: Quasi-uniform resolution indeed help better exploit modern architectures.
+\item [Yumeng]: Changed as suggested.
+
+\item [Page 2, Line 4]: quasi-uniform resolution doesn't help better exploit modern architectures; they're largely orthogonal concepts.
+Quasi-uniform resolution just improves the global time step restriction.
+
+\item [Yumeng]: Quasi-uniform resolution indeed help better exploit modern architectures.
 The lat-lon grid leads to convergence of resolution toward poles and restricts the time stepping.
@@
 
+\item [Page 2, Line 27]: I feel the strong need for a caveat here.
+I appreciate the desire to create a simpler test case not involving spherical geometry, because the cubed-sphere really is easy to mess up.
+However, my personal experience is that dimensional splitting  on the cubed-sphere incurs significantly more error compared to dimensional splitting on a plane.
+This could potentially be because spherical geometry coupled with non-orthogonality incurs more splitting error than non-orthogonality alone.
+Without directly confirming the applicability of this framework to the cubed-sphere by correlating error measures between them, I think the authors need to qualify more clearly that this applicability remains a *hypothesis.*
+
+\item [Yumeng]: The error on the sphere indeed is larger than on the plane for dimensionally splitting scheme, even on the lat-lon grid.
+On cubed sphere, the mesh around corner of each panel has abrupt changes and is not differentiable.
+\cite{TN17} implemented shallow water equations on generalized curvilinear coordinate on the plane as a test before the implementation on the sphere.
+The comparison between plane and non-orthogonal mesh with abrupt change is to show the difference between orthogonal grid and non-orthogonal grid, which has similar feature as cubed sphere.
+The test certainly cannot be equivalent to the comparison between orthogonal lat-lon grid and cubed-sphere grid.
+
+\item [Page 2, Line 40]: Why would remapping otherwise be needed? This sounds like Yin-Yang kind of stuff. If so, please explicitly mention that.
+
+\item [Yumeng]:  The remapping here refers to integrating the reconstruction function to get the mass fluxes along the trajectory of each cell face.
+The cumulative mass technique \citep{CW84,LLM95} avoids the mass calculation along the trajectory and we just need to get the integral at departure cell with simple add/subtraction from cumulative mass.
+We did not remap between different grids like Yin-Yang.
+
+\item [Page 2, Line 43]: temporally second-order accurate
+
+\item [Yumeng]: changed as suggested
+
+\item [Page 2, Line 45]: The scheme in this paper isn't really dimensionally split, only the *reconstruction* is.
+Katta et al use dimensionally split sweeps to create a genuinely multi-dimensional interpolant (minus a few of the cross terms), and then they evolve in a genuinely multi-dimensional way.
+This is an important distinction because (1) they do not incur time error in the same way as truly split schemes and (2) they include a diagonal sweep for some of the cross terms, whereas truly split schemes generally do not.
+
+\item [Yumeng]: Agreed. We deleted the reference and the description herein.
+
+\item [Page 2, Line 46]: There's another paper that actually does do dimensional splitting on the cubed-sphere:
+
+<http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>
+
+And they incur very large L-inf errors when the advected shape passes cubed-sphere corners.
+Their errors are smaller with their Strang carry-over scheme than one would experience with an alternating Strang splitting, but the errors are still quite high.
+Also, these errors at the cubed-sphere corners reduce significantly at smaller time steps, as expected, but it puts a bit of a damper on the desire to use larger time steps.
+
+\item [Yumeng]: 
+On this paper, on page 460: 
+``It is difficult to design a numerical scheme that preserves a constant field in the dimensional splitting framework.
+It is our ongoing work to design a nonsplitting SLDG scheme that preserves the constant field when the velocity field is nondivergent.''
+Is this because of the filter or the splitting? How is the performance of the scheme under 1-D divergnet flow, because the 2-D non-divergent field could lead to 1-D divergent flow in each dimension.
+
+On page 464, ``It is clear that the second order of convergence is observed, which comes from the splitting error.''
+I believe in solid body rotation test case on Cartesian grid, the splitting error is zero in COSMIC splitting if the splitting scheme is 2nd order in time.
+So I think the splitting scheme is not comparable to the COSMIC splitting.
+
+On page 469, ``This is an indication that such error comes from the dimensional splitting, and there exists a certain symmetry property for different dimensional-splitting orderings.
+Such symmetry property, together with the symmetry of the cosine bell profile, may contribute to the dropping of the $l_\infty$ error after the cosine bell passed the corner.''
+If this is the reason for the large error, COSMIC splitting may give a better results as COSMIC splitting get a symmetric results by averaging two splitting scheme.
+
+\item [Page 2, Line 55]: It's important to note that all Schar et al did was compress all of the errors to the lowest few levels, where the metric Jacobian determinants vary even more.
+
+\item [Yumeng]: We add words here as suggested.
+
+\item [Page 2, Line 57]: Remapping is where the error in incurred here. The advantage is that one can do that less often with floating layers.
+
+\item [Yumeng]: We add words here as suggested.
+
+\item [Page 3, Line 4]: Ullrich \& Norman, 2012 QJRMS performed a multi-dimensional FFSL method with hyperdiffusion out to a CFL of roughly 2.5
+
+\item [Yumeng]: Not treated.
+
+\item [Page 3, Line 7]: I'm not so sure this is true.
+The reconstruction cost would be the same regardless.
+It's the integration cost that would be proportional with CFL.
+Of course, as the CFL increases, integration increasingly becomes more and more of the model cost.
+So this is asymptotically true, but not necessarily true for more moderate CFL numbers.
+
+\item [Yumeng]: The cost is proportional to the CFL because the flux-form scheme for higher CFL has to integrate larger areas as the trajectory of departure cell is longer.
+
+\item [Page 3, Line 14]: A third disadvantage is that they require MPI\_all\_reduce for the linear solves, which is increasingly prohibitive in its expense on large computers.
+
+\item [Yumeng]: We added statements as suggested.
+
+\item [Page 3, Line 14]: I believe this has been solved with SSP implicit RK and SSP DIRK schemes, though, right?
+
+\item [Yumeng]: We deleted the relevant statement for monotonicity of method of lines schemes.
+
+\item [Page 3, Line 31]: It's worth mentioning that significantly less special treatment is needed for element-based schemes (e.g., Galerkin) because no halo is needed.
+They merely transform vector quantities at element edges that coincide with panel boundaries.
+
+\item [Yumeng]: We added statements as suggested.
+
+\item [Page 3, Line 45]: What do you mean by maximally? Do you mean ``highly'', or is something being optimized here?
+
+\item [Yumeng]: We changed it to highly distorted grids.
+
+\item [Page 5, Line 1]: Thanks for mentioning this.
+Do you think using quadrature in time would be worthwhile?
+It'd have to be separate over each individual cell because it requires continuity, so it'd be pretty expensive
+
+\item [Yumeng]: Yes, higher order scheme for departure point would be expensive. On the other hand, the original flux-form semi-Lagrangian scheme by \citep{LR96} also just uses first order Euler.
+
+\item [Page 5, Line 27]: Is this done for sake of expense, since advective SL schemes only need one upwind evaluation, and not an integral; or are the other reasons for this?
+
+\item [Yumeng]: Using the same scheme both for inner and outer operator can avoid stability problem.
+Such treatment makes the use of PPM in both inner and outer operator easier because finite volume scheme requires flux at cell boundaries instead of flux at cell center.
+
+\item [Page 6, Line 15]: I find it confusing that you show a hexagonal mesh here and yet are only using rectangular meshes in the study.
+
+\item [Yumeng]: Leave it to Hilary.
+
+\item [Page 6, Line 40]: Can you use $a_1$, $a_2$, etc here, because ``$i$'' has a meaning that could otherwise be confusing.
+
+\item [Yumeng]: Leave it to Hilary.
+
+\item [Page 8, Line 57]: It seems to me that the implicit scheme just isn't a very good comparison point.
+I think it would be more helpful to see a smaller time step, fully explicit, multi-dimensional comparison point.
+
+If the purpose of this paper were solely to introduce new test cases, I think the existing methods would be OK.
+But since the purpose is to also introduce a new scheme, I think having a better comparison point is important.
+Since for many models, the state of the art is using smaller-time-step explicit schemes, it's important to see how the accuracy compares to that.
+
+\item [Yumeng]: When time step is 1s, the Courant number is less than one and most explicit scheme without semi-Lagrangian extension should be able to handle such Courant number.
+Phase error is only obvious for Courant number larger than one.
+In the case of $\Delta t = 1$ s, a fully explicit, multi-dimensional scheme for smaller time step will lead to similar results.
+
+\item [Page 10, Line 55]: Have you thought about using a more accurate implicit time stepping than CN? It would probably improve the phase errors.
+
+\item [Yumeng]: Yes, a better time stepping scheme may improve phase error, but it may also introduce a more complicated scheme.
+The reason we chose CN is due to the consideration of complex of implementation.
+The implementation difficulty of CN is comparable to COSMIC splitting with PPM.
+
+\item [Page 13, Line 1]: This is really a comparative statement, and I think the baseline is fairly poor.
+
+\item [Yumeng]: As stated, the baseline is based on the complex of implementation.
+
+\item [Page 13, Line 19]: Does it really ``maximize'' the non-orthogonality or just make for a ``highly'' non-orthogonal test case?
+
+\item [Yumeng]: We use such statement because the BTF has the maximum distortion compared with other terrain-following coordinate or other treatment of orography.
+
+\item [Page 15, Line 36]: I think it's worth mentioning why we dimensionally split off the vertical: the high aspect ratio present in most weather/climate models.
+
+\item [Yumeng]: We add explanation at the start of this section as suggested.
+
+\item [Page 22, Line 41]: I'd remove words like ``very'' and ``extremely.'' They aren't well-defined, and it's better to use comparative language with a clear baseline.
+
+\item [Yumeng]: We remove the word extremely.
+
+\item [E-mail, Reviewer 1]: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: <http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>.
+Additionally, Katta et al was cited as a dimensionally split scheme, but it is not.
+The reconstruction is performed in dimensional sweeps, but the scheme itself is multi-dimensional.
+
+\item [Yumeng]: We removed the citation and relevant description.
+
+\item [E-mail, Reviewer 1]: Most importantly, to say the newly proposed dimensionally split scheme is accurate compared to multi-dimensional schemes, I believe a better comparison point is needed.
+Users select the scheme that gives the best results with the fewest resources.
+I believe the authors need to include a time-explicit semi-discrete (method of lines) multi-dimensional scheme with cubic interpolants using smaller time steps to serve as a comparison point for this scheme.
+I know the focus is on large time steps, but it's necessary to consider if one can, in fact, get more accuracy for a given amount of work by simply using the status quo with smaller time steps.
+
+\item [Yumeng]: The method of lines scheme is chosen to achieve similar amount of work as dimensionally split scheme.
+PPM on uniform grid only need a 4-point stencil for each cell edge in 1-D, while cubic interpolants will lead to a much larger stencil in 2-D.
+The results should be similar for smaller time steps as we already have tests for Courant number less than one.
+On the other hand, dimensionally splitting scheme only uses first-order Euler, which is less accurate than Crank-Nicolson method for method of lines scheme.
+
+\item [E-mail, Reviewer 1]: It is stated that results in this 2-D framework will generalize well to the cubed-sphere, but I believe that in order to truly make that statement, one would have to implement a given scheme on both meshes and then compare.
+My experience on the cubed-sphere mesh is that it.
+
+\item [Yumeng]: The non-orthogonal mesh has similar feature as edges at cubed-sphere: discontinuity in Jacobian.
+The results should be able to show a glimpse of the property of COSMIC splitting, especially for long time steps.
+
+\item [E-mail, Reviewer 2]: The first questions that come to my mind when reading this paper are, ``why another advection scheme?'' and ``why another advection test case?'' given the multitude of existing schemes and tests.
+The demonstration of the good accuracy of standard dimensionally-split FFSL schemes is certainly useful, especially considering the well-publicized shortcomings of such methods.
+
+\item [Yumeng]:
+
+\item [E-mail, Reviewer 2]:  However, I am concerned that the comparison is somewhat limited by the fact that the multi-dimensional scheme is of lower-order than the dimensionally-split scheme.
+How can the lower order of accuracy of the multi-dimensional scheme justified?
+
+\item [Yumeng]:
+
+\item [E-mail, Reviewer 2]: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
+The effect of this is most visible in Figure 4, where the more accurate dimensionally-split scheme has more grid-scale noise than the lower-order multi-dimensional scheme, likely because of the larger implicit diffusion in the lower-order scheme.
+A proper monotone FFSL scheme would greatly reduce these ripples with only a modest introduction of diffusive errors.
+
+\item [Yumeng]: Yes, monotonicity preservation is important for modern solvers.
+The shape change is more likely to be phase error from time stepping scheme.
+However, it is unfair to compare scheme with limiter for FFSL with multi-dimensional scheme without proper limiter.
+
+\item [E-mail, Reviewer 2]: Given that a fundamentally new scheme is not presented here, a better title might be something like, ``Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps''
+
+\item [Yumeng]:
+
+\item [E-mail, Reviewer 2]: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
+LR96 is designed for cartesian and lat-lon grids.
+Also, only LR96 uses FFSL methods, while PL07 has dropped the semi-Lagrangian extension (which was most useful near the poles of lat-lon grids).
+
+\item [Yumeng]: One wrong reference of this kind at last page has changed to Lin (2007).
+
+\item [E-mail, Reviewer 2]: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
+
+\item [Yumeng]:  Yes.(?)
+
+\item [E-mail, Reviewer 2]: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
+\end{description}
 \bibliographystyle{abbrvnat}

### 2017-04-27 revision/advectionPaper.tex
modifying advectionPaper.lyx and export to tex file and reformatt of reply_to_reviewer.tex

@@
-%% LyX 2.0.8.1 created this file.  For more info, see http://www.lyx.org/.
+%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
 %% Do not edit unless you really know what you are doing.
@@
 \setcounter{tocdepth}{3}
+\usepackage{color}
 \usepackage{url}
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
@@
 \usepackage[authoryear]{natbib}
+\PassOptionsToPackage{normalem}{ulem}
+\usepackage{ulem}
 
@@
 
-\usepackage{amssymb}
-
 \usepackage{lineno}
-%\linenumbers
+\linenumbers
 
@@
 mass fluxes along the domain \foreignlanguage{english}{\citep{LLM95}.
-}This is done without remapping which can be an expensive procedure,
-finding a conservative map between fields on overlapping meshes\foreignlanguage{english}{.
+}This is done without remapping \textcolor{red}{on each cell a long
+trajectory} which can be an expensive procedure\textcolor{red}{\sout{,
+finding a conservative map between fields on overlapping meshes}}\foreignlanguage{english}{.
 }One-dimensional schemes can be used with operator splitting to create
@@
 on logically rectangular, multidimensional meshes. Dimensionally split
-schemes have been found to give good accuracy on non-orthogonal meshes
-such as the cubed-sphere \citep{PL07,KNK15} with special treatment
-over cube edges. \citet{PL07} use the average of two one-sided schemes
-at cube edges whereas \citet{KNK15} create ghost cells outside each
-cube panel boundary. Without this special treatment, dimensionally
-split schemes do not account properly for mesh distortion and errors
-may occur. Accuracy between second and fourth order was found in practice
-on a range of test cases on the cubed-sphere by \citet{KNK15}.
+schemes have been found to give good accuracy \textcolor{red}{for
+Courant number less than one }on non-orthogonal meshes such as the
+cubed-sphere\textcolor{red}{{} }\textcolor{red}{\sout{\mbox{\citep{PL07}}}}\textcolor{red}{{}
+}with special treatment over cube edges. \citet{PL07} use the average
+of two one-sided schemes at cube edges \textcolor{red}{\sout{whereas
+\mbox{\citet{KNK15}} create ghost cells outside each cube panel boundary}}.
+\textcolor{red}{\citet{WNQ14} reported large error growth around
+cube edges for large Courant number using an alternating Strang splitting.}
+Without this special treatment, dimensionally split schemes do not
+account properly for mesh distortion and errors may occur. Accuracy
+between second and fourth order was found in practice on a range of
+test cases on the cubed-sphere by \citet{KNK15}.
 
@@
 is orography. A common solution has been to make the terrain following
-layers as smooth as possible, reducing non-orthogonality \citep[eg][]{SLF+02}
-or to use floating Lagrangian vertical co-ordinates \citep{Lin04}.
-Dimension splitting may account for some of the errors over orography
-reported by \citet{KUJ14} although they do not cite this as a reason
-for errors.
+layers as smooth as possible, reducing non-orthogonality \citep[eg][]{SLF+02}\textcolor{red}{{}
+away from the orography and increase the distortion around the orography
+}or to use floating Lagrangian vertical co-ordinates \citep{Lin04},
+\textcolor{red}{which} \textcolor{red}{incurs error from remapping
+between grids}. Dimension splitting may account for some of the errors
+over orography reported by \citet{KUJ14} although they do not cite
+this as a reason for errors.
 
@@
 solution of a matrix equation would be needed for every advected tracer.
-Two other disadvantages of implicit time-stepping are the large phase
-errors when long time-steps are used \citep[eg][]{DB12,LWW14} and
-the difficulty of achieving monotonicity. This is in contrast to semi-Lagrangian
-or FFSL schemes which maintain accuracy with long time-steps \citep{Pur76,PS84,LLM95}
-although monotonocity with long time-steps is still challenging \citep{Bott10}.
-Conservative semi-Lagrangian naturally extends to long time-steps
-but the conservative remapping is complicated and expensive, particularly
-on non-rectangular meshes and will not be investigated here. \citet{LUJ+14}
-described how the FFSL technique with a long time-step can be made
-equivalent to the conservative semi-Lagrangian.
+\textcolor{black}{Two} other disadvantages of implicit time-stepping
+are the large phase errors when long time-steps are used \citep[eg][]{DB12,LWW14}\textcolor{red}{\sout{,
+the difficulty of achieving monotonicity}} and \textcolor{red}{the
+inconvinient for parallelization due to linear solvers}. This is in
+contrast to semi-Lagrangian or FFSL schemes which maintain accuracy
+with long time-steps \citep{Pur76,PS84,LLM95} although monotonocity
+with long time-steps is still challenging \citep{Bott10}. Conservative
+semi-Lagrangian naturally extends to long time-steps but the conservative
+remapping is complicated and expensive, particularly on non-rectangular
+meshes and will not be investigated here. \citet{LUJ+14} described
+how the FFSL technique with a long time-step can be made equivalent
+to the conservative semi-Lagrangian.
 
@@
 On the cubed-sphere, special treatment is needed for the cube edges
-\citep[eg][]{LR96,KNK15}. Developing a transport scheme to the extent
-that it can be used on a multi-panel cubed-sphere with special treatment
-of cube edges is a considerable undertaking. Hence there is a need
-for more challenging advection test cases which are simpler to implement,
-without the need for spherical meshes. We therefore propose some modifications
-of existing test cases to use distorted meshes, or distorted co-ordinate
-systems, on a logically rectangular, two-dimensional plane.
+\textcolor{red}{for finite volume schemes}\citep[eg][]{LR96,KNK15}
+\textcolor{red}{while less effort is needed for element-based method
+like finite element method or discontinuous Garlerkin method (reference?)}.
+Developing a transport scheme to the extent that it can be used on
+a multi-panel cubed-sphere with special treatment of cube edges is
+a considerable undertaking. Hence there is a need for more challenging
+advection test cases which are simpler to implement, without the need
+for spherical meshes. We therefore propose some modifications of existing
+test cases to use distorted meshes, or distorted co-ordinate systems,
+on a logically rectangular, two-dimensional plane.
 
@@
 \citep{SLF+02}, examining sensitivity to time-step, resolution and
-mountain height, all on the maximally distorted basic terrain following
-mesh (section \ref{sub:overOrog}) and a modification of the deformational
-flow test case of \citet{LSPT12} for a periodic rectangular plane
-(section \ref{sub:deformFlow}). Some estimates are made of computational
-cost in section \ref{sub:cost} and final conclusions are drawn in
-section \ref{sec:concs}.
+mountain height, all on the \textcolor{red}{highly} distorted basic
+terrain following mesh (section \ref{sub:overOrog}) and a modification
+of the deformational flow test case of \citet{LSPT12} for a periodic
+rectangular plane (section \ref{sub:deformFlow}). Some estimates
+are made of computational cost in section \ref{sub:cost} and final
+conclusions are drawn in section \ref{sec:concs}.
 
@@
 
-%
+\selectlanguage{english}%
 In order to cope with long time-steps, we follow \citet{LLM95} and
@@
 
-%
+\selectlanguage{british}%
 Mesh distortions can be included in the advection equation with a
@@
 \noindent \begin{centering}
-\includegraphics{stencils}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{stencils.pdf}
 \par\end{centering}
@@
 \begin{centering}
-\includegraphics[width=0.49\textwidth,height=0.49\textwidth]{solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh}
+\includegraphics[width=0.49\textwidth,height=0.49\textwidth,bb = 0 0 200 100, draft, type=eps]{solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf}
 \par\end{centering}
@@
 \begin{figure}
-\includegraphics{sbrc1.png}
-
-\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=1\text{ s}$ giving a maximum Courant number of around 1.
-The contours show the tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
-The dashed contour shows the initial conditions.
-The min and max text in each plot gives the minimum and maximum errors.
-\label{fig:solidBody}}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{sbrc1.pdf}
+
+\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
+meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=1\text{ s}$
+giving a maximum Courant number of around 1. The contours show the
+tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
+The dashed contour shows the initial conditions. The min and max text
+in each plot gives the minimum and maximum errors. \label{fig:solidBody}}
 \end{figure}
@@
 \begin{figure}
-\includegraphics{sbrc10.png}
-
-\caption{Results of the solid body rotation test case on orthogonal and non-orthogonalmeshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$ giving a maximum Courant number of around 10.
-The contours show the tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
-The dashed contour shows the initial conditions.
-The min and max text in each plot gives the minimum and maximum errors.
-\label{fig:solidBody-dt10}}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{sbrc10.pdf}
+
+\caption{Results of the solid body rotation test case on orthogonal and non-orthogonal
+meshes every 100 seconds for meshes of $100\times100$ cells and $\Delta t=10\text{ s}$
+giving a maximum Courant number of around 10. The contours show the
+tracer value from 0.1 every 0.1 to 0.9 and the colours show the errors.
+The dashed contour shows the initial conditions. The min and max text
+in each plot gives the minimum and maximum errors. \label{fig:solidBody-dt10}}
 \end{figure}
@@
 \begin{centering}
-\includegraphics{sbr_dx}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{sbr_dx.pdf}
 \par\end{centering}
 
-\caption{Convergence with spatial resolution of the $\ell_{2}$ and $\ell_{\infty}$ errors for the solid body rotation test case at $t=500\ \text{seconds}$ ($5/6$ of one revolution) on orthogonal and non-orthogonal meshes
+\caption{Convergence with spatial resolution of the $\ell_{2}$ and $\ell_{\infty}$
+errors for the solid body rotation test case at $t=500\ \text{seconds}$
+($5/6$ of one revolution) on orthogonal and non-orthogonal meshes
 using multi-dimensional and dimensionally split schemes. \label{fig:solidBodyConverge} }
@@
 \begin{centering}
-\includegraphics{sbr_dt}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{sbr_dt.pdf}
 \par\end{centering}
 
-\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$errors for the solid body rotation test case at $t=500\ \text{seconds}$ ($5/6$ of one revolution) using orthogonal and non-orthogonal meshes of $100\times100$ cells using multi-dimensional and dimensionally split schemes.
-Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$ to $ $$c=10$ (for $\Delta t=10\ \text{s})$.
-\label{fig:solidBodyConverge-dt} }
+\caption{Convergence with temporal resolution of the $\ell_{2}$ and $\ell_{\infty}$
+errors for the solid body rotation test case at $t=500\ \text{seconds}$
+($5/6$ of one revolution) using orthogonal and non-orthogonal meshes
+of $100\times100$ cells using multi-dimensional and dimensionally
+split schemes. Maximum Courant numbers range from $c=0.25$ (for $\Delta t=0.25\ \text{s})$
+to $ $$c=10$ (for $\Delta t=10\ \text{s})$. \label{fig:solidBodyConverge-dt} }
 \end{figure}
@@
 split schemes. Horizontal-vertical split schemes are commonly used
-in this context \citep[eg.][]{DEE+12,WGZ+13}. We present results
-of the \citet{SLF+02} horizontal advection over orography test case
-for a range of resolutions and Courant numbers for the dimensionally
-split and multi-dimensional schemes.
+in this context \citep[eg.][]{DEE+12,WGZ+13} \textcolor{red}{since
+most weather/climate models use higher resolution in the vertical
+compared with horizontal}. We present results of the \citet{SLF+02}
+horizontal advection over orography test case for a range of resolutions
+and Courant numbers for the dimensionally split and multi-dimensional
+schemes.
 
@@
 \begin{figure}
-\includegraphics{overMountains.png}
-
-\caption{Tracer contours after 0, 5000 and 10000s for the horizontal advection over orography and errors in colour.
-Different plots use different time-steps.
-Spatial resolution is $\Delta x=1\ $km, $\Delta z=500\ $m.
-Maximum Courant numbers, horizontal Courant number, $c$, and maximum deformational Courant number, $c_{d}$ are also given.
-\label{fig:horizontalAdvectionWithDt}}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{overMountains.pdf}
+
+\caption{Tracer contours after 0, 5000 and 10000s for the horizontal advection
+over orography and errors in colour. Different plots use different
+time-steps. Spatial resolution is $\Delta x=1\ $km, $\Delta z=500\ $m.
+Maximum Courant numbers, horizontal Courant number, $c$, and maximum
+deformational Courant number, $c_{d}$ are also given. \label{fig:horizontalAdvectionWithDt}}
 \end{figure}
@@
 \noindent \begin{centering}
-\includegraphics{overMountains_dx}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{overMountains_dx.pdf}
 \par\end{centering}
 
-\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metricsfor horizontal advection over a 3000m and a 6000m mountains after 10,000s for the dimensionally split (PPM) scheme and multi-dimensional MOL scheme, all using a horizontal Courant number of 0.25.
-Lines also show theoretical first- and second-order convergence are also shown.
+\caption{Convergence of the $\ell_{2}$ and $\ell_{\infty}$ error metrics
+for horizontal advection over a 3000m and a 6000m mountains after
+10,000s for the dimensionally split (PPM) scheme and multi-dimensional
+MOL scheme, all using a horizontal Courant number of 0.25. Lines also
+show theoretical first- and second-order convergence are also shown.
 \label{fig:horizontalAdvectConverge}}
@@
 \begin{figure}
-\includegraphics{deform_init}
-
-\caption{The non-orthognal mesh/co-ordinate system for the deformational flow with $120\times60$ cells and the initial tracer conditions.\label{fig:deformGrid}}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{deform_init.pdf}
+
+\caption{The non-orthognal mesh/co-ordinate system for the deformational flow
+with $120\times60$ cells and the initial tracer conditions.\label{fig:deformGrid}}
 \end{figure}
@@
 \begin{figure}
-\includegraphics{deform.png}
-
-\caption{Results of the deformational flow test case after 1, 2, 3, 4 and 5 time units on the non-orthogonal meshes of 480$\times$240 cells using the multi-dimensional and split schemes.
-\label{fig:deformResults}}
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{deform.pdf}
+
+\caption{Results of the deformational flow test case after 1, 2, 3, 4 and 5
+time units on the non-orthogonal meshes of 480$\times$240 cells using
+the multi-dimensional and split schemes. \label{fig:deformResults}}
 \end{figure}
@@
 \begin{figure}
-\includegraphics{deform_dx}
-
-\caption{Convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$ errors for the deformational flow on orthogonal and non-orthogonal meshes using multi-dimensional and dimensionally split schemes.
-\label{fig:deformationalConverge} }
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{deform_dx.pdf}
+
+\caption{Convergence with resolution of the $\ell_{2}$ and $\ell_{\infty}$
+errors for the deformational flow on orthogonal and non-orthogonal
+meshes using multi-dimensional and dimensionally split schemes. \label{fig:deformationalConverge} }
 \end{figure}
@@
 \begin{figure}
-\includegraphics{deform_dt}
-
-\caption{Variation of $\ell_{2}$ and $\ell_{\infty}$ errors with time-step for the deformational flow on orthogonal and non-orthogonal meshes of $120\times60$ cells using multi-dimensional and dimensionally split schemes.
-Different time-steps give maximum Courant numbers of 0.25, 0.5, 1, 2, 5 and 10.
-\label{fig:deformational_dt} }
+\includegraphics[bb = 0 0 200 100, draft, type=eps]{deform_dt.pdf}
+
+\caption{Variation of $\ell_{2}$ and $\ell_{\infty}$ errors with time-step
+for the deformational flow on orthogonal and non-orthogonal meshes
+of $120\times60$ cells using multi-dimensional and dimensionally
+split schemes. Different time-steps give maximum Courant numbers of
+0.25, 0.5, 1, 2, 5 and 10. \label{fig:deformational_dt} }
 \end{figure}
@@
 
-The dimensionally split scheme is extremely accurate on orthogonal
-meshes and retains accuracy when long time-steps are used. However
-on distorted meshes, particularly at changes in direction such as
-those that appear at cube sphere edges or over orography, the split
-scheme loses some accuracy. In contrast, the multi-dimensional scheme
-is almost entirely insensitive to mesh distortion and asymptotes to
-second-order convergence at high resolution. As is expected for implicit
-time-stepping, phase errors occur when using long time-steps but the
-spatially well resolved features are advected at the correct speed
-and the multi-dimensional scheme is always stable.
+The dimensionally split scheme is \textcolor{red}{\sout{extremely}}
+accurate on orthogonal meshes and retains accuracy when long time-steps
+are used. However on distorted meshes, particularly at changes in
+direction such as those that appear at cube sphere edges or over orography,
+the split scheme loses some accuracy. In contrast, the multi-dimensional
+scheme is almost entirely insensitive to mesh distortion and asymptotes
+to second-order convergence at high resolution. As is expected for
+implicit time-stepping, phase errors occur when using long time-steps
+but the spatially well resolved features are advected at the correct
+speed and the multi-dimensional scheme is always stable.
 
@@
 split schemes are used on cubed-sphere meshes, they do usually have
-special interpolations at cube edges \citep[eg][]{LR96}. However
+special interpolations at cube edges \citet[eg][]{PL07}. However
 orography is everywhere and so special treatment over steep orography
@@
 
-\section*{Acknowledgements}
-
-Thanks to Nigel Wood of the UK Met Office for advice and proof reading. Weller acknowledges support from NERC grant NE/K006797/1.
-
 \bibliographystyle{abbrvnat}

### 2017-05-18 revision/reply_to_reviewer.tex
More responses to reviewer comments

@@
-\documentclass[12pt,a4paper]{article}
-\usepackage{mathptmx}
-\usepackage[T1]{fontenc}
-\usepackage[latin9]{inputenc}
-\setcounter{secnumdepth}{3}
-\setcounter{tocdepth}{3}
+\documentclass[12pt]{article}
+\usepackage[margin=2cm]{geometry}
 \usepackage{url}
@@
 \usepackage{graphicx}
-\usepackage{esint}
 \usepackage[authoryear]{natbib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
-\usepackage { fancybox}
-\usepackage[export]{adjustbox}
-
 \usepackage{todonotes}
 
-\usepackage{amssymb}
+\begin{document}
 
-\usepackage{lineno}
-%\linenumbers
+\title{Responses to reviewer comments on artile QJ-17-0017\\
+Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme for Atmospheric Transport}
+\author{Yumeng Chen \and Hilary Weller \and Stephen Pring \and and James Shaw}
+\maketitle
 
-\makeatother
+\section*{Responses to reviewer 1}
 
-% \usepackage{babel}
-\begin{document}
 \begin{description}
+
+\item [Comment]:
+I like the subject matter, and we are certainly in need of new ways of gaining efficiency for transport algorithms and for full dynamics for that matter.
+I think dimensionally split semi-Lagrangian schemes could be very competitive, but I have concerns about their accuracy on the cubed-sphere (as do the authors).
+I like the ingenuity of the test cases as well.
+I have three major concerns regarding the manuscript though.
+I have annotated these and other concerns directly to the PDF document, which is attached to this review.
+I think the content is important and needed and that the manuscript should be published, but subject to what I think amount to major revisions.
+
+\item [Response]:
+Many thanks for your insightful comments. We hope that we have now addressed these concerns.
+
+\item [Comment 1]: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: \url{http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1}.
+Additionally, Katta et al was cited as a dimensionally split scheme, but it is not.
+The reconstruction is performed in dimensional sweeps, but the scheme itself is multi-dimensional.
+
+\item [Response]: 
+Thank-you for pointing out the \cite{GNQ14} paper.
+This is now cited in the introduction.
+Transport errors are dominated by errors in the reconstruction so if the reconstruction is performed in dimensional sweeps, then I would describe the \cite{KNK15} scheme as using dimension splitting rather than being fully multi-dimensional.
+We have put a caveat in pointing out that the reconstruction is in dimensional sweeps.
+
+\item [Comment 2]:
+It is stated that results in this 2-D framework will generalize well to the cubed-sphere, but I believe that in order to truly make that statement, one would have to implement a given scheme on both meshes and then compare.
+My experience on the cubed-sphere mesh is that it.
+
+\item [Response]:
+Agreed.
+Our test cases on a plane will highlight errors due to non-orthogonality and due to discontinuities at cube edges but cannot help to describe errors at cube corners.
+A sentence has been added to the introduction:
+\begin{quote}
+These test cases will mimic the problems encountered at cubed-sphere edges but will not give any information about errors that might occur at cubed-sphere corners. 
+\end{quote}
+A similar caveat has been inserted in the second paragraph of the abstract, the first paragraph of section 3 and in the third paragraph of the conclusions.
+
+\item [Comment 3]: Most importantly, to say the newly proposed dimensionally split scheme is accurate compared to multi-dimensional schemes, I believe a better comparison point is needed.
+Users select the scheme that gives the best results with the fewest resources.
+I believe the authors need to include a time-explicit semi-discrete (method of lines) multi-dimensional scheme with cubic interpolants using smaller time steps to serve as a comparison point for this scheme.
+I know the focus is on large time steps, but it's necessary to consider if one can, in fact, get more accuracy for a given amount of work by simply using the status quo with smaller time steps.
+
+\item [Response]: 
+We now compare the dimensionally split scheme with the multi-dimensional scheme when the multi-dimensional scheme is run with explicit and implicit time-stepping.
+New results are described in section 3.
+The errors of the multi-dimensional scheme (which is now described in full by \cite{SWMD17}) are insensitive to the choice of time-stepping type (implicit or explicit) for Courant numbers smaller than two.
+These comparisons have also revealed that the implicit solve is a small part of the total cost of the advection.
+This is likely because the solvers in OpenFOAM are well optimised but there are cache misses associated with the explicit part of the advection.
+The multi-dimensional scheme uses cubic interpolants but is not overall a third-order accurate.
+The multi-dimensional scheme is designed for arbitrary meshes.
+In order to achieve third-order accuracy on arbitrary meshes either multi-moments or Gauss points within each cell are needed.
+This would increase the complexity and cost significantly.
+The multi-dimensional scheme run explicitly has similar cost and complexity to the dimensionally split scheme.
+
+\end{description}
+
+\section*{Responses to reviewer 1 pdf annotations}
+
+\begin{enumerate}
 \item [Page 1, Line 41]: Might want to say ``distorted, non-orthogonal meshes'' to make it's clear that you are indeed directly testing non-orthogonality.
@@
 
-\item [E-mail, Reviewer 1]: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: <http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>.
-Additionally, Katta et al was cited as a dimensionally split scheme, but it is not.
-The reconstruction is performed in dimensional sweeps, but the scheme itself is multi-dimensional.
+\end{enumerate}
 
-\item [Yumeng]: We removed the citation and relevant description.
+\section*{Responses to reviewer 2 (Lucas Harris)}
 
-\item [E-mail, Reviewer 1]: Most importantly, to say the newly proposed dimensionally split scheme is accurate compared to multi-dimensional schemes, I believe a better comparison point is needed.
-Users select the scheme that gives the best results with the fewest resources.
-I believe the authors need to include a time-explicit semi-discrete (method of lines) multi-dimensional scheme with cubic interpolants using smaller time steps to serve as a comparison point for this scheme.
-I know the focus is on large time steps, but it's necessary to consider if one can, in fact, get more accuracy for a given amount of work by simply using the status quo with smaller time steps.
-
-\item [Yumeng]: The method of lines scheme is chosen to achieve similar amount of work as dimensionally split scheme.
-PPM on uniform grid only need a 4-point stencil for each cell edge in 1-D, while cubic interpolants will lead to a much larger stencil in 2-D.
-The results should be similar for smaller time steps as we already have tests for Courant number less than one.
-On the other hand, dimensionally splitting scheme only uses first-order Euler, which is less accurate than Crank-Nicolson method for method of lines scheme.
-
-\item [E-mail, Reviewer 1]: It is stated that results in this 2-D framework will generalize well to the cubed-sphere, but I believe that in order to truly make that statement, one would have to implement a given scheme on both meshes and then compare.
-My experience on the cubed-sphere mesh is that it.
-
-\item [Yumeng]: The non-orthogonal mesh has similar feature as edges at cubed-sphere: discontinuity in Jacobian.
-The results should be able to show a glimpse of the property of COSMIC splitting, especially for long time steps.
-
-\item [E-mail, Reviewer 2]: The first questions that come to my mind when reading this paper are, ``why another advection scheme?'' and ``why another advection test case?'' given the multitude of existing schemes and tests.
+\begin{enumerate}
+\item [Major comment 1]: The first questions that come to my mind when reading this paper are, ``why another advection scheme?'' and ``why another advection test case?'' given the multitude of existing schemes and tests.
 The demonstration of the good accuracy of standard dimensionally-split FFSL schemes is certainly useful, especially considering the well-publicized shortcomings of such methods.
-
-\item [Yumeng]:
-
-\item [E-mail, Reviewer 2]:  However, I am concerned that the comparison is somewhat limited by the fact that the multi-dimensional scheme is of lower-order than the dimensionally-split scheme.
+However, I am concerned that the comparison is somewhat limited by the fact that the multi-dimensional scheme is of lower-order than the dimensionally-split scheme.
 How can the lower order of accuracy of the multi-dimensional scheme justified?
 
-\item [Yumeng]:
+\item [Response]: 
+Our initial motivation was to demonstrate the large errors associated with dimensionally split schemes on non-orthogonal meshes with discontinuities in co-ordinate transforms which we thought had not previously been demonstrated in the literature.
+We could not find these large errors despite comparing results on orthogonal and non-orthogonal meshes.
 
-\item [E-mail, Reviewer 2]: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
+Accuracy higher than second order is not straightforward for a multi-dimensional scheme suitable for arbitrary meshes.
+The complexity and cost of the multi-dimensional scheme is similar to that of the dimensionally split scheme.
+However, in fact, neither scheme has accuracy higher than second order although they both have aspects that are third-order accurate.
+The dimensionally split scheme is third or forth order accurate in one dimension but the COSMIC splitting reduces the order of accuracy.
+We thought that discontinuities in the coordinate system would reduce the order of accuracy to first but it appears to be second order.
+The multi-dimensional scheme uses cubic interpolations but since we use the cell centre as the Gauss point for each cell, the multi-dimensional scheme is limited to second order accuracy.
+We have added this text to the end of the first paragraph of section 2:
+\begin{quote}
+The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the order to two.
+The multi-dimensional scheme uses cubic interpolations but uses cell centoids as Gauss points so the order of accuracy is limited to two.
+When run explicitly, the cost of both schemes is similar (see section 3.4).
+These schemes are therefore considered suitable for comparison.
+\end{quote}
+
+\item [Major comment 2]: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
 The effect of this is most visible in Figure 4, where the more accurate dimensionally-split scheme has more grid-scale noise than the lower-order multi-dimensional scheme, likely because of the larger implicit diffusion in the lower-order scheme.
@@
 
-\item [Yumeng]: Yes, monotonicity preservation is important for modern solvers.
-The shape change is more likely to be phase error from time stepping scheme.
-However, it is unfair to compare scheme with limiter for FFSL with multi-dimensional scheme without proper limiter.
+\item [Response]: We agree that advection schemes without monotonicity preservation are of limited use.
+While some of the dimensionally split scheme results have oscillations, they are still more accurate than the smooth results from the multi-dimensional scheme.
+Therefore including a good monotone limiter shouldn't change the conclusions of this paper.
+I am afraid that we no longer have the resources to implement a good monotone limiter for the dimensionally split scheme.
+We feel that these points are already covered sufficiently in section 3.1:
+\begin{quote}
+The dimensionally split scheme outperforms the multi-dimensional scheme on both meshes due to the higher-order accuracy of the split scheme.
+The dimensionally split scheme introduces a small error at 300 seconds where the tracer goes through the change in direction of the mesh which would be ameliorated if we were using monotonicity constraints.
+The second-order, multi-dimensional scheme shows phase lag but errors are almost entirely insensitive the mesh distortions.
+\end{quote}
 
-\item [E-mail, Reviewer 2]: Given that a fundamentally new scheme is not presented here, a better title might be something like, ``Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps''
+\item [Minor Comment 1]: Given that a fundamentally new scheme is not presented here, a better title might be something like, ``Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps''
 
-\item [Yumeng]:
+\item [Response]: Agreed, thanks.
 
-\item [E-mail, Reviewer 2]: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
+\item [Minor Comment 2]: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
 LR96 is designed for cartesian and lat-lon grids.
@@
 
-\item [Yumeng]: One wrong reference of this kind at last page has changed to Lin (2007).
+\item [Response]: Thanks, changed.
 
-\item [E-mail, Reviewer 2]: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
+\item [Minor Comment 3]: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
 
-\item [Yumeng]:  Yes.(?)
+\item [Response]: Yes. This is now clarified in section 3.2:
+\begin{quote}
+The width and height of the orography is kept constant.
+\end{quote}
 
-\item [E-mail, Reviewer 2]: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
-\end{description}
+\item [Minor Comment 4]: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
+
+\item [Response]: Thank-you. Following a request from reviewer 1, we also now include a comparison of wall clock time for the multi-dimensional scheme run implicitly and explicitly in serial. 
+\end{enumerate}
+
 \bibliographystyle{abbrvnat}

### 2017-05-19 revision/reply_to_reviewer.tex
Hilary's first draft of replies to reviewer comments (following Yumeng's first draft). Next need to update the figures and comment on the explicit results

@@
 \usepackage[margin=2cm]{geometry}
-\usepackage{url}
+\usepackage{url,times}
 \usepackage{amsmath}
@@
 
-\item [Response]:
-Many thanks for your insightful comments. We hope that we have now addressed these concerns.
-
-\item [Comment 1]: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: \url{http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1}.
-Additionally, Katta et al was cited as a dimensionally split scheme, but it is not.
+\item [Comment 1]: This is the most minor, but there is a study for semi-Lagrangian dimensionally split transport on the cubed-sphere that needs to be considered and discussed: \cite{GNQ14}.
+Additionally, \cite{KNK15} was cited as a dimensionally split scheme, but it is not.
 The reconstruction is performed in dimensional sweeps, but the scheme itself is multi-dimensional.
@@
 \item [Response]: 
-Thank-you for pointing out the \cite{GNQ14} paper.
-This is now cited in the introduction.
-Transport errors are dominated by errors in the reconstruction so if the reconstruction is performed in dimensional sweeps, then I would describe the \cite{KNK15} scheme as using dimension splitting rather than being fully multi-dimensional.
-We have put a caveat in pointing out that the reconstruction is in dimensional sweeps.
+Thank-you for pointing out the \cite{GNQ14} paper. See response to pdf annotations below.
 
@@
 Our test cases on a plane will highlight errors due to non-orthogonality and due to discontinuities at cube edges but cannot help to describe errors at cube corners.
-A sentence has been added to the introduction:
+Some sentences have been added to the introduction:
+\begin{quote}
+We assume that a large fraction of the numerical errors associated with terrain following co-ordinates and the cubed-sphere grid are due to the mesh non-orthoganlity and distortion.
+However there will also be errors due to cubed-sphere corners and inaccurate representation of spherical geometry.
+These are not addressed here.
+\end{quote}
+and
 \begin{quote}
@@
 These comparisons have also revealed that the implicit solve is a small part of the total cost of the advection.
-This is likely because the solvers in OpenFOAM are well optimised but there are cache misses associated with the explicit part of the advection.
+This is likely because the linear equation solvers in OpenFOAM are well optimised, our test cases are run in serial and there are cache misses associated with the explicit part of the advection which uses extended stencils.
 The multi-dimensional scheme uses cubic interpolants but is not overall a third-order accurate.
@@
 
+Page and line numbers refer to the original annotated document.
+
 \begin{enumerate}
@@
 
-\item [Yumeng]: Changed as suggested.
+\item [Response]: Changed as suggested.
 
@@
 
-\item [Yumeng]: Quasi-uniform resolution indeed help better exploit modern architectures.
+\item [Response]: Quasi-uniform resolution indeed help better exploit modern architectures.
 The lat-lon grid leads to convergence of resolution toward poles and restricts the time stepping.
@@
 The modern computers use increasing number of cores and the efficiency is limited by the scalability.
-Implicit solvers and semi-Lagrangian advection result in massive data communications \citep{ST12}, which limits the scalability around poles.
+Implicit solvers and semi-Lagrangian advection result in massive data communications on lat-lon grids \citep{ST12}, which limits the scalability around poles.
+
+I know what you mean but, in reality, the modern computer architectures really are driving the need for quasi-uniform meshes and this is well documented in the references that we cite.
+We have changed the sentence to:
+\begin{quote}
+Weather and climate models are being developed on quasi-uniform meshes in order to equidistribute and reduce the storage needed for the domain of dependence of the model degrees of freedom and hence better exploit modern computers \cite[eg][]{WTC12,LUJ+14,ST12,SG11,KNK15}.
+\end{quote}
 
@@
 
-\item [Yumeng]: The error on the sphere indeed is larger than on the plane for dimensionally splitting scheme, even on the lat-lon grid.
-On cubed sphere, the mesh around corner of each panel has abrupt changes and is not differentiable.
-\cite{TN17} implemented shallow water equations on generalized curvilinear coordinate on the plane as a test before the implementation on the sphere.
-The comparison between plane and non-orthogonal mesh with abrupt change is to show the difference between orthogonal grid and non-orthogonal grid, which has similar feature as cubed sphere.
-The test certainly cannot be equivalent to the comparison between orthogonal lat-lon grid and cubed-sphere grid.
+\item [Response]: Agreed. We have added the following to the end of this paragraph:
+\begin{quote}
+We assume that a large fraction of the numerical errors associated with terrain following co-ordinates and the cubed-sphere grid are due to the mesh non-orthoganlity and distortion.
+However there will also be errors due to cubed-sphere corners and inaccurate representation of spherical geometry.
+There are not addressed here.
+\end{quote}
 
@@
 
-\item [Yumeng]:  The remapping here refers to integrating the reconstruction function to get the mass fluxes along the trajectory of each cell face.
-The cumulative mass technique \citep{CW84,LLM95} avoids the mass calculation along the trajectory and we just need to get the integral at departure cell with simple add/subtraction from cumulative mass.
-We did not remap between different grids like Yin-Yang.
+\item [Response]: I was thinking about schemes like SLICE and CSLAM here but on reflection it is not best to discuss these schemes at this point so this comment has been removed.
 
@@
 
-\item [Yumeng]: changed as suggested
+\item [Response]: changed as suggested
 
-\item [Page 2, Line 45]: The scheme in this paper isn't really dimensionally split, only the *reconstruction* is.
-Katta et al use dimensionally split sweeps to create a genuinely multi-dimensional interpolant (minus a few of the cross terms), and then they evolve in a genuinely multi-dimensional way.
+\item [Page 2, Line 45]: These guys \cite[]{PL07} appear to be using a sort of predictor-corrector approach where one predicts in the $x$-direction to $\Delta t/2$, then corrects with $y$ over the time-step.
+Then they swap the dimensions and average the final result.
+Their errors are certainly more uniform over time, but they are larger (using $90\times 90$ DOFs per panel) than \cite{GNQ14} (who used $80\times 80$ DOFs per panel).
+
+\item [Page 2, Line 46]: The scheme \cite[]{KNK15} in this paper isn't really dimensionally split, only the *reconstruction* is.
+\cite{KNK15} use dimensionally split sweeps to create a genuinely multi-dimensional interpolant (minus a few of the cross terms), and then they evolve in a genuinely multi-dimensional way.
 This is an important distinction because (1) they do not incur time error in the same way as truly split schemes and (2) they include a diagonal sweep for some of the cross terms, whereas truly split schemes generally do not.
 
-\item [Yumeng]: Agreed. We deleted the reference and the description herein.
+\item [Response]: These sentences have been changed to:
+\begin{quote}
+Dimensionally split schemes have been found to give good accuracy on non-orthogonal meshes such as the cubed-sphere if special treatment is applied over cube edges to improve accuracy: \cite{PL07} use the average of two one-sided schemes at cube edges.
+\cite{KNK15} create ghost cells outside each cube panel boundary and achieve accuracy between second and fourth order.
+\cite{KNK15} use a multi-dimensional transport scheme with dimensionally split reconstructions.
+\cite[eg]{GNQ14} report error growth around cube edges using Strang splitting.
+\end{quote}
 
@@
 
-<http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1>
+\url{http://journals.ametsoc.org/doi/abs/10.1175/MWR-D-13-00048.1}
 
@@
 
-\item [Yumeng]: 
-On this paper, on page 460: 
-``It is difficult to design a numerical scheme that preserves a constant field in the dimensional splitting framework.
+\item [Response]: Thank-you for pointing out this reference. Included as described above.
+
+%On this paper, on page 460: 
+%``It is difficult to design a numerical scheme that preserves a constant field in the dimensional splitting framework.
 It is our ongoing work to design a nonsplitting SLDG scheme that preserves the constant field when the velocity field is nondivergent.''
-Is this because of the filter or the splitting? How is the performance of the scheme under 1-D divergnet flow, because the 2-D non-divergent field could lead to 1-D divergent flow in each dimension.
+%Is this because of the filter or the splitting? How is the performance of the scheme under 1-D divergnet flow, because the 2-D non-divergent field could lead to 1-D divergent flow in each dimension.
 
-On page 464, ``It is clear that the second order of convergence is observed, which comes from the splitting error.''
-I believe in solid body rotation test case on Cartesian grid, the splitting error is zero in COSMIC splitting if the splitting scheme is 2nd order in time.
+%On page 464, ``It is clear that the second order of convergence is observed, which comes from the splitting error.''
+%I believe in solid body rotation test case on Cartesian grid, the splitting error is zero in COSMIC splitting if the splitting scheme is 2nd order in time.
 So I think the splitting scheme is not comparable to the COSMIC splitting.
 
-On page 469, ``This is an indication that such error comes from the dimensional splitting, and there exists a certain symmetry property for different dimensional-splitting orderings.
+%On page 469, ``This is an indication that such error comes from the dimensional splitting, and there exists a certain symmetry property for different dimensional-splitting orderings.
 Such symmetry property, together with the symmetry of the cosine bell profile, may contribute to the dropping of the $l_\infty$ error after the cosine bell passed the corner.''
-If this is the reason for the large error, COSMIC splitting may give a better results as COSMIC splitting get a symmetric results by averaging two splitting scheme.
+%If this is the reason for the large error, COSMIC splitting may give a better results as COSMIC splitting get a symmetric results by averaging two splitting scheme.
 
-\item [Page 2, Line 55]: It's important to note that all Schar et al did was compress all of the errors to the lowest few levels, where the metric Jacobian determinants vary even more.
+\item [Page 2, Line 55]: It's important to note that all \cite{SLF+02}  did was compress all of the errors to the lowest few levels, where the metric Jacobian determinants vary even more.
 
-\item [Yumeng]: We add words here as suggested.
+\item [Response] Agreed. Added the phrase ``away from the ground''.
 
@@
 
-\item [Yumeng]: We add words here as suggested.
+\item [Response]: There are of course errors from remapping but we are unaware that these errors are worse where the grid is highly distorted over orography. Therefore we have left the text as it is.
 
-\item [Page 3, Line 4]: Ullrich \& Norman, 2012 QJRMS performed a multi-dimensional FFSL method with hyperdiffusion out to a CFL of roughly 2.5
+\item [Page 3, Line 4]: \cite{UN14} performed a multi-dimensional FFSL method with hyperdiffusion out to a CFL of roughly 2.5
 
-\item [Yumeng]: Not treated.
+\item [Response]. Thank-you. We have added:
+\begin{quote}
+a notable exception being \cite{UN14}, whose FFSL semi-Lagrangian method is stable for Courant numbers up to around 2.5.
+\end{quote}
 
@@
 
-\item [Yumeng]: The cost is proportional to the CFL because the flux-form scheme for higher CFL has to integrate larger areas as the trajectory of departure cell is longer.
+\item [Response]: Agreed. Added ``integration cost''
 
@@
 
-\item [Yumeng]: We added statements as suggested.
+\item [Response]: Added ``the parallelization cost of linear solvers''.
 
@@
 
-\item [Yumeng]: We deleted the relevant statement for monotonicity of method of lines schemes.
+\item [Response]: I don't think so because monotonic schemes are non-linear. We have left this as ``the \emph{difficulty} of achieving monotonicity'' so as not to exclude the possibility.
 
@@
 
-\item [Yumeng]: We added statements as suggested.
+\item [Response]: Actually we wanted to talk about testing advection schemes here rather than properties of advection schemes so the sentence has been replaced by:
+\begin{quote}
+The theoretical properties of dimensionally split advection schemes are often tested on uniform, orthogonal meshes \cite[eg][]{LLM96} which is inadequate for a scheme that will eventually be used on a cubed-sphere.
+\end{quote}
 
@@
 
-\item [Yumeng]: We changed it to highly distorted grids.
+\item [Response]: Replaced by:
+\begin{quote}
+without smoothing of terrain following layers
+\end{quote}
 
@@
 
-\item [Yumeng]: Yes, higher order scheme for departure point would be expensive. On the other hand, the original flux-form semi-Lagrangian scheme by \citep{LR96} also just uses first order Euler.
+\item [Response]: We have added:
+\begin{quote}
+The next step would be to use a second-order predictor-corrector departure point calculation \cite[eg][]{MDW+10}.
+\end{quote}
 
@@
 
-\item [Yumeng]: Using the same scheme both for inner and outer operator can avoid stability problem.
-Such treatment makes the use of PPM in both inner and outer operator easier because finite volume scheme requires flux at cell boundaries instead of flux at cell center.
+\item [Response]: COSMIC splitting is defined as using both advective and conservative advection operators in order to maintain stability, conservation, preservation of constants and second-order accuracy.
+This is already described so we haven't changed the text here.
 
@@
 
-\item [Yumeng]: Leave it to Hilary.
+\item [Response]: Removed the hexagonal mesh.
 
@@
 
-\item [Yumeng]: Leave it to Hilary.
+\item [Response]: Yes, done.
 
@@
 
-\item [Yumeng]: When time step is 1s, the Courant number is less than one and most explicit scheme without semi-Lagrangian extension should be able to handle such Courant number.
-Phase error is only obvious for Courant number larger than one.
-In the case of $\Delta t = 1$ s, a fully explicit, multi-dimensional scheme for smaller time step will lead to similar results.
+\item [Response]: See response to comment 3 above.
 
@@
 
-\item [Yumeng]: Yes, a better time stepping scheme may improve phase error, but it may also introduce a more complicated scheme.
-The reason we chose CN is due to the consideration of complex of implementation.
-The implementation difficulty of CN is comparable to COSMIC splitting with PPM.
+\item [Response]: Good idea.
+Simpler than improving the spatial accuracy of the multi-dimensional scheme.
+However improving the spatial accuracy is a higher priority since for Courant numbers less than one, the multi-dimensional scheme errors are insensitive to time-step.
+No change made to manuscript.
 
@@
 
-\item [Yumeng]: As stated, the baseline is based on the complex of implementation.
+\item [Response]: We agree that the baseline could be better. However the two schemes do have similar complexity and cost. The final sentence has been changed to:
+\begin{quote}
+In contrast, the multi-dimensional scheme is insensitive to the orthogonality, converges with second-order \textcolor{blue}{as expected} and suffers from phase errors at large Courant numbers \textcolor{blue}{due to the use of Crank-Nicolson implicit time-stepping}.
+\end{quote}
 
@@
 
-\item [Yumeng]: We use such statement because the BTF has the maximum distortion compared with other terrain-following coordinate or other treatment of orography.
+\item [Response]: We have changed this to:
+\begin{quote}
+maximises the non-orthogonality that can be introduced by terrain following coordinates.
+\end{quote}
+Smoother terrain following coordinates are more orthogonal but less uniform near the ground.
 
@@
 
-\item [Yumeng]: We add explanation at the start of this section as suggested.
+\item [Response]: We add explanation at the start of this section as suggested.
 
@@
 
-\item [Yumeng]: We remove the word extremely.
+\item [Response]: Agreed. We removed the words ``very'' and "extremely'' in the following locations:
+\begin{itemize}
+\item One instance of  ``very'' in the abstract.
+\item One instance of  ``very'' in the conclusions.
+\item One instance of  ``extremely'' in the conclusions.
+\end{itemize}
+Also caveats have been added to the first paragraph of the conclusions explaining the limitations of the comparison.
 
@@
 However, in fact, neither scheme has accuracy higher than second order although they both have aspects that are third-order accurate.
-The dimensionally split scheme is third or forth order accurate in one dimension but the COSMIC splitting reduces the order of accuracy.
+The dimensionally split scheme is third or forth order accurate in one dimension but the COSMIC splitting reduces the temporal order of accuracy even on an orthogonal grid.
 We thought that discontinuities in the coordinate system would reduce the order of accuracy to first but it appears to be second order.
 The multi-dimensional scheme uses cubic interpolations but since we use the cell centre as the Gauss point for each cell, the multi-dimensional scheme is limited to second order accuracy.
-We have added this text to the end of the first paragraph of section 2:
+We have added this text to the end of the first paragraph of section 2 and similar wording to the beginning of the conclusions:
 \begin{quote}
-The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the order to two.
+The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the temporal order to two.
 The multi-dimensional scheme uses cubic interpolations but uses cell centoids as Gauss points so the order of accuracy is limited to two.

### 2017-05-19 revision/getGraphics.sh

@@ -1,0 +1,110 @@
+#!/bin/bash -e
+
+# Root directory where graphics files can be found
+ROOTDIR=/home/hilary/OpenFOAM/*/run
+
+#List of files to convert to png
+pngFiles=(
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/0/T.pdf
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/1/T.pdf
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/2/T.pdf
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/3/T.pdf
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/4/T.pdf
+deformingAdvectionCyclic/nonOrthogW_c1/480x240/5/T.pdf
+)
+
+# List of files to copy and rename
+cpFiles=(
+deformingAdvectionCyclic/nonOrthogW_c1/120x60/0/TUmesh.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt1000CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt100CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt200CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt25CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt500CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt50CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography_coarse/save/cubicUpwindCPCFit_dt50CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography_fine/save/cubicUpwindCPCFit_dt12.5CN_up/10000/Tall.pdf
+vSlice/horizontalAdvection/orography_vfine/save/cubicUpwindCPCFit_dt6.25CN_up/10000/Tall.pdf
+
+vSlice/horizontalAdvection/highOrography/save/cubicUpwindCPCFit_dt1000N_up/10000/TallTall.pdf
+vSlice/horizontalAdvection/highOrography/save/cubicUpwindCPCFit_dt100N_up/10000/TallTall.pdf
[…]

### 2017-05-19 revision/makeFigs.sh

@@ -1,0 +1,21 @@
+#!/bin/bash -e
+cd makeFigs
+
+for file in stencils sbrc1 sbrc10 sbr_dx sbr_dt overMountains overMountains_dx deform_init deform deform_dx deform_dt; do
+    lyx --export pdflatex -f $file
+    pdflatex $file
+    pdfcrop $file.pdf ../figures/$file.pdf
+    rm $file.pdf
+    evince ../figures/$file.pdf &
+done
+
+cd ..
+
+cp -u HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf .
+
+eps2png2 deform.pdf
+eps2png2 sbrc10.pdf
+eps2png2 sbrc1.pdf
+
+zip figures.zip stencil.pdf solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf sbrc1.png sbrc10.png sbr_dx.pdf sbr_dt.pdf overMountains.png overMountains_dx.pdf deform_init.pdf deform.png deform_dx.pdf deform_dt.pdf
+

### 2017-05-25 revision/reply_to_reviewer.tex
Hilary's first draft of revisions and replies to reviewer comments

@@
 \usepackage{todonotes}
+\usepackage[normalem]{ulem}
+\usepackage{color}
 
@@
 \begin{quote}
-These test cases will mimic the problems encountered at cubed-sphere edges but will not give any information about errors that might occur at cubed-sphere corners. 
+These test cases will mimic the problems encountered at cubed-sphere edges but will not give any information about additional errors that might occur at cubed-sphere corners. 
 \end{quote}
@@
 \item [Response]:
-We now compare the dimensionally split scheme with the multi-dimensional scheme when the multi-dimensional scheme is run with explicit and implicit time-stepping.
+An important result from this paper is that the dimensionally split scheme loses very little accuracy when grid distortions are introduced.
+This result does not rely on the accuracy of the multi-dimensional scheme.
+Text has been added throughout the paper to clarify this point.
+The text in blue has been added:
+\begin{itemize}
+\item In the abstract:
+\begin{quote}
+Mesh distortions are expected to accentuate the errors associated with dimension splitting, however, \textcolor{blue}{the accuracy of} the dimensionally split scheme \sout{is very accurate on orthogonal meshes and accuracy} decreases only a little in the presence of mesh distortions.
+\end{quote}
+\item In the introduction:
+\begin{quote}
+we show the effect of dimension splitting errors using a FFSL dimensionally split scheme on a number of test-cases which use distorted \textcolor{blue}{and un-distorted} meshes.
+\end{quote}
+\item In the conclusions
+\begin{quote}
+The dimensionally split scheme is most accurate on orthogonal meshes but only loses a little accuracy on highly distorted meshes, despite a first-order departure point calculation.
+\end{quote}
+\end{itemize}
+
+We also now compare the dimensionally split scheme with the multi-dimensional scheme when the multi-dimensional scheme is run with explicit and implicit time-stepping.
 New results are described in section 3.
@@
 These comparisons have also revealed that the implicit solve is a small part of the total cost of the advection.
-This is likely because the linear equation solvers in OpenFOAM are well optimised, our test cases are run in serial and there are cache misses associated with the explicit part of the advection which uses extended stencils.
-The multi-dimensional scheme uses cubic interpolants but is not overall a third-order accurate.
-The multi-dimensional scheme is designed for arbitrary meshes.
-In order to achieve third-order accuracy on arbitrary meshes either multi-moments or Gauss points within each cell are needed.
-This would increase the complexity and cost significantly.
-The multi-dimensional scheme run explicitly has similar cost and complexity to the dimensionally split scheme.
+This is now discussed in section 3.4:
 
+\begin{quote}
+We also compared the real computational cost of the implicit and explicit versions of the multi-dimensional scheme running the deformational test case in serial on a non-orthogonal mesh with a maximum Courant number of 1.
+At the resolution of $120\times 60$, the model using implicit time-stepping is 5\%
+more expensive that the explicit model. This drops to 2.5\%
+at the resolution of $960\times 480$.
+Despite the implicit scheme having four times as many multiplies as calculated for table 3, the cost is insignificant.
+This demonstrates both the danger of only looking at the number of multiplies and the danger of using results of timing code.
+The linear equation solvers on OpenFOAM have undergone years of optimisation whereas the stencils for calculating the cubic fit have been developed less and entail unpredictable memory access patterns and likely frequent cache misses.
+\end{quote}
+
+We have added some description of why we think it suitable to compare the dimensionally split and multi-dimensional schemes at the beginning of section 2:
+\begin{quote}
+The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the temporal order to two.
+The multi-dimensional scheme uses cubic interpolations but uses cell centoids as Gauss points so the order of accuracy is limited to two.
+When run explicitly, the cost of both schemes is similar (see section 3.4).
+These schemes are therefore considered suitable for comparison.
+\end{quote}
+
+Several additions have been made as we are now using implicit and explicit time-stepping for the multi-dimensional scheme:
+
+\begin{itemize}
+\item At the beginning of section 2:
+\begin{quote}
+an explicit, second-order Runge-Kutta time-stepping scheme and with
+\end{quote}
+
+\item
+Section 2.3.1 has been split into two, the first to describe the implicit and explicit time-stepping and the second to describe the matrix solvers and tolerances.
+
+\item Added to section 3.1:
+\begin{quote}
+The results in figure 3 use the implicit time-stepping but the results using explicit time-stepping are similar. 
+\end{quote}
+and
+\begin{quote}
+The explicit multi-dimensional scheme is not stable for this time-step.
+\end{quote}
+and
+\begin{quote}
+Results of the explicit multi-dimensional scheme are also shown in figure 5 for the shorter time-step on the non-orthogonal mesh.
+The explicit RK2 scheme is more accurate than the implicit (Crank-Nicolson) scheme with the same order of convergence but does not come close to the accuracy of the dimensionally split scheme.
+\end{quote}
+and
+\begin{quote}
+The explicit version of the multi-dimensional scheme has errors that decrease as time-step increases due to the spatial discretisation errors that are being introduced at every time-step. 
+\end{quote}
+
+\item 
+Added to section 3.2:
+\begin{quote}
+The multi-dimensional scheme uses implicit time-stepping.
+\end{quote}
+
+\item 
+Added to section 3.3:
+\begin{quote}
+(with explicit time-stepping although the results using implicit time-stepping look identical)
+\end{quote}
+and
+\begin{quote}
+The explicit multi-dimensional scheme is stable at this Courant number and errors are almost identical to those of the implicit scheme.
+\end{quote}
+and
+\begin{quote}
+Again this shows that the explicit time-stepping gives almost identical errors to the implicit time-stepping at time-steps at which it is stable.
+\end{quote}
+
+Added to the conclusions:
+\begin{quote}
+Computational cost is estimated by the number of multiplies.\\
+...\\
+However in practice we found that the implicit solution contributed only 2.5\%-5\% of the over-all cost of the multi-dimensional model.
+\end{quote}
+\end{itemize}
 \end{description}
@@
 
-\item [Response]: Quasi-uniform resolution indeed help better exploit modern architectures.
-The lat-lon grid leads to convergence of resolution toward poles and restricts the time stepping.
-These problems are usually solved by semi-implicit semi-Lagrangian schemes.
-The modern computers use increasing number of cores and the efficiency is limited by the scalability.
-Implicit solvers and semi-Lagrangian advection result in massive data communications on lat-lon grids \citep{ST12}, which limits the scalability around poles.
-
-I know what you mean but, in reality, the modern computer architectures really are driving the need for quasi-uniform meshes and this is well documented in the references that we cite.
-We have changed the sentence to:
+\item [Response]: Have added the following to the beginning of the introduction:
 \begin{quote}
-Weather and climate models are being developed on quasi-uniform meshes in order to equidistribute and reduce the storage needed for the domain of dependence of the model degrees of freedom and hence better exploit modern computers \cite[eg][]{WTC12,LUJ+14,ST12,SG11,KNK15}.
+Latitude-longitude meshes needed measures to enable long time-steps near the poles such as polar filtering, semi-implicit and semi-Lagrangian \cite[]{DCM+05}.
+These methods necessarily have large domains of dependence near the poles which leads to poor parallel scaling at high resolution.
+Quasi-uniform meshes avoid parallel scaling bottlenecks near the poles but lead to non-orthogonal meshes (eg the cubed-sphere) or different structures (eg icosahedra).
 \end{quote}
@@
 
-\item [Response]: I was thinking about schemes like SLICE and CSLAM here but on reflection it is not best to discuss these schemes at this point so this comment has been removed.
+\item [Response]: Was thinking about schemes like SLICE and CSLAM here but on reflection it is not best to discuss these schemes at this point so this comment has been removed.
 
@@
 Their errors are certainly more uniform over time, but they are larger (using $90\times 90$ DOFs per panel) than \cite{GNQ14} (who used $80\times 80$ DOFs per panel).
+
+\item [Response]: See below
 
@@
 
-\item [Response]: Agreed. Added ``integration cost''
+\item [Response]: Agreed. Added ``integration cost'' and deleted:
+\begin{quote}
+This technique therefore offers no advantage over using a smaller time-step.
+\end{quote}
 
@@
 without smoothing of terrain following layers
+\end{quote}
+Also made a modification in the conclusions:
+\begin{quote}
+using a basic terrain following mesh in order to maximise the distortions \textcolor{blue}{due to terrain following coordinates}.
 \end{quote}
@@
 \begin{enumerate}
-\item [Major comment 1]: The first questions that come to my mind when reading this paper are, ``why another advection scheme?'' and ``why another advection test case?'' given the multitude of existing schemes and tests.
+\item [Major 1]: The first questions that come to my mind when reading this paper are, ``why another advection scheme?'' and ``why another advection test case?'' given the multitude of existing schemes and tests.
 The demonstration of the good accuracy of standard dimensionally-split FFSL schemes is certainly useful, especially considering the well-publicized shortcomings of such methods.
@@
 \item [Response]: 
-Our initial motivation was to demonstrate the large errors associated with dimensionally split schemes on non-orthogonal meshes with discontinuities in co-ordinate transforms which we thought had not previously been demonstrated in the literature.
+Our original motivation was to demonstrate the large errors associated with dimensionally split schemes on non-orthogonal meshes with discontinuities in co-ordinate transforms which we thought had not previously been demonstrated in the literature.
 We could not find these large errors despite comparing results on orthogonal and non-orthogonal meshes.
+Therefore an important result from this paper is that the dimensionally split scheme loses very little
+accuracy when grid distortions are introduced.
+This result does not rely on the accuracy of the multi-dimensional scheme.
+Modifications to the manuscript have been made in order to make this point, as described in responses to reviewer 1's comment 3.
 
-Accuracy higher than second order is not straightforward for a multi-dimensional scheme suitable for arbitrary meshes.
-The complexity and cost of the multi-dimensional scheme is similar to that of the dimensionally split scheme.
-However, in fact, neither scheme has accuracy higher than second order although they both have aspects that are third-order accurate.
-The dimensionally split scheme is third or forth order accurate in one dimension but the COSMIC splitting reduces the temporal order of accuracy even on an orthogonal grid.
-We thought that discontinuities in the coordinate system would reduce the order of accuracy to first but it appears to be second order.
-The multi-dimensional scheme uses cubic interpolations but since we use the cell centre as the Gauss point for each cell, the multi-dimensional scheme is limited to second order accuracy.
-We have added this text to the end of the first paragraph of section 2 and similar wording to the beginning of the conclusions:
-\begin{quote}
-The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the temporal order to two.
-The multi-dimensional scheme uses cubic interpolations but uses cell centoids as Gauss points so the order of accuracy is limited to two.
-When run explicitly, the cost of both schemes is similar (see section 3.4).
-These schemes are therefore considered suitable for comparison.
-\end{quote}
-
-\item [Major comment 2]: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
+\item [Major  2]: It is also questionable how useful it is to compare advection schemes without any form of monotonicity preservation, which is a necessity for modern solvers.
 The effect of this is most visible in Figure 4, where the more accurate dimensionally-split scheme has more grid-scale noise than the lower-order multi-dimensional scheme, likely because of the larger implicit diffusion in the lower-order scheme.
@@
 
-\item [Minor Comment 1]: Given that a fundamentally new scheme is not presented here, a better title might be something like, ``Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps''
+\item [Minor  1]: Given that a fundamentally new scheme is not presented here, a better title might be something like, ``Comparison of dimensionally-split and multi-dimensional atmospheric transport schemes for long time-steps''
 
-\item [Response]: Agreed, thanks.
+\item [Response]: Agreed, done.
 
-\item [Minor Comment 2]: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
+\item [Minor  2]: Several references to Lin and Rood (1996) should be to Putman and Lin (2007), since only PL07 uses the cubed-sphere.
 LR96 is designed for cartesian and lat-lon grids.
@@
 
-\item [Minor Comment 3]: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
+\item [Minor  3]: Figure 8: Is the variation in grid spacing done while keeping the width of the orography constant?
 
@@
 
-\item [Minor Comment 4]: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
+\item [Minor  4]: Section 3.4: The discussion of computational cost of the algorithms is a very useful addition to the paper. The brief discussion of 3D extensions is also very useful.
 
 \item [Response]: Thank-you. Following a request from reviewer 1, we also now include a comparison of wall clock time for the multi-dimensional scheme run implicitly and explicitly in serial. 
+\end{enumerate}
+
+\section*{Other revisions}
+
+There have been some other changes that have come up based on proof reading and code updates made in other projects.
+\begin{enumerate}
+\item Also added git repository for the setup of the test cases that are run using OpenFOAM.
+\item The test cases have been re-run (without changing the results) using an updated version of the code. Since we are running with a slightly newer version, we have added:
+\begin{quote}
+The stabilisation procedures described by \cite{SWMD17} are in the code but are not activated for these test cases because the meshes are sufficiently regular.
+\end{quote}
+
+\item Modified the first sentence of section 3.2 to remove uncertainty:
+\begin{quote}
+Non-orthogonal meshes (or co-ordinate systems) are usually necessary for representing orography with terrain following coordinates. This is a challenge for dimensionally split schemes.
+\end{quote}
+
+\item At the end of 3.3:
+\begin{quote}
+The multi-dimensional scheme is competitive at small Courant numbers but the semi-Lagrangian nature of the split scheme means that \textcolor{blue}{errors are very low when the Courant number is} close to one.
+\end{quote}
+
+\item Two very similar paragraphs in the conclusions are combined.
+
 \end{enumerate}

### 2017-05-25 revision/makeFigs/deform.tex
Hilary's first draft of revisions and replies to reviewer comments

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{array}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
$t$ & Dimensionally split & Multi-dimensional\tabularnewline
1 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_1_T}\tabularnewline
2 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_2_T}\tabularnewline
3 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T3} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_3_T}\tabularnewline
4 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T4} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_4_T}\tabularnewline
5 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T5} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_5_T}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_legends_T}}\tabularnewline
\end{tabular*}
\end{document}

### 2017-05-25 revision/makeFigs/deform_dx.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/deform.tex
@@
 \setlength{\parindent}{0pt}
-\usepackage{array}
 \usepackage{amssymb}
@@
 \begin{document}
-\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.03\textwidth}cc}
-$t$ & Dimensionally split & Multi-dimensional\tabularnewline
-1 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T1} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_1_T}\tabularnewline
-2 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T2} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_2_T}\tabularnewline
-3 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T3} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_3_T}\tabularnewline
-4 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T4} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_4_T}\tabularnewline
-5 & \includegraphics[width=0.49\textwidth,valign=m]{../YumengsGraphics/deformationalData/plots/T5} & \includegraphics[width=0.49\textwidth,valign=m]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_480x240_c1_explicit_5_T}\tabularnewline
-\multicolumn{3}{c}{\includegraphics[width=0.98\textwidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_legends_T}}\tabularnewline
-\end{tabular*}
+\begin{tabular}{cc}
+$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
+\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
+$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
+\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
+\end{tabular}
 \end{document}

### 2017-05-25 revision/makeFigs/deform_init.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/deform_dx.tex
@@
 \makeatletter
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
-%% Because html converters don't know tabularnewline
-\providecommand{\tabularnewline}{\\}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
@@
 \begin{document}
-\begin{tabular}{cc}
-$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
-\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
-$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
-\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
-\end{tabular}
+\includegraphics[width=1\textwidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_nonOrthogW_120x60_c1_implicit_0_TUmesh}
+
+\includegraphics[width=1\textwidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_legends_TUmesh_T}
 \end{document}

### 2017-05-25 revision/makeFigs/overMountains.tex
Hilary's first draft of revisions and replies to reviewer comments

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,british,times]{article}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[pdftex,a4paper]{geometry}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\pagestyle{empty}
\setcounter{secnumdepth}{5}
\setcounter{tocdepth}{5}
\setlength{\parskip}{\smallskipamount}
\setlength{\parindent}{0pt}
\usepackage{array}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage { fancybox}
\usepackage[pdftex]{graphicx,color}
\usepackage{rotating}
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}
\usepackage{amsmath}

\makeatother

\usepackage{babel}
\begin{document}
\begin{tabular*}{1\linewidth}{@{\extracolsep{\fill}}>{\centering}m{0.14\textwidth}>{\centering}p{0.42\textwidth}c}
 & Dimensionally split & Multi-dimensional\tabularnewline
$\Delta t=25\ $s

max $c$ = 0.74

horiz $c$ =0.25

$c_{d}=0.22$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt25CN_up_10000_Tall}\tabularnewline
$\Delta t=50\ $s

max $c$ = 1.48

horiz $c$ =0.5

$c_{d}=0.44$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt50CN_up_10000_Tall}\tabularnewline
$\Delta t=100\ $s

max $c$ = 2.96

horiz $c$ =1

$c_{d}=0.87$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt100CN_up_10000_Tall}\tabularnewline
$\Delta t=200\ $s

max $c$ = 5.93

horiz $c$ =2

$c_{d}=1.75$ & \includegraphics[bb=0bp 0bp 998bp 442bp,clip,width=0.39\textwidth,height=0.2\textwidth,valign=m]{../YumengsGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall} & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt200CN_up_10000_Tall}\tabularnewline
$\Delta t=500\ $s

max $c$ = 14.8

horiz $c$ =5

$c_{d}=4.4$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt500CN_up_10000_Tall}\tabularnewline
$\Delta t=1000\ $s

max $c$ = 29.6

horiz $c$ =10

$c_{d}=8.7$ & unstable & \includegraphics[width=0.42\textwidth,valign=m]{../HilarysGraphics/vSlice_horizontalAdvection_orography_save_cubicUpwindCPCFit_dt1000CN_up_10000_Tall}\tabularnewline
\multicolumn{3}{c}{\includegraphics[width=0.85\textwidth]{../HilarysGraphics/vSlice_horizontalAdvection_orography_legends_Tall_TdiffAll}}\tabularnewline
\end{tabular*}
\end{document}

### 2017-05-25 revision/makeFigs/sbr_dt.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/deform_dx.tex
@@
 \begin{document}
+\begin{center}
 \begin{tabular}{cc}
-$\ell_{2}$ error, $c\le1$ & $\ell_{\infty}$ error, $c\le1$\tabularnewline
-\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c1} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c1}\tabularnewline
-$\ell_{2}$ error, $c\le10$ & $\ell_{\infty}$ error, $c\le10$\tabularnewline
-\includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_l2errors_c10} & \includegraphics[width=0.48\linewidth]{../HilarysGraphics/multiDadvectCases_deformingAdvectionCyclic_plots_linferrors_c10}\tabularnewline
+$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
+\includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
 \end{tabular}
+\par\end{center}
 \end{document}

### 2017-05-25 revision/makeFigs/sbr_dx.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/sbr_dt.tex
@@
 \begin{tabular}{cc}
-$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
-\includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
+$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx1$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx1$\tabularnewline
+\includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errorsC1} & \includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrorsC1}\tabularnewline
+$\ell_{2}$ errors, convergence with $\Delta x$ for $c\approx10$ & $\ell_{\infty}$ errors, convergence with $\Delta x$ for $c\approx10$\tabularnewline
+\includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errorsC10} & \includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrorsC10}\tabularnewline
 \end{tabular}

### 2017-05-25 revision/makeFigs/sbrc1.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/sbr_dt.tex
@@
 \begin{document}
-\begin{center}
 \begin{tabular}{cc}
-$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
-\includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
+Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
+\includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
+Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
+\includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_orthogonal_100x100_dt_1_Tcontours} & \includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_nonOrthog_100x100_dt_1_Tcontours}\tabularnewline
+\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_legends_Tdiff}}\tabularnewline
 \end{tabular}
-\par\end{center}
 \end{document}

### 2017-05-25 revision/makeFigs/sbrc10.tex
Hilary's first draft of revisions and replies to reviewer comments

--- 2017-05-25 revision/makeFigs/sbr_dt.tex
@@
 \begin{document}
-\begin{center}
 \begin{tabular}{cc}
-$\ell_{2}$ errors & $\ell_{\infty}$ errors\tabularnewline
-\includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_l2errors_dt} & \includegraphics[width=0.49\textwidth]{/mnt/homer2/hilary/latex/myPapers/YumengHilaryAdvectionPaper/revision/HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_plots_linferrors_dt}\tabularnewline
+Orthogonal, dimensionally split scheme & Non-orthogonal, dimensionally split scheme\tabularnewline
+\includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{../YumengsGraphics/solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
+Orthogonal, multi-dimensional scheme & Non-orthogonal, multi-dimensional scheme\tabularnewline
+\includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_orthogonal_100x100_dt_10_TcontoursBig} & \includegraphics[width=0.49\textwidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_nonOrthog_100x100_dt_10_TcontoursBig}\tabularnewline
+\multicolumn{2}{c}{\includegraphics[width=0.9\linewidth]{../HilarysGraphics/multiDadvectCases_solidBodyRotationOnPlane_legends_TdiffBig_Tdiff}}\tabularnewline
 \end{tabular}
-\par\end{center}
 \end{document}

### 2017-05-25 revision/getGraphics.sh

@@ -5,17 +5,17 @@ ROOTDIR=/home/hilary/OpenFOAM/*/run
 
 #List of files to convert to png
 pngFiles=(
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/0/T.pdf
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/1/T.pdf
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/2/T.pdf
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/3/T.pdf
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/4/T.pdf
-deformingAdvectionCyclic/nonOrthogW_c1/480x240/5/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/0/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/1/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/2/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/3/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/4/T.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/480x240/c1_explicit/5/T.pdf
 )
 
 # List of files to copy and rename
 cpFiles=(
-deformingAdvectionCyclic/nonOrthogW_c1/120x60/0/TUmesh.pdf
+multiDadvectCases/deformingAdvectionCyclic/nonOrthogW/120x60/c1_implicit/0/TUmesh.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt1000CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt100CN_up/10000/Tall.pdf
 vSlice/horizontalAdvection/orography/save/cubicUpwindCPCFit_dt200CN_up/10000/Tall.pdf
@@ -36,32 +36,32 @@ vSlice/horizontalAdvection/highOrography_coarse/save/cubicUpwindCPCFit_dt50N_up/
 vSlice/horizontalAdvection/highOrography_fine/save/cubicUpwindCPCFit_dt12.5N_up/10000/TallTall.pdf
[…]

### 2017-05-25 revision/makeFigs.sh

@@ -2,20 +2,24 @@
 cd makeFigs
 
 for file in stencils sbrc1 sbrc10 sbr_dx sbr_dt overMountains overMountains_dx deform_init deform deform_dx deform_dt; do
+    echo $file
     lyx --export pdflatex -f $file
     pdflatex $file
     pdfcrop $file.pdf ../figures/$file.pdf
     rm $file.pdf
-    evince ../figures/$file.pdf &
+    gv ../figures/$file.pdf &
 done
 
 cd ..
 
-cp -u HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf .
+cp -u HilarysGraphics/solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf figures
 
-eps2png2 deform.pdf
-eps2png2 sbrc10.pdf
-eps2png2 sbrc1.pdf
+eps2png figures/deform.pdf
+eps2png figures/sbrc10.pdf
+eps2png figures/sbrc1.pdf
+eps2png figures/overMountains.pdf
 
-zip figures.zip stencil.pdf solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf sbrc1.png sbrc10.png sbr_dx.pdf sbr_dt.pdf overMountains.png overMountains_dx.pdf deform_init.pdf deform.png deform_dx.pdf deform_dt.pdf
+rm figures/deform.pdf figures/sbrc10.pdf figures/sbrc1.pdf figures/overMountains.pdf
+
+#zip figures.zip stencil.pdf solidBodyRotationOnPlane_nonOrthog_50x50_analytic_constant_mesh.pdf sbrc1.png sbrc10.png sbr_dx.pdf sbr_dt.pdf overMountains.png overMountains_dx.pdf deform_init.pdf deform.png deform_dx.pdf deform_dt.pdf
 

### 2017-06-07 revision/reply_to_reviewer.tex
Incoroprated James's comments on the revised manuscript

@@
 
-\title{Responses to reviewer comments on artile QJ-17-0017\\
+\title{Responses to reviewer comments on article QJ-17-0017\\
 Dimension Splitting and a Long Time-Step Multi-Dimensional Scheme for Atmospheric Transport}
@@
 \item [Response]:
-Agreed.
-Our test cases on a plane will highlight errors due to non-orthogonality and due to discontinuities at cube edges but cannot help to describe errors at cube corners.
-Some sentences have been added to the introduction:
+Agreed. Our test cases on a plane will highlight errors due to non-orthogonality and due to discontinuities at cube edges. Some sentences have been added to the introduction:
 \begin{quote}
-We assume that a large fraction of the numerical errors associated with terrain following co-ordinates and the cubed-sphere grid are due to the mesh non-orthoganlity and distortion.
-However there will also be errors due to cubed-sphere corners and inaccurate representation of spherical geometry.
-These are not addressed here.
+We assume that a large fraction of the numerical errors associated with terrain following co-ordinates and the cubed-sphere grid are due to the mesh non-orthogonality and distortion.
+Other sources of numerical errors, such as those associated with the representation of spherical geometry, are not addressed here.
 \end{quote}
@@
 \begin{quote}
-These test cases will mimic the problems encountered at cubed-sphere edges but will not give any information about additional errors that might occur at cubed-sphere corners. 
+These test cases will mimic the problems encountered near cubed-sphere edges. 
 \end{quote}
@@
 \begin{quote}
-we show the effect of dimension splitting errors using a FFSL dimensionally split scheme on a number of test-cases which use distorted \textcolor{blue}{and un-distorted} meshes.
+we show the effect of dimension splitting errors using a FFSL dimensionally split scheme on a number of test-cases which use distorted \textcolor{blue}{and undistorted} meshes.
 \end{quote}
@@
 \begin{quote}
-We also compared the real computational cost of the implicit and explicit versions of the multi-dimensional scheme running the deformational test case in serial on a non-orthogonal mesh with a maximum Courant number of 1.
+We also measured the wall-clock computation cost, comparing the implicit and explicit versions of the multi-dimensional scheme running the deformational test case in serial on a non-orthogonal mesh with a maximum Courant number of 1.
 At the resolution of $120\times 60$, the model using implicit time-stepping is 5\%
@@
 at the resolution of $960\times 480$.
-Despite the implicit scheme having four times as many multiplies as calculated for table 3, the cost is insignificant.
+Despite the implicit scheme having four times as many multiplies as calculated for table 3, the difference in wall-clock time is negligible.
 This demonstrates both the danger of only looking at the number of multiplies and the danger of using results of timing code.
-The linear equation solvers on OpenFOAM have undergone years of optimisation whereas the stencils for calculating the cubic fit have been developed less and entail unpredictable memory access patterns and likely frequent cache misses.
+The linear equation solvers on OpenFOAM have undergone years of optimisation whereas the stencils for calculating the cubic fit have not been optimised and entail unpredictable memory access patterns and likely frequent cache misses.
 \end{quote}
@@
 The dimensionally split scheme is third-order accurate in one dimension but the COSMIC splitting reduces the temporal order to two.
-The multi-dimensional scheme uses cubic interpolations but uses cell centoids as Gauss points so the order of accuracy is limited to two.
-When run explicitly, the cost of both schemes is similar (see section 3.4).
+The multi-dimensional scheme uses cubic interpolations but uses cell centroids as Gauss points so the order of accuracy is limited to two.
+When explicit time-stepping is used, the cost of both schemes is similar (see section 3.4).
 These schemes are therefore considered suitable for comparison.
@@
 \begin{quote}
-The results in figure 3 use the implicit time-stepping but the results using explicit time-stepping are similar. 
+The results in figure 3 use implicit time-stepping but the results using explicit time-stepping are similar (not shown). 
 \end{quote}
@@
 \begin{quote}
-Again this shows that the explicit time-stepping gives almost identical errors to the implicit time-stepping at time-steps at which it is stable.
+Again this shows that the explicit time-stepping gives almost identical errors to the implicit time-stepping for time-steps at which it is stable.
 \end{quote}
@@
 ...\\
-However in practice we found that the implicit solution contributed only 2.5\%-5\% of the over-all cost of the multi-dimensional model.
+However in practice we found that the implicit solution contributed only 2.5\%-5\% of the total wall-clock computation time.
 \end{quote}
@@
 
-\item [Response]: Have added the following to the beginning of the introduction:
+\item [Response]: Have changed the beginning of the introduction to:
 \begin{quote}
-Latitude-longitude meshes needed measures to enable long time-steps near the poles such as polar filtering, semi-implicit and semi-Lagrangian \cite[]{DCM+05}.
+Many traditional weather and climate models use latitude-longitude meshes, but new models are being developed on quasi-uniform meshes in order to better exploit modern computers \cite[eg][]{WTC12,LUJ+14,ST12,SG11,KNK15}.
+Latitude-longitude meshes need measures to enable long time-steps near the poles such as polar filtering, semi-implicit and semi-Lagrangian methods \cite[]{DCM+05}.
 These methods necessarily have large domains of dependence near the poles which leads to poor parallel scaling at high resolution.
@@
 \begin{quote}
-Dimensionally split schemes have been found to give good accuracy on non-orthogonal meshes such as the cubed-sphere if special treatment is applied over cube edges to improve accuracy: \cite{PL07} use the average of two one-sided schemes at cube edges.
-\cite{KNK15} create ghost cells outside each cube panel boundary and achieve accuracy between second and fourth order.
+Dimensionally split schemes have been found to give good accuracy on non-orthogonal meshes such as the cubed-sphere if special treatment is applied over cube edges to improve accuracy: for example, \cite{PL07} use the average of two one-sided schemes at cube edges.
 \cite{KNK15} use a multi-dimensional transport scheme with dimensionally split reconstructions.
-\cite[eg]{GNQ14} report error growth around cube edges using Strang splitting.
+They create ghost cells outside each cube panel boundary and achieve accuracy between second and fourth order.
+\cite{GNQ14} report error growth around cube edges using Strang splitting.
 \end{quote}
@@
 \begin{quote}
-The next step would be to use a second-order predictor-corrector departure point calculation \cite[eg][]{MDW+10}.
+The departure point calculation could be improved by using a second-order predictor-corrector departure point calculation \cite[eg][]{MDW+10}.
 \end{quote}
@@
 \begin{quote}
-In contrast, the multi-dimensional scheme is insensitive to the orthogonality, converges with second-order \textcolor{blue}{as expected} and suffers from phase errors at large Courant numbers \textcolor{blue}{due to the use of Crank-Nicolson implicit time-stepping}.
+The multi-dimensional scheme is insensitive to the orthogonality, converges with second-order \textcolor{blue}{as expected} and suffers from phase errors at large Courant numbers \textcolor{blue}{due to the use of Crank-Nicolson implicit time-stepping}.
 \end{quote}
